### 2015-01-28 Programming Project A/Provided Files/LaTex Template/sample.tex
initial commit

%% Based on a TeXnicCenter-Template by Tino Weinkauf.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% HEADER
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\documentclass[onecolumn, 11pt, conference, compsocconf]{IEEEtran}
\documentclass[onecolumn, 12pt, article]{IEEEtran}
\usepackage{times,color,amsmath,amssymb,amsthm,comment,graphicx,cite,multirow}
%\documentclass[letterpaper,oneside,11pt]{letter}
% Alternative Options:
%	Paper Size: a4paper / a5paper / b5paper / letterpaper / legalpaper / executivepaper
% Duplex: oneside / twoside
% Base Font Size: 10pt / 11pt / 12pt

%% Language %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[USenglish]{babel} %francais, polish, spanish, ...
\usepackage[T1]{fontenc}
\usepackage[ansinew]{inputenc}

\usepackage{lmodern} %Type1-font for non-english texts and characters
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{float}

%% Packages for Graphics & Figures %%%%%%%%%%%%%%%%%%%%%%%%%%
%\usepackage{graphicx} %%For loading graphic files
%\usepackage{subfig} %%Subfigures inside a figure
%\usepackage{tikz} %%Generate vector graphics from within LaTeX

%% Please note:
%% Images can be included using \includegraphics{filename}
%% resp. using the dialog in the Insert menu.
%% 
%% The mode "LaTeX => PDF" allows the following formats:
%%   .jpg  .png  .pdf  .mps
%% 
%% The modes "LaTeX => DVI", "LaTeX => PS" und "LaTeX => PS => PDF"
%% allow the following formats:
%%   .eps  .ps  .bmp  .pict  .pntg

%% Math Packages %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
%\usepackage{array,MnSymbol}

%% Line Spacing %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{setspace}
\singlespacing        %% 1-spacing (default)
%\onehalfspacing       %% 1,5-spacing
%\doublespacing        %% 2-spacing

%% Other Packages %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\usepackage{a4wide} %%Smaller margins = more text per page.
\usepackage{fancyhdr} %%Fancy headings
\usepackage{listings}

%
% Theorem like environments
%
\newtheorem{problem}{Problem}%
%\numberwithin{problem}{section}
\newtheorem{theorem}{Theorem}%
\newtheorem{acknowledgment}{Acknowledgment}%
%\newtheorem{algorithm}{Algorithm}%
\newtheorem{assumption}{Assumption}%
\newtheorem{axiom}{Axiom}%
\newtheorem{case}{Case}%
\numberwithin{case}{problem}
\newtheorem{claim}{Claim}%
\newtheorem{conclusion}{Conclusion}
\newtheorem{condition}{Condition}
\numberwithin{condition}{problem}
\numberwithin{condition}{subsection}
\newtheorem{conjecture}{Conjecture}
\newtheorem{corollary}{Corollary}
\newtheorem{criterion}{Criterion}
\newtheorem{definition}{Definition}
\numberwithin{definition}{section}
\newtheorem{example}{Example}
\newtheorem{exercise}{Exercise}%
\newtheorem{lemma}{Lemma}%
\newtheorem{notation}{Notation}%
\theoremstyle{remark}
\newtheorem{question}{Question}%
\numberwithin{question}{problem}
\theoremstyle{plain}
\newtheorem{answer}{Answer}%
\numberwithin{answer}{problem}
\newtheorem{proposition}{Proposition}%
\newtheorem{remark}{Remark}%
\newtheorem{solution}{Solution}%
\numberwithin{solution}{section}
\newtheorem{summary}{Summary}%
\numberwithin{equation}{section}%
\newtheorem{option}{Option}

\raggedbottom
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Remarks
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% TODO:
% 1. Edit the used packages and their options (see above).
% 2. If you want, add a BibTeX-File to the project
%    (e.g., 'literature.bib').
% 3. Happy TeXing!
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Options / Modifications
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\input{options} %You need a file 'options.tex' for this
%% ==> TeXnicCenter supplies some possible option files
%% ==> with its-templates (File | New from Template...).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% DOCUMENT
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

%\pagestyle{empty} %No headings for the first pages.

%% Title Page %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% ==> Write your text here or include other files.

%% The simple version:
\title{Meaningful Title for this Paper}
\author{John Doe}
\date{} %%If commented, the current date is used.
\maketitle
%\fancyhead()

\pagestyle{fancy} %Now display headings: headings / fancy / ...
\fancyhead[R]{Insertion Sort, page \thepage}
\fancyhead[L]{John Doe}

\begin{abstract}

The abstract is written \emph{after} the report and gives a high--level overview.  An abstract is optional for these reports, but it is good practice.  Look up some published papers and read the abstracts to get an idea of how to write one.

\end{abstract}
\section{Motivation}
There has to be some good reasons to do this work and publish it.  Why?  Never refer to any report for this class as a class project, lab project, or assignment.  It is a report, a work, or an article.  Try to be professional.

\section{Background}
What exactly is a good description of the problem?  What work has been done on it?  In the past, others have reported algorithms to perform an Insertion Sort~\cite{cormen2001introduction}.  Remember:  If you copy it, it's plagiarism until you cite it.

\section{Procedure}
How did you do this wonderful thing anyway?  This is where you put your pseudocode (see Algorithm~\ref{prot:auth}), pre-conditions, post conditions, and Invariants.  Describe the idea in a general way.  How you implement your procedure goes in the Experimental Analysis.
\begin{algorithm}
\caption {\textsc{Build-Max-Heap}(A)}
\label{prot:auth}

\begin{algorithmic}[1]
\Procedure{Build-Max-Heap}{A}%\Comment{The g.c.d. of a and b}
\State{assert(A is not empty)}
\State {$heapsize = A.length$}
 \For {$i = \left\lfloor heapsize / 2 \right\rfloor  \text{\textbf{ downto }} 2$} 
     \State  Max-Heapify(A,$ i$)
    \State{assert(Heap-Test(A,$i$)}
 \EndFor
\State{assert(Heap-Test(A,$ 1$)}
\EndProcedure
\end{algorithmic}

\end{algorithm}

\section{Testing}
\subsection{Testing Plan and Results}
This can be done as a table.  Something like:\\\\
\small{\begin{tabular}{|l|l|l|l|}
\hline Test input & Expected Results & Actual Results & Corrective Actions \\ 
\hline  Empty array & nothing returned  & nothing was returned & no action taken  \\ 
\hline All elements identical & same array  & Program faulted & off--by--one error corrected  \\ 
\hline 
\end{tabular}}
\subsection{Problems Encountered}
Normally, you would not see these in a paper, but our projects are a special case.  You might see this in a paper that developed a simulator for running virtual experiments.  If you had no problems, state ``No significant problems were encountered.''  Syntax errors and typo's are not problems.  You can give a brief description of how you fixed the problem, for example ``Updating to the latest version of java corrected the issues with program faults.''  Use your best judgement on what is a problem and what is not.   If you got help with your problem, acknowledge it here or cite the help as a conversation\footnote{Look up how to cite conversations.  Also avoid footnotes.}.

\section{Experimental Analysis}
IF you wish, you can include a paragraph about the specifications of the platform you used to run your experiments.  Be very brief:  ``These timings were carried out on a 427 cubic inch supercharged Dell MT708 with dual overhead BlueRay drives.  The power of the CUP was wasted by running Windows 8.1 Pro with all the latest updates as of January 12, 1995.  Oracle Java 1.0 was used as the programming language/compiler.''

Here you need to graph your results and explain them.  This is the main part of your paper.   Graphs should be inserted as figures and referenced by ``...agreed well with the expected results, see Figure~\ref{fig:figureforprinting}.  Be careful of color-coding parts of graphs as most printing is still black and white.  Notice how I used both color and dashes to differentate the two lines on the graph.   You should be able to get away with one graph with two lines.  Mark $n_0$ on the graph if you can.

\begin{figure}[!]
\centering
\includegraphics[width=0.7\linewidth]{./figure.pdf}
\caption{A Sample Graph}
\label{fig:figureforprinting}
\end{figure}
Current style is to place tables \emph{before} they are referenced with the caption above the table.  Figures are place \emph{after} they are referenced and the caption is below the figure.  If you have time, read the IEEE Style Sheet in the same folder with this sample.

Explain clearly what your results mean.  If you have a result that makes no sense, clearly state why it makes no sense.  If you can, give a plausible explanation as to what may have cause the strange result.  Do not try to hide a bad result.

\section{Conclusions}
Here you answer any questions raised in the previous sections.   You should end with a wrap--up and firm statement of what a great accomplishment this is.

% references section
%\baselineskip=.55\baselineskip

\bibliographystyle{IEEEtran}
\bibliography{references}

\newpage
\section*{Appendix}
An Appendix starts on a new page after everything else.  This is were your code will be.\\\\

\lstinputlisting{fragment.java}
\end{document}

### 2015-01-28 Programming Project 1/TwoPowers.java

@@ -1,0 +1,23 @@
+/*
+*	@author: Preston Stosur-Bassett
+*	@date: 9.1.15
+*	@course: CS215 Algorithms
+*	@instructor: Dr. Howser
+*	@class: TwoPowers
+*	@description: TwoPowers prints the value of 2 raised to the power of 0-10.
+*/
+
+import java.lang.Math;
+
+public class TwoPowers {
+	public static void main(String args[]) {
+		System.out.println("Printing powers of 2 from 0-10...");
+		//Run this loop 10 times to print out all the values of 2 from the power of 0 to the power of 10
+		for(double i = 0; i < 11; i++) {
+			//Raise 2 to the power of the current iterator value
+			double ans = Math.pow(2, i);
+			//Print the calculated values
+			System.out.println("2^"+i+" ==> "+ans);
+		}
+	}
+}

### 2015-01-28 Programming Project A/Code/Debug.java

@@ -1,0 +1,41 @@
+/*
+*	@Author Preston Stosur-Bassett
+*	@Date Jan 21, 2015
+*	@Class Debug
+*	@Description This class will help debugging by being able to turn on and turn off debug messages easily
+*/
+
+public class Debug {
+	boolean debugOn; //Variable to keep track of whether or not debug is on
+
+	/*
+	*	@Description constructor method that sets the default value of debugOn to false so that debug statements will not automatically print
+	*/
+	public void Debug() {
+		debugOn = false;
+	}
+
+	/*
+	*	@Description turn on debugging print statements
+	*/
+	public void turnOn() {
+		debugOn = true;
+	}
+
+	/*
+	*	@Description turn off debugging print statements
+	*/
+	public void turnOff() {
+		debugOn = false;
+	}
+
+	/*
+	*	@Description will print messages only when debugOn boolean is set to true
+	*	@param String message the string to print when debugging is turned on
+	*/
+	public void print(String message) {
+		if(debugOn == true) {
+			System.out.println(message);
+		}
+	}
+}

### 2015-01-28 Programming Project A/Code/DummyData.java

@@ -1,0 +1,33 @@
+/*
+*	@Author Preston Stosur-Bassett
+*	@Date Jan 25, 2015
+*	@Class DummyData
+*	@Description This class contains methods to generate dummy data given a set of parameters. 
+*/
+
+import java.util.ArrayList;
+
+public class DummyData {
+
+	/*
+	*	@Description runArrayList will take an ArrayList of Integer Objects and add a given amount of values to it
+	*	@param int start the starting value to add to the array list
+	*	@param int end the ending value to denote when to stop adding to the array list
+	*	@param int iterator the amount to increase <code>start</code> variable by after each iteration
+	*	@param ArrayList<Integer> list the list to add value to and return
+	*	@return ArrayList<Integer> the list after it has been updated with the given data
+	*/
+	public static ArrayList<Integer> runArrayList(int start, int end, int iterator, ArrayList<Integer> list) {
+		int counter = start;
+		while(start < end) {
+			Integer intToAdd = new Integer(counter);
+			list.add(intToAdd);
+
+			//Count up on the iterators
+			counter++;
+			start = start + iterator;
+		}
+
+		return list;
+	}
+}

### 2015-01-28 Programming Project A/Code/Sort.java

@@ -1,0 +1,33 @@
+/*
+*	@Author: Preston Stosur-Bassett
+*	@Date: Jan 24, 2015
+*	@Class: Sort
+*	@Description: This class will contain many methods that will sort generic data types using common sorting algorithms.
+*/
+
+import java.util.ArrayList;
+
+public class Sort <T extends Comparable> {
+	/*
+	*	@Pre-Condition: ArrayList<T> unsorted is an unsorted ArrayList of a comparable data type that is non-empty
+	*	@Post-Condition: ArrayList<T> will return a permutation of <code>unsorted</code> that will be in increasing order
+	*	@Description: Insertion will sort an ArrayList of generic type T in increasing order using an insertion sort
+	*	@param ArrayList<T> unsorted is a non-empty unsorted array list of T, where T is a comparable type
+	*	@return sorted is a permutation of <code>unsorted</code> where all the elements are sorted in increasing order
+	*/
+	public static <T> ArrayList<T> insertion(ArrayList<T> unsorted) {
+		ArrayList<T> sorted = unsorted;
+		if(sorted.size() > 1) {
+			for(int i = 1; i < sorted.size(); i++) {
+				T value = sorted.get(i);
+				int j = i - 1;
+				while((value.compareTo(sorted.get(j)) < 0) && j > 0) {
+					sorted.set(j+1, sorted.get(j));
+					j--;
+				}
+				sorted.set(j+1, value);
+			}
+		}
+		return sorted;
+	}
+}

### 2015-01-28 Programming Project A/Code/TestDriver.java

@@ -1,0 +1,35 @@
+/*
+*	@Author Preston Stosur-Bassett
+*	@Date Jan 25, 2015
+*	@Class TestDriver
+*	@Description this class will create test data to run through and test the other classes in this directory.
+*/
+
+import java.util.ArrayList;
+
+public class TestDriver {
+	public static void main(String args[]) {
+		//Turn on debugger
+		Debug debugger = new Debug();
+		debugger.turnOn();
+
+		//Test debug
+		debugger.print("Debug is on.");
+
+		//Create test data for sorts class
+		ArrayList<Integer> testList = new ArrayList<Integer>();
+		Integer oneOhone = new Integer(101);
+		Integer oneEleven = new Integer(111);
+		Integer oneOhOhFive = new Integer(1005);
+		testList = DummyData.runArrayList(0, 10, 2, testList);
+
+		System.out.println("Un-Sorted List");
+		System.out.println(testList);
+
+		//Test sort List
+		testList = Sort.insertion(testList);
+
+		System.out.println("Sorted List");
+		System.out.println(testList);
+	}
+}

### 2015-01-28 Programming Project A/Provided Files/LaTex Template/fragment.java

@@ -1,0 +1,29 @@
+  //Sample routine with assert statements
+  public void Build_Max_Heap(<T> A[])
+   {
+//Author:  			G Howser
+//Purpose:  		Builds a Max Heap in A[] with root A[1]
+//Heap_Test:		Returns "true" iff A[j] Max Heap for j, j+1, j+2, ... n
+//
+//Pre condition:	A[] is a non-empty array of integers
+//Invariant:		A[j] is root of a Max Heap for j=i+1, i+2, ... n
+    if (debug)
+     {
+      assert (((A[1] > 0) || (A[1] <= 0)));
+     };
+    int heap_size;
+    int n = A.length - 1;
+    heap_size = A.length - 1; //NOTE:  zero relative arrays but we are ignoring A[0]
+    for (int i = Math.floor((n / 2.0)); i--; i < 1)
+     {
+      Max_Heapify(A, i);
+      if (debug)
+       {
+        assert ((Heap_Test(A[], i));
+       };
+//Post condition:  A[] is a Max Heap
+      if (debug)
+       {
+        assert ((Heap_Test(A[], 1));
+       };
+	 }

### 2015-01-28 Programming Project A/Code/Debug.java
updated debug class

@@ -5,7 +5,7 @@
 *	@Description This class will help debugging by being able to turn on and turn off debug messages easily
 */
 
-public class Debug {
+public class Debug<T> {
 	boolean debugOn; //Variable to keep track of whether or not debug is on
 
 	/*
@@ -36,6 +36,19 @@ public class Debug {
 	public void print(String message) {
 		if(debugOn == true) {
 			System.out.println(message);
+		}
+	}
+
+	/*
+	*	@Pre-Condition <code>T expected</code> and <code>T actual</code> are both of the same type T
+	*	@Post-Condition If <code>T expected</code> and <code>T actual</code> are found to be equal, the program moves on, otherwise the program halts with <code>AssertionError</code> is thrown
+	*	@Description runs an assert statement against an expected value and the actual value that are passed as parameters only when <code>debugOn == true</code>
+	*	@param T expected the expected value to assert against the actual value
+	*	@param T actualt he actual value to assert against the expected value
+	*/
+	public void assertIt(T expected, T actual) {
+		if(debugOn == true) {
+			assert actual == expected;
 		}
 	}
 }

### 2015-01-28 Programming Project A/Code/Sort.java
fixed issue with generic compareto

@@ -7,7 +7,7 @@
 
 import java.util.ArrayList;
 
-public class Sort <T extends Comparable> {
+public class Sort<T extends Comparable<T>> {
 	/*
 	*	@Pre-Condition: ArrayList<T> unsorted is an unsorted ArrayList of a comparable data type that is non-empty
 	*	@Post-Condition: ArrayList<T> will return a permutation of <code>unsorted</code> that will be in increasing order
@@ -15,7 +15,7 @@ public class Sort <T extends Comparable> {
 	*	@param ArrayList<T> unsorted is a non-empty unsorted array list of T, where T is a comparable type
 	*	@return sorted is a permutation of <code>unsorted</code> where all the elements are sorted in increasing order
 	*/
-	public static <T> ArrayList<T> insertion(ArrayList<T> unsorted) {
+	public ArrayList<T> insertion(ArrayList<T> unsorted) {
 		ArrayList<T> sorted = unsorted;
 		if(sorted.size() > 1) {
 			for(int i = 1; i < sorted.size(); i++) {

### 2015-01-28 Programming Project A/Code/TestDriver.java

@@ -21,13 +21,17 @@ public class TestDriver {
 		Integer oneOhone = new Integer(101);
 		Integer oneEleven = new Integer(111);
 		Integer oneOhOhFive = new Integer(1005);
+		testList.add(oneOhone);
+		testList.add(oneEleven);
+		testList.add(oneOhOhFive);
 		testList = DummyData.runArrayList(0, 10, 2, testList);
 
 		System.out.println("Un-Sorted List");
 		System.out.println(testList);
 
 		//Test sort List
-		testList = Sort.insertion(testList);
+		Sort sorter = new Sort<Integer>();
+		testList = sorter.insertion(testList);
 
 		System.out.println("Sorted List");
 		System.out.println(testList);

### 2015-01-28 Programming Project A/Code/Driver.java
added official driver class

@@ -1,0 +1,17 @@
+/*
+* @Author Preston Stosur-Bassett
+* @Date Jan 27, 2015
+* @Description This is the main driver for this program. It will officially test the functionality of the other classes in this program.
+* @Note: It is important to note that TestDriver.java tests the functtionality of each class for debugging purposes only. This is the official driver.
+* @Class: Driver
+*/
+
+public class Driver {
+
+  /*
+  * @Description: This is the main driver for the program to show functionality for the entire program
+  */
+  public static void main(String[] args) {
+    //Do things here
+  }
+}

### 2015-01-28 Programming Project A/Code/Debug.java
working version and everything

@@ -33,7 +33,7 @@ public class Debug<T> {
 	*	@Description will print messages only when debugOn boolean is set to true
 	*	@param String message the string to print when debugging is turned on
 	*/
-	public void print(String message) {
+	public void print(T message) {
 		if(debugOn == true) {
 			System.out.println(message);
 		}

### 2015-01-28 Programming Project A/Code/DummyData.java

@@ -2,10 +2,11 @@
 *	@Author Preston Stosur-Bassett
 *	@Date Jan 25, 2015
 *	@Class DummyData
-*	@Description This class contains methods to generate dummy data given a set of parameters. 
+*	@Description This class contains methods to generate dummy data given a set of parameters.
 */
 
 import java.util.ArrayList;
+import java.util.Random;
 
 public class DummyData {
 
@@ -13,19 +14,19 @@ public class DummyData {
 	*	@Description runArrayList will take an ArrayList of Integer Objects and add a given amount of values to it
 	*	@param int start the starting value to add to the array list
 	*	@param int end the ending value to denote when to stop adding to the array list
-	*	@param int iterator the amount to increase <code>start</code> variable by after each iteration
+	*	@param int min the minimum value of the randomly generated data.
+	*	@param int max the maximum value of the randomly generated data.
 	*	@param ArrayList<Integer> list the list to add value to and return
 	*	@return ArrayList<Integer> the list after it has been updated with the given data
 	*/
-	public static ArrayList<Integer> runArrayList(int start, int end, int iterator, ArrayList<Integer> list) {
-		int counter = start;
+	public static ArrayList<Integer> runArrayList(int start, int end, int min, int max, ArrayList<Integer> list) {
+		Random random = new Random();
 		while(start < end) {
-			Integer intToAdd = new Integer(counter);
+			Integer intToAdd = new Integer(random.nextInt((max - min + 1) + min));
 			list.add(intToAdd);
 
 			//Count up on the iterators
-			counter++;
-			start = start + iterator;
+			start++;
 		}
 
 		return list;

### 2015-01-28 Programming Project A/Code/Sort.java

@@ -21,7 +21,7 @@ public class Sort<T extends Comparable<T>> {
 			for(int i = 1; i < sorted.size(); i++) {
 				T value = sorted.get(i);
 				int j = i - 1;
-				while((value.compareTo(sorted.get(j)) < 0) && j > 0) {
+				while(j >= 0 && (value.compareTo(sorted.get(j)) < 0)) {
 					sorted.set(j+1, sorted.get(j));
 					j--;
 				}

### 2015-01-28 Programming Project A/Code/TestDriver.java

@@ -18,22 +18,18 @@ public class TestDriver {
 
 		//Create test data for sorts class
 		ArrayList<Integer> testList = new ArrayList<Integer>();
-		Integer oneOhone = new Integer(101);
-		Integer oneEleven = new Integer(111);
-		Integer oneOhOhFive = new Integer(1005);
-		testList.add(oneOhone);
-		testList.add(oneEleven);
-		testList.add(oneOhOhFive);
-		testList = DummyData.runArrayList(0, 10, 2, testList);
+		testList = DummyData.runArrayList(0, 1000, 0, 100000, testList);
 
 		System.out.println("Un-Sorted List");
 		System.out.println(testList);
 
 		//Test sort List
 		Sort sorter = new Sort<Integer>();
+		Stopwatch watchStopper = new Stopwatch();
 		testList = sorter.insertion(testList);
 
 		System.out.println("Sorted List");
 		System.out.println(testList);
+		System.out.println("Time to complete: "+watchStopper.elapsedTime());
 	}
 }

### 2015-01-29 Programming Project A/Code/TestDriver.java
preparing for testing

@@ -8,6 +8,10 @@
 import java.util.ArrayList;
 
 public class TestDriver {
+
+	/*
+	*	@Description This serves as the Driver function for this program, run this class to execute the program
+	*/
 	public static void main(String args[]) {
 		//Turn on debugger
 		Debug debugger = new Debug();
@@ -18,7 +22,7 @@ public class TestDriver {
 
 		//Create test data for sorts class
 		ArrayList<Integer> testList = new ArrayList<Integer>();
-		testList = DummyData.runArrayList(0, 1000, 0, 100000, testList);
+		testList = DummyData.runArrayList(0, 1, 0, 100, testList);
 
 		System.out.println("Un-Sorted List");
 		System.out.println(testList);

### 2015-01-30 Programming Project A/Insertion Sort Analysis in Java.tex
added latex files

--- 2015-01-28 Programming Project A/Provided Files/LaTex Template/sample.tex
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% DOCUMENT
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% BEGIN DOCUMENT
 \begin{document}
 
-%\pagestyle{empty} %No headings for the first pages.
+%% Title Page
+\title{Insertion Sort Analysis in Java}
+\author{Preston Stosur-Bassett}
+\date{February 2, 2015}
+\maketitle
 
-%% Title Page %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% ==> Write your text here or include other files.
+\pagestyle{fancy}
+\fancyhead[R]{Insertion Sort Analysis in Java, page \thepage}
+\fancyhead[L]{Preston Stosur-Bassett}
 
-%% The simple version:
-\title{Meaningful Title for this Paper}
-\author{John Doe}
-\date{} %%If commented, the current date is used.
-\maketitle
-%\fancyhead()
-
-\pagestyle{fancy} %Now display headings: headings / fancy / ...
-\fancyhead[R]{Insertion Sort, page \thepage}
-\fancyhead[L]{John Doe}
-
+%% BEGIN ABSTRACT
 \begin{abstract}
 
-The abstract is written \emph{after} the report and gives a high--level overview.
-An abstract is optional for these reports, but it is good practice.
-Look up some published papers and read the abstracts to get an idea of how to write one.
+%% END ABSTRACT
+\end{abstract}
 
-\end{abstract}
+%% BEGIN MOTIVATION
 \section{Motivation}
-There has to be some good reasons to do this work and publish it.
-Why?
-Never refer to any report for this class as a class project, lab project, or assignment.
-It is a report, a work, or an article.
-Try to be professional.
 
+%% END MOTIVATION
+
+%% BEGIN BACKGROUND
 \section{Background}
-What exactly is a good description of the problem?
-What work has been done on it?
-In the past, others have reported algorithms to perform an Insertion Sort~\cite{cormen2001introduction}.
-Remember:  If you copy it, it's plagiarism until you cite it.
 
+%% END BACKGROUND
+
+%% BEGIN PROCEDURE
 \section{Procedure}
-How did you do this wonderful thing anyway?
-This is where you put your pseudocode (see Algorithm~\ref{prot:auth}), pre-conditions, post conditions, and Invariants.
-Describe the idea in a general way.
-How you implement your procedure goes in the Experimental Analysis.
-\begin{algorithm}
-\caption {\textsc{Build-Max-Heap}(A)}
-\label{prot:auth}
 
-\begin{algorithmic}[1]
-\Procedure{Build-Max-Heap}{A}%\Comment{The g.c.d. of a and b}
-\State{assert(A is not empty)}
-\State {$heapsize = A.length$}
- \For {$i = \left\lfloor heapsize / 2 \right\rfloor  \text{\textbf{ downto }} 2$} 
-     \State  Max-Heapify(A,$ i$)
-    \State{assert(Heap-Test(A,$i$)}
- \EndFor
-\State{assert(Heap-Test(A,$ 1$)}
-\EndProcedure
-\end{algorithmic}
+%% END PROCEDURE
 
-\end{algorithm}
-
-\section{Testing}
-\subsection{Testing Plan and Results}
-This can be done as a table.  Something like:\\\\
-\small{\begin{tabular}{|l|l|l|l|}
-\hline Test input & Expected Results & Actual Results & Corrective Actions \\ 
-\hline  Empty array & nothing returned  & nothing was returned & no action taken  \\ 
-\hline All elements identical & same array  & Program faulted & off--by--one error corrected  \\ 
-\hline 
-\end{tabular}}
-\subsection{Problems Encountered}
-Normally, you would not see these in a paper, but our projects are a special case.
-You might see this in a paper that developed a simulator for running virtual experiments.
-If you had no problems, state ``No significant problems were encountered.''  Syntax errors and typo's are not problems.
-You can give a brief description of how you fixed the problem, for example ``Updating to the latest version of java corrected the issues with program faults.''  Use your best judgement on what is a problem and what is not.
-If you got help with your problem, acknowledge it here or cite the help as a conversation\footnote{Look up how to cite conversations.
-Also avoid footnotes.}.
-
-\section{Experimental Analysis}
-IF you wish, you can include a paragraph about the specifications of the platform you used to run your experiments.
-Be very brief:  ``These timings were carried out on a 427 cubic inch supercharged Dell MT708 with dual overhead BlueRay drives.
-The power of the CUP was wasted by running Windows 8.1 Pro with all the latest updates as of January 12, 1995.
-Oracle Java 1.0 was used as the programming language/compiler.''
-
-Here you need to graph your results and explain them.
-This is the main part of your paper.
-Graphs should be inserted as figures and referenced by ``...agreed well with the expected results, see Figure~\ref{fig:figureforprinting}.
-Be careful of color-coding parts of graphs as most printing is still black and white.
-Notice how I used both color and dashes to differentate the two lines on the graph.
-You should be able to get away with one graph with two lines.
-Mark $n_0$ on the graph if you can.
-
-\begin{figure}[!]
-\centering
-\includegraphics[width=0.7\linewidth]{./figure.pdf}
-\caption{A Sample Graph}
-\label{fig:figureforprinting}
-\end{figure}
-Current style is to place tables \emph{before} they are referenced with the caption above the table.
-Figures are place \emph{after} they are referenced and the caption is below the figure.
-If you have time, read the IEEE Style Sheet in the same folder with this sample.
-
-Explain clearly what your results mean.
-If you have a result that makes no sense, clearly state why it makes no sense.
-If you can, give a plausible explanation as to what may have cause the strange result.
-Do not try to hide a bad result.
-
-\section{Conclusions}
-Here you answer any questions raised in the previous sections.   You should end with a wrap--up and firm statement of what a great accomplishment this is.
-
-% references section
-%\baselineskip=.55\baselineskip
-
-\bibliographystyle{IEEEtran}
-\bibliography{references}
-
-\newpage
-\section*{Appendix}
-An Appendix starts on a new page after everything else.  This is were your code will be.\\\\
-
-\lstinputlisting{fragment.java}
+%% END DOCUMENT
 \end{document}

### 2015-01-30 Programming Project A/Code/TestDriver.java

@@ -21,8 +21,9 @@ public class TestDriver {
 		debugger.print("Debug is on.");
 
 		//Create test data for sorts class
+		//Note that this Array is for testing purposes only, the Algorithm can handle all Comparable Generic Types
 		ArrayList<Integer> testList = new ArrayList<Integer>();
-		testList = DummyData.runArrayList(0, 1, 0, 100, testList);
+		testList = DummyData.runArrayList(0, 1000000, 0, 1000000, testList);
 
 		System.out.println("Un-Sorted List");
 		System.out.println(testList);
@@ -32,6 +33,7 @@ public class TestDriver {
 		Stopwatch watchStopper = new Stopwatch();
 		testList = sorter.insertion(testList);
 
+		//Print out the results.
 		System.out.println("Sorted List");
 		System.out.println(testList);
 		System.out.println("Time to complete: "+watchStopper.elapsedTime());

### 2015-01-30 Programming Project A/Code/DummyData.java
tested more values

@@ -11,13 +11,13 @@ import java.util.Random;
 public class DummyData {
 
 	/*
-	*	@Description runArrayList will take an ArrayList of Integer Objects and add a given amount of values to it
+	*	@Description runArrayList<Integer> will take an ArrayList of Integer Objects and add a given amount of values to it
 	*	@param int start the starting value to add to the array list
 	*	@param int end the ending value to denote when to stop adding to the array list
 	*	@param int min the minimum value of the randomly generated data.
 	*	@param int max the maximum value of the randomly generated data.
 	*	@param ArrayList<Integer> list the list to add value to and return
-	*	@return ArrayList<Integer> the list after it has been updated with the given data
+	*	@return ArrayList<Integer> the list after it has been updated with the randomly generated data
 	*/
 	public static ArrayList<Integer> runArrayList(int start, int end, int min, int max, ArrayList<Integer> list) {
 		Random random = new Random();
@@ -25,7 +25,28 @@ public class DummyData {
 			Integer intToAdd = new Integer(random.nextInt((max - min + 1) + min));
 			list.add(intToAdd);
 
-			//Count up on the iterators
+			//Count up on the iterator
+			start++;
+		}
+
+		return list;
+	}
+
+	/*
+	*	@Description: runArrayList<String> will take an ArrayList of String Objects and add a given amount of String numerical values to it
+	*	@param int start the starting value to add to the array list
+	*	@param int end the ending value to denote when to stop adding to the array list
+	*	@param ArrayList<String> list the list to add String values to and return
+	*	@return ArrayList<String> the list after it has been updated with the randomly generated numerical String values
+	*/
+	public static ArrayList<String> runArrayList(int start, int end, ArrayList<String> list) {
+		Random random = new Random();
+		while(start < end) {
+			Integer intToString = new Integer(random.nextInt((1000000 - 1) + 1));
+			String intString = String.valueOf(intToString);
+			list.add(intString);
+
+			//Count up on the iterator
 			start++;
 		}
 

### 2015-01-30 Programming Project A/Code/TestDriver.java

@@ -22,14 +22,17 @@ public class TestDriver {
 
 		//Create test data for sorts class
 		//Note that this Array is for testing purposes only, the Algorithm can handle all Comparable Generic Types
-		ArrayList<Integer> testList = new ArrayList<Integer>();
-		testList = DummyData.runArrayList(0, 1000000, 0, 1000000, testList);
+		//TEMP: ArrayList<Integer> testList = new ArrayList<Integer>();
+		//TEMP: testList = DummyData.runArrayList(0, 0, 0, 1000000, testList);
+		ArrayList<String> testList = new ArrayList<String>();
+		testList = DummyData.runArrayList(0, 1000, testList);
 
 		System.out.println("Un-Sorted List");
 		System.out.println(testList);
 
 		//Test sort List
-		Sort sorter = new Sort<Integer>();
+		//Sort sorter = new Sort<Integer>();
+		Sort sorter = new Sort<String>();
 		Stopwatch watchStopper = new Stopwatch();
 		testList = sorter.insertion(testList);
 

### 2015-01-30 Programming Project A/Code/Sort.java
added minor loop invariant documentation

@@ -15,19 +15,27 @@ public class Sort<T extends Comparable<T>> {
 	*	@param ArrayList<T> unsorted is a non-empty unsorted array list of T, where T is a comparable type
 	*	@return sorted is a permutation of <code>unsorted</code> where all the elements are sorted in increasing order
 	*/
+	// INVARIANT (Outer-Loop): sorted[0 ... i - 1] will contain all the same data as unsorted[0 ... i - 1] but sorted in stricly non-decreasing order.
+	// INVARIANT (Inner-Loop):
 	public ArrayList<T> insertion(ArrayList<T> unsorted) {
 		ArrayList<T> sorted = unsorted;
 		if(sorted.size() > 1) {
+			// INITIALIZATION (Outer-Loop): The invariant holds because i = 1, and there is one element in the subarray of sorted[0 ... i - 1] and unsorted[0 ... i - 1], which is already in strictly non-decreasing order
 			for(int i = 1; i < sorted.size(); i++) {
+				// MAINTENANCE (Outer-Loop):
 				T value = sorted.get(i);
 				int j = i - 1;
+				// INITIALIZATION (Inner-Loop):
 				while(j >= 0 && (value.compareTo(sorted.get(j)) < 0)) {
+					// MAINTENANCE: (Inner-Loop):
 					sorted.set(j+1, sorted.get(j));
 					j--;
 				}
+				// TERMINATION (Inner-Loop):
 				sorted.set(j+1, value);
 			}
 		}
+		// TERMINATION (Outer-Loop):
 		return sorted;
 	}
 }

### 2015-01-30 Programming Project A/Code/Debug.java
updated assertions

@@ -48,7 +48,7 @@ public class Debug<T> {
 	*/
 	public void assertIt(T expected, T actual) {
 		if(debugOn == true) {
-			assert actual == expected;
+			assert actual.equals(expected);
 		}
 	}
 }

### 2015-01-30 Programming Project A/Code/Sort.java

@@ -6,6 +6,7 @@
 */
 
 import java.util.ArrayList;
+import java.util.List;
 
 public class Sort<T extends Comparable<T>> {
 	/*
@@ -18,10 +19,16 @@ public class Sort<T extends Comparable<T>> {
 	// INVARIANT (Outer-Loop): sorted[0 ... i - 1] will contain all the same data as unsorted[0 ... i - 1] but sorted in stricly non-decreasing order.
 	// INVARIANT (Inner-Loop):
 	public ArrayList<T> insertion(ArrayList<T> unsorted) {
+		Debug debugger = new Debug<List<T>>();
+		debugger.turnOn();
 		ArrayList<T> sorted = unsorted;
 		if(sorted.size() > 1) {
-			// INITIALIZATION (Outer-Loop): The invariant holds because i = 1, and there is one element in the subarray of sorted[0 ... i - 1] and unsorted[0 ... i - 1], which is already in strictly non-decreasing order
 			for(int i = 1; i < sorted.size(); i++) {
+				// INITIALIZATION (Outer-Loop): The invariant holds because i = 1, and there is one element in the subarray of sorted[0 ... i - 1] and unsorted[0 ... i - 1], which is already in strictly non-decreasing order
+				List<T> subSorted = sorted.subList(0, i - 1);
+				List<T> subUnsorted = unsorted.subList(0, i - 1);
+				debugger.assertIt(subUnsorted, subSorted);
+
 				// MAINTENANCE (Outer-Loop):
 				T value = sorted.get(i);
 				int j = i - 1;

### 2015-01-30 Programming Project A/Code/TestDriver.java

@@ -22,17 +22,17 @@ public class TestDriver {
 
 		//Create test data for sorts class
 		//Note that this Array is for testing purposes only, the Algorithm can handle all Comparable Generic Types
-		//TEMP: ArrayList<Integer> testList = new ArrayList<Integer>();
-		//TEMP: testList = DummyData.runArrayList(0, 0, 0, 1000000, testList);
-		ArrayList<String> testList = new ArrayList<String>();
-		testList = DummyData.runArrayList(0, 1000, testList);
+		ArrayList<Integer> testList = new ArrayList<Integer>();
+		testList = DummyData.runArrayList(0, 5, 0, 10, testList);
+		//ArrayList<String> testList = new ArrayList<String>();
+		//testList = DummyData.runArrayList(0, 1000, testList);
 
 		System.out.println("Un-Sorted List");
 		System.out.println(testList);
 
 		//Test sort List
-		//Sort sorter = new Sort<Integer>();
-		Sort sorter = new Sort<String>();
+		Sort sorter = new Sort<Integer>();
+		//Sort sorter = new Sort<String>();
 		Stopwatch watchStopper = new Stopwatch();
 		testList = sorter.insertion(testList);
 

### 2015-01-30 Programming Project A/Code/Sort.java
added more assertion statements

@@ -23,13 +23,19 @@ public class Sort<T extends Comparable<T>> {
 		debugger.turnOn();
 		ArrayList<T> sorted = unsorted;
 		if(sorted.size() > 1) {
-			for(int i = 1; i < sorted.size(); i++) {
-				// INITIALIZATION (Outer-Loop): The invariant holds because i = 1, and there is one element in the subarray of sorted[0 ... i - 1] and unsorted[0 ... i - 1], which is already in strictly non-decreasing order
-				List<T> subSorted = sorted.subList(0, i - 1);
-				List<T> subUnsorted = unsorted.subList(0, i - 1);
-				debugger.assertIt(subUnsorted, subSorted);
+			// INITIALIZATION (Outer-Loop): The invariant holds because i = 1, and there is one element in the subarray of sorted[0 ... i - 1] and unsorted[0 ... i - 1], which is already in strictly non-decreasing order
+			List<T> subSortedI = sorted.subList(0, 0);
+			List<T> subUnsortedI = unsorted.subList(0, 0);
+			debugger.assertIt(subUnsortedI, subSortedI);
 
-				// MAINTENANCE (Outer-Loop):
+			int i = 1;
+			while(i < sorted.size()) {
+				// MAINTENANCE (Outer-Loop): At the beginning of each iteration of the loop, the loop invariant is maintained because the subarray of sorted[0 ... i - 1] contains all the same elements as unsorted[0 ... i - 1] and is in strictly non-decreasing order
+				// TODO: ensure list is is stricly non decreasing order
+				List<T> subSortedM = sorted.subList(0, i - 1);
+				List<T> subUnsortedM = unsorted.subList(0, i - 1);
+				debugger.assertIt(subUnsortedM, subSortedM);
+
 				T value = sorted.get(i);
 				int j = i - 1;
 				// INITIALIZATION (Inner-Loop):
@@ -40,9 +46,18 @@ public class Sort<T extends Comparable<T>> {
 				}
 				// TERMINATION (Inner-Loop):
 				sorted.set(j+1, value);
+
+				//Count up on the iterator
+				i++;
 			}
+			/* TERMINATION (Outer-Loop): When the loop terminates, i is equal to sorted.size() meaning the entire array has been traversed and that the guard has been negated.
+				The negation of the guard implies that sorted[0 ... i - 1] contains all the elements of unsorted[0 ... i - 1] in strictly non-decreasing order. */
[…]

### 2015-01-31 Programming Project A/Code/Debug.java
running into some assertOrder errors

@@ -5,6 +5,8 @@
 *	@Description This class will help debugging by being able to turn on and turn off debug messages easily
 */
 
+import java.util.List;
+
 public class Debug<T> {
 	boolean debugOn; //Variable to keep track of whether or not debug is on
 
@@ -46,9 +48,23 @@ public class Debug<T> {
 	*	@param T expected the expected value to assert against the actual value
 	*	@param T actualt he actual value to assert against the expected value
 	*/
-	public void assertIt(T expected, T actual) {
+	public void assertEquals(T expected, T actual) {
 		if(debugOn == true) {
 			assert actual.equals(expected);
+		}
+	}
+
+	/*
+	*
+	*/
+	public void assertOrder(List<Integer> actual) {
+		if(debugOn == true) {
+			int i = actual.size();
+			while(i > 1) {
+				assert (actual.get(i).compareTo(actual.get(i - 1) > 1));
+
+				i--;
+			}
 		}
 	}
 }

### 2015-01-31 Programming Project A/Code/DummyData.java

@@ -21,6 +21,7 @@ public class DummyData {
 	*/
 	public static ArrayList<Integer> runArrayList(int start, int end, int min, int max, ArrayList<Integer> list) {
 		Random random = new Random();
+		// TODO: Write loop invariant
 		while(start < end) {
 			Integer intToAdd = new Integer(random.nextInt((max - min + 1) + min));
 			list.add(intToAdd);
@@ -41,6 +42,7 @@ public class DummyData {
 	*/
 	public static ArrayList<String> runArrayList(int start, int end, ArrayList<String> list) {
 		Random random = new Random();
+		// TODO: Write loop invariant
 		while(start < end) {
 			Integer intToString = new Integer(random.nextInt((1000000 - 1) + 1));
 			String intString = String.valueOf(intToString);

### 2015-01-31 Programming Project A/Code/Sort.java

@@ -16,47 +16,55 @@ public class Sort<T extends Comparable<T>> {
 	*	@param ArrayList<T> unsorted is a non-empty unsorted array list of T, where T is a comparable type
 	*	@return sorted is a permutation of <code>unsorted</code> where all the elements are sorted in increasing order
 	*/
-	// INVARIANT (Outer-Loop): sorted[0 ... i - 1] will contain all the same data as unsorted[0 ... i - 1] but sorted in stricly non-decreasing order.
-	// INVARIANT (Inner-Loop):
+	// INVARIANT (Outer-Loop): The pre condition implies that sorted[0 ... i - 1] will contain all the same data as unsorted[0 ... i - 1].
+	// INVARIANT (Inner-Loop): sorted[0 ... j] is sorted in stricly non-decreasing order.
 	public ArrayList<T> insertion(ArrayList<T> unsorted) {
 		Debug debugger = new Debug<List<T>>();
 		debugger.turnOn();
 		ArrayList<T> sorted = unsorted;
 		if(sorted.size() > 1) {
-			// INITIALIZATION (Outer-Loop): The invariant holds because i = 1, and there is one element in the subarray of sorted[0 ... i - 1] and unsorted[0 ... i - 1], which is already in strictly non-decreasing order
-			List<T> subSortedI = sorted.subList(0, 0);
-			List<T> subUnsortedI = unsorted.subList(0, 0);
-			debugger.assertIt(subUnsortedI, subSortedI);
-
 			int i = 1;
+			/* INITIALIZATION (Outer-Loop): The invariant holds because i = 1, and there is one element in the subarray of sorted[0 ... i - 1] and unsorted[0 ... i - 1], */
+			List<T> subSortedOI = sorted.subList(0, i - 1);
+			List<T> subUnsortedOI = unsorted.subList(0, i - 1);
+			debugger.assertEquals(subUnsortedOI, subSortedOI);
+
 			while(i < sorted.size()) {
-				// MAINTENANCE (Outer-Loop): At the beginning of each iteration of the loop, the loop invariant is maintained because the subarray of sorted[0 ... i - 1] contains all the same elements as unsorted[0 ... i - 1] and is in strictly non-decreasing order
-				// TODO: ensure list is is stricly non decreasing order
-				List<T> subSortedM = sorted.subList(0, i - 1);
-				List<T> subUnsortedM = unsorted.subList(0, i - 1);
-				debugger.assertIt(subUnsortedM, subSortedM);
[…]

### 2015-01-31 Programming Project A/Code/Debug.java
now im getting different errors

@@ -61,7 +61,7 @@ public class Debug<T> {
 		if(debugOn == true) {
 			int i = actual.size();
 			while(i > 1) {
-				assert (actual.get(i).compareTo(actual.get(i - 1) > 1));
+				assert actual.get(i).compareTo(actual.get(i - 1)) > 1;
 
 				i--;
 			}

### 2015-01-31 Programming Project A/Code/Debug.java
fixed errors, but now my assertion is throwing an exception

@@ -61,7 +61,7 @@ public class Debug<T> {
 		if(debugOn == true) {
 			int i = actual.size();
 			while(i > 1) {
-				assert actual.get(i).compareTo(actual.get(i - 1)) > 1;
+				assert actual.get(i - 1).compareTo(actual.get(i - 2)) > 1;
 
 				i--;
 			}

### 2015-02-01 Programming Project A/Code/Debug.java
fixed assertion errors

@@ -61,7 +61,7 @@ public class Debug<T> {
 		if(debugOn == true) {
 			int i = actual.size();
 			while(i > 1) {
-				assert actual.get(i - 1).compareTo(actual.get(i - 2)) > 1;
+				assert actual.get(i - 1).compareTo(actual.get(i - 2)) > 0;
 
 				i--;
 			}

### 2015-02-01 Programming Project A/Code/Debug.java
updated documentation

@@ -55,13 +55,16 @@ public class Debug<T> {
 	}
 
 	/*
-	*
+	* @Pre-Condition: <code>List<Integer> actual</code> is a iterable list of Integer objects
+	*	@Post-Conditions: If the List of Integer objects is in stricly non-decreasing order, the program moves on normally, if not, the program halts with an <code>AssertionError</code>
+	* @Description: runs an assertion statement against a list of Integer objects to ensure that for <code>k = actual.size(); A[k - 2] <= A[k - 1];</code>
+	*	@param List<Integer> actual the list to assert is in stricly non-decreasing order
 	*/
 	public void assertOrder(List<Integer> actual) {
 		if(debugOn == true) {
 			int i = actual.size();
 			while(i > 1) {
-				assert actual.get(i - 1).compareTo(actual.get(i - 2)) > 0;
+				assert actual.get(i - 1).compareTo(actual.get(i - 2)) >= 0;
 
 				i--;
 			}

### 2015-02-01 Programming Project A/Insertion Sort Analysis in Java.tex
wrote half of paper, updated code to support more testing

@@
 \begin{abstract}
-
+%% TODO: Write abstract
 %% END ABSTRACT
@@
 \section{Motivation}
-
+In order to show how an algorithm might run on a given set of hardware, and how the algorithm will perform when given large amounts of data, algorithms are analysed.
+More specifically, insertion sort is analysed to determined what the run time might be for sorting large amounts of data on any modern computer.
 %% END MOTIVATION
@@
 \section{Background}
-
+%%% TODO: !! THIS SECTION SEEMS INCOMPLETE, PLEASE FINISH !! 
+A sorting algorithm is used to sort data with a natural order.
+One such sorting algorithm is insertion sort, which sorts by iterating through a list of data, taking the current position and repositioning it into a more appropriate place in the list.
+How will this procedure perform when handling high volumes of data?
+How will it perform when executed on different machines?
+Because insertion sort is a more basic sorting algorithm, numerous papers and articles have been written answering these two questions.
 %% END BACKGROUND
@@
 \section{Procedure}
+An insertion sort can be implemented in a multitude of languages using the pseudocode provided below.
+\newline
+\textbf{Insertion Sort Pre-Condition}: A is a non-empty array of data with a natural order.
+\newline
+\textbf{Insertion Sort Post-Condition}: A' is a permutation of A (containing all the same elements) in strictly non-decreasing order.
+\begin{algorithm}
+\caption {\textsc{Insertion-Sort}(A)}
+\begin{algorithmic}[1]
+\Procedure{Insertion-Sort}{A}
+\State{$i = 2$}
+\While{$i$ upto $A.length$}
+	\State{$key = A[i]$}
+	\State{$j = i - 1$}
+	\While{$j$ downto $1$ and $key < A[j]$}
+		\State{$A[j + 1] = A[j]$}
+		\State{$j = j - 1$}
+	\EndWhile
+	\State{$A[j+1] = key$}
+	\State{$i = i + 1$}
+\EndWhile
+\EndProcedure
+\end{algorithmic}
+\end{algorithm}
+\newline
+\textbf{Outer-Loop Invariant}: The subarray A'[1 ... i - 1] contains all the same elements as the subarray A[1 .. i - 1].
+\newline 
+\textbf{Outer-Loop Initialization}: The outer-loop invariant holds because A'[1 ... i - 1] and A[1 ... i - 1] both contain the same one element.
+\newline
+\textbf{Outer-Loop Maintenance}: The outer-loop invariant holds because A'[1 ... i - 1] and A[1 ... i - 1] both contain the same elements, although they maybe in different orders.
+\newline
+\textbf{Outer-Loop Termination}: When the outer-loop terminates, i = A.length, which implies that the entire array has been traversed and the guard has been negated.
+The negation of the guard implies that A'[1 ...
+i - 1] contains all the elements in A[1 ...
+i - 1].
+\newline
+\newline
+\textbf{Inner-Loop Invariant}: A'[1 ... j] is sorted in strictly non-decreasing order.
+\newline
+\textbf{Inner-Loop Initialization}: Before the first iteration of the loop, j = 1, meaning the subarray A'[1 ... j] contains exactly one element, which is already sorted.
+\newline
+\textbf{Inner-Loop Maintenance}: At the beginning of each iteration of the loop the inner-loop invariants holds because j counts down from i and A'[j+1] is swapped with A'[j] only if A'[j+1] is less than A[j].
+\newline
+\textbf{Inner-Loop Termination}: The negation of the implies that j = A.length and A'[1 ...
+j] has been entirely traversed and sorted in strictly non-decreasing order which maintains the inner-loop invariant.
+\newline
+\newline
+\textbf{Conclusion}: The termination of both the inner and outer loops implies that the entire array has been traversed, A' is a permutation of A containing all the same elements in strictly non-decreasing order.
+This satisfies the post condition.
+%% END PROCEDURE
 
-%% END PROCEDURE
+%% BEGIN TESTING
+\section{Testing}
+\subsection{Testing Plan and Results}
+All arrays used in testing are Java ArrayList<Integer> unless otherwise specified.
+All times are recorded in milliseconds using a stopwatch class borrowed from [NEED CITATION].
+%% TODO: Add Citation %%
+It is important to note that the stopwatch class used takes the elapsed real-time between the start of the insertion sort algorithm and the end of the insertion sort algorithm as opposed to taking the elapsed processor-time because these tests were run on a multi-core computer.
+In the table below A denotes Array. Times in the table below are given as averages out of 10 trials.
+\begin{center}
+\begin{tabular}{|l|l|l|l|}
+\hline Tested Input & Expected Results & Actual Results & Time \\
+\hline Empty A & Empty A & Empty A & 0.0003 \\
+\hline A of 1000 Strings & Sorted A 1000 Strings & Sorted A 1000 Strings & 0.021 \\
+\hline A 1 Element & Original A & Original A & 0.0003 \\
+\hline A 10 Elements & Sorted A 10 Elements & Sorted A 10 Elements & 0.0005 \\
+\hline A 100 Elements & Sorted A 100 Elements & Sorted A 100 Elements & 0.0021 \\
+\hline A 1000 Elements & Sorted A 1000 Elements & Sorted A 1000 Elements & 0.019 \\
+\hline A 10000 Elements & Sorted A 10000 Elements & Sorted A 10000 Elements & 0.129 \\
+\hline A 100000 Elements & Sorted A 100000 Elements & Sorted A 100000 Elements & 6.4923 \\
+\hline A 1000000 Elements & Sorted A 1000000 Elements & Sorted A 1000000 Elements & 2135.5007 \\
+\hline A 10000000 Elements & Sorted A 10000000 Elements & OS Crash & N/A \\
+\hline A 1000 Identical Elements & Original Array & Original Array & 0.0052 \\
+\hline
+\end{tabular}
+\end{center}
+\subsection{Problems Encountered}
+
+%% END TESTING
+
+%% BEGIN EXPERIMENTAL ANALYSIS
+\section{Experimental Analysis}
+
+%% END EXPERIMENTAL ANALYSIS
+
+%% BEGIN CONCLUSION
+\section{Conclusions}
+
+%% END CONCLUSION
+
+\newpage
+
+%% BEGIN REFERENCES
+\section*{References}
+
+%% END REFERENCES
+
+\newpage
+
+%% BEGIN APPENDIX
+\section*{Appendix}
+
+%% END APPENDIX
 

### 2015-02-01 Programming Project A/Code/Debug.java

@@ -63,6 +63,7 @@ public class Debug<T> {
 	public void assertOrder(List<Integer> actual) {
 		if(debugOn == true) {
 			int i = actual.size();
+			// TODO: Write invariant for this loop
 			while(i > 1) {
 				assert actual.get(i - 1).compareTo(actual.get(i - 2)) >= 0;
 

### 2015-02-01 Programming Project A/Code/DummyData.java

@@ -21,14 +21,17 @@ public class DummyData {
 	*/
 	public static ArrayList<Integer> runArrayList(int start, int end, int min, int max, ArrayList<Integer> list) {
 		Random random = new Random();
-		// TODO: Write loop invariant
+		// INVARIANT: TODO: Come up with invariant
+		// INITIALIZATION: TODO: Assert initialization
 		while(start < end) {
+			// MAINTANANCE: TODO: Assert Maintanance
 			Integer intToAdd = new Integer(random.nextInt((max - min + 1) + min));
 			list.add(intToAdd);
 
 			//Count up on the iterator
 			start++;
 		}
+		//TERMINATION: TODO: Assert Termination
 
 		return list;
 	}
@@ -47,6 +50,28 @@ public class DummyData {
 			Integer intToString = new Integer(random.nextInt((1000000 - 1) + 1));
 			String intString = String.valueOf(intToString);
 			list.add(intString);
+
+			//Count up on the iterator
+			start++;
+		}
+
+		return list;
+	}
+
+	/*
+	*	@Description: identicalElement will take an element and add it to the ArrayList<Integer> for a given amount of times
+	*	@param int start the starting value to be used as an iterator for the loop
+	*	@param int end the ending value to denote when to stop adding elements to the array
+	*	@param int element the element to add over and over again to the array
+	*	@param ArrayList<Integer> list the list to add elements to
+	*	@return ArrayList<Integer> the list after it has been updated with the given data
+	*/
+	public static ArrayList<Integer> identicalElement(int start, int end, int element, ArrayList<Integer> list) {
+		// TODO: Write loop invariant
+		//The element to add over and over again
+		Integer iden = new Integer(element);
+		while(start < end) {
+			list.add(iden);
 
 			//Count up on the iterator
 			start++;

### 2015-02-01 Programming Project A/Code/Sort.java

@@ -20,7 +20,7 @@ public class Sort<T extends Comparable<T>> {
 	// INVARIANT (Inner-Loop): sorted[0 ... j] is sorted in stricly non-decreasing order.
 	public ArrayList<T> insertion(ArrayList<T> unsorted) {
 		Debug debugger = new Debug<List<T>>();
-		debugger.turnOn();
+		//debugger.turnOn();
 		ArrayList<T> sorted = unsorted;
 		if(sorted.size() > 1) {
 			int i = 1;

### 2015-02-01 Programming Project A/Code/TestDriver.java

@@ -23,7 +23,8 @@ public class TestDriver {
 		//Create test data for sorts class
 		//Note that this Array is for testing purposes only, the Algorithm can handle all Comparable Generic Types
 		ArrayList<Integer> testList = new ArrayList<Integer>();
-		testList = DummyData.runArrayList(0, 5, 0, 10, testList);
+		testList = DummyData.runArrayList(0, 10, 0, 10, testList);
+		//testList = DummyData.identicalElement(0, 1000, 10, testList);
 		//ArrayList<String> testList = new ArrayList<String>();
 		//testList = DummyData.runArrayList(0, 1000, testList);
 

### 2015-02-01 Programming Project A/Insertion Sort Analysis in Java.tex
wrote more latex

@@
 \usepackage{listings}
+\usepackage{capt-of}
 
@@
 \section{Background}
-%%% TODO: !! THIS SECTION SEEMS INCOMPLETE, PLEASE FINISH !! 
 A sorting algorithm is used to sort data with a natural order.
@@
 Because insertion sort is a more basic sorting algorithm, numerous papers and articles have been written answering these two questions.
+The insertion sort demonstrated below (see Algorithm 1) was implemented in Java and executed on an HP SprectreXT TouchSmart with 4 Intel Core i7 at 1.GHz running Ubuntu Gnome 14.10 64-bit.
 %% END BACKGROUND
@@
 In the table below A denotes Array. Times in the table below are given as averages out of 10 trials.
+\newline
+\captionof{table}{Test Results}
 \begin{center}

### 2015-02-02 Programming Project A/Insertion Sort Analysis in Java.tex
added graph to latex
added more content to paper, reformeted a few things

@@
 Because insertion sort is a more basic sorting algorithm, numerous papers and articles have been written answering these two questions.
-The insertion sort demonstrated below (see Algorithm 1) was implemented in Java and executed on an HP SprectreXT TouchSmart with 4 Intel Core i7 at 1.GHz running Ubuntu Gnome 14.10 64-bit.
 %% END BACKGROUND
@@
 \caption {\textsc{Insertion-Sort}(A)}
+\label{algo:insertionsort}
 \begin{algorithmic}[1]
@@
 \end{center}
+
 \subsection{Problems Encountered}
-
+One major issue encountered during the development of this insertion sort was that after completing the sort, A' was completely sorted properly except for the first element in the array.
+No matter what value the first element of A had, it did not change position in A'.
+For example, if A[5, 6, 3, 4, 7] was passed to the insertion sort algorithm, the returned array would look like A'[5, 3, 4, 6, 7].
+Changing the guard for the inner for loop (see Algorithm 1 line 6) from \textit{key < A[j] and j downto 1} to \textit{j downto 1 and key < A[j]} corrected this issue.
 %% END TESTING
@@
 \section{Experimental Analysis}
+ The insertion sort demonstrated in Algorithm 1 was implemented in Java and executed on an HP SprectreXT TouchSmart with 4 Core Intel i7 processor clocked at 1.9GHz running Ubuntu Gnome 14.10 64-bit.
+ \begin{figure}[!]
+\begin{center}
+\includegraphics[width=0.7\linewidth]{insertion_sort_time_analysis.pdf}
+\end{center}
+\captionof{figure}{Insertion Sort Time Analysis}
+\label{fig:insertionsorttimeanalysis}
+\end{figure}
 

### 2015-02-02 Programming Project A/Insertion Sort Analysis in Java.tex
updated appendix

@@
 \section*{Appendix}
-
+\lstinputlisting[caption=TestDriver,
+label={TestDriver.java},
+breaklines=true,
+]{Code/TestDriver.java}
+\lstinputlisting[caption=Debug,
+label={Debug.java},
+breaklines=true,
+]{Code/Debug.java}
+\lstinputlisting[caption=DummyData,
+label={DummyData.java},
+breaklines=true,
+]{Code/DummyData.java}
+\lstinputlisting[caption=Stopwatch,
+label={Stopwatch.java}
+breaklines=true,
+]{Code/Stopwatch.java}
+\lstinputlisting[caption=Sort,
+label={Sort.java},
+breaklines=true,
+]{Code/Sort.java}
 %% END APPENDIX

### 2015-02-03 Programming Project A/Insertion Sort Analysis in Java.tex
added content to paper

@@
 \Procedure{Insertion-Sort}{A}
+\If{$A.length <= 1$}
+\State{\Return{$A$}}
+\EndIf
 \State{$i = 2$}
@@
 \begin{center}
-\includegraphics[width=0.7\linewidth]{insertion_sort_time_analysis.pdf}
+\includegraphics[scale=.75]{insertion_sort_time_analysis.pdf}
 \end{center}
@@
 
+The expected growth of insertion sort as the number of elements (n) grows large can be represented as $ \theta(n^2) $ where $ \theta() $ represents the asymptotically tightly bound running time.
+At $n_0$ the algorithm took 0.0003 milliseconds to complete.
+At $n_1$ the algorithm also took 0.0003 milliseconds to complete.
+For both these values of $n$, the algorithm runs at a constant time because no for-loops are executed.
+As $n$ grows larger the time to complete the experimental data correlates quite accurately with the expected growth.
+For $n_{1*10^6} $ the data matches up perfectly.
+This is to be expected however, because as n grows larger the constants and lower orders of the actual running time of the insertion sort start to effect the the running time less and less because the highest order of $n^2$ is so large.
 %% END EXPERIMENTAL ANALYSIS
@@
 ]{Code/DummyData.java}
+
+The class Stopwatch has not been altered from its original form.
+\newline
 \lstinputlisting[caption=Stopwatch,

### 2015-02-03 Programming Project A/Code/Debug.java
added more invariants to code

@@ -61,14 +61,58 @@ public class Debug<T> {
 	*	@param List<Integer> actual the list to assert is in stricly non-decreasing order
 	*/
 	public void assertOrder(List<Integer> actual) {
+		//TODO: Write invariant
 		if(debugOn == true) {
 			int i = actual.size();
-			// TODO: Write invariant for this loop
 			while(i > 1) {
 				assert actual.get(i - 1).compareTo(actual.get(i - 2)) >= 0;
 
 				i--;
 			}
+		}
+	}
+
+	/*
+	* @Description: asserts that the first arguement is stricly greator than the second arguement
+	*	@param int large an integer primative value to assert is strictly greator than the second arguement
+	*	@param int small an integer primative value to assert the first arguement is strictly greator than.
+	*/
+	public void assertStrictGreat(int large, int small) {
+		if(debugOn == true) {
+			assert large > small;
+		}
+	}
+
+	/*
+	*	@Description: asserts that the first arguement is strictly less than the second arguement
+	*	@param int small an integer primative value to assert is stricly less than the second arguement
+	*	@param int large an integer primative value to assert the first arguement is strictly less than.
+	*/
+	public void assertStrictLess(int small, int large) {
+		if(debugOn == true) {
+			assert small < large;
+		}
+	}
+
+	/*
+	*	@Description: asserts that the first arguement is greator than or equal to the second arguement
+	*	@param int large an integer primative value to assert is greator than or equal to the second arguement
+	*	@param int small an integer primative value to assert the first arguement is greator than or equal to.
+	*/
+	public void assertGreatEqual(int large, int small) {
+		if(debugOn == true) {
+			assert large >= small;
+		}
+	}
+
+	/*
+	*	@Description: asserts that the first arguement is less than or equal to the second arguement
+	*	@param int small an integer primative value to assert is less than or equal to the second arguement
+	*	@param int large an integer primative value to assert the first arguement is less than or equal to.
+	*/
+	public void assertLessEqual(int small, int large) {
+		if(debugOn == true) {
+			assert small <= large;
 		}
 	}
 }

### 2015-02-03 Programming Project A/Code/Debug.java
no invariants for this class

@@ -61,7 +61,6 @@ public class Debug<T> {
 	*	@param List<Integer> actual the list to assert is in stricly non-decreasing order
 	*/
 	public void assertOrder(List<Integer> actual) {
-		//TODO: Write invariant
 		if(debugOn == true) {
 			int i = actual.size();
 			while(i > 1) {

### 2015-02-04 Programming Project A/Insertion Sort Analysis in Java.tex
finished experimental analysis section

@@
 This is to be expected however, because as n grows larger the constants and lower orders of the actual running time of the insertion sort start to effect the the running time less and less because the highest order of $n^2$ is so large.
+It is expected for lower values of n to not correlate well with $ \theta(n^2) $ because the system running the algorithm may be executing superfolous commands such as checking for system.
+For the data graphed above (see Fig.
+1) this was in fact the case.
+To aid this issue, averages of 10 trails were taken and used in Figure 1.
 %% END EXPERIMENTAL ANALYSIS

### 2015-02-04 Programming Project A/Insertion Sort Analysis in Java.tex
proof read, added more content

@@
 A sorting algorithm is used to sort data with a natural order.
-One such sorting algorithm is insertion sort, which sorts by iterating through a list of data, taking the current position and repositioning it into a more appropriate place in the list.
+One such sorting algorithm is insertion sort, which sorts by iterating through a list of data, taking the current position, and repositioning it into a more appropriate place in the list.
 How will this procedure perform when handling high volumes of data?
@@
 \newline
-\textbf{Outer-Loop Maintenance}: The outer-loop invariant holds because A'[1 ... i - 1] and A[1 ... i - 1] both contain the same elements, although they maybe in different orders.
+\textbf{Outer-Loop Maintenance}: The outer-loop invariant holds because A'[1 ... i - 1] and A[1 ... i - 1] both contain the same elements, although they may be in different orders.
 \newline
@@
 \newline
-\textbf{Inner-Loop Maintenance}: At the beginning of each iteration of the loop the inner-loop invariants holds because j counts down from i and A'[j+1] is swapped with A'[j] only if A'[j+1] is less than A[j].
+\textbf{Inner-Loop Maintenance}: At the beginning of each iteration of the loop the inner-loop invariant holds because j counts down from i, and A'[j+1] is swapped with A'[j] only if A'[j+1] is less than A[j].
 \newline
-\textbf{Inner-Loop Termination}: The negation of the implies that j = A.length and A'[1 ...
-j] has been entirely traversed and sorted in strictly non-decreasing order which maintains the inner-loop invariant.
+\textbf{Inner-Loop Termination}: The negation of the guard implies that j = A.length and that A'[1 ...
+j] has been entirely traversed and sorted in strictly non-decreasing order, which maintains the inner-loop invariant.
 \newline
@@
 It is important to note that the stopwatch class used takes the elapsed real-time between the start of the insertion sort algorithm and the end of the insertion sort algorithm as opposed to taking the elapsed processor-time because these tests were run on a multi-core computer.
-In the table below A denotes Array. Times in the table below are given as averages out of 10 trials.
+In the table below, A denotes Array. Times in the table below are given as averages out of 10 trials.
 \newline
@@
 \subsection{Problems Encountered}
-One major issue encountered during the development of this insertion sort was that after completing the sort, A' was completely sorted properly except for the first element in the array.
+One major issue encountered during the development of this insertion sort was that after completing the sort, A' was sorted properly except for the first element in the array.
 No matter what value the first element of A had, it did not change position in A'.
@@
 Changing the guard for the inner for loop (see Algorithm 1 line 6) from \textit{key < A[j] and j downto 1} to \textit{j downto 1 and key < A[j]} corrected this issue.
+
+Another issue encountered during the development of this program was
+
+%% Also mention generics
+
 %% END TESTING
@@
 For $n_{1*10^6} $ the data matches up perfectly.
-This is to be expected however, because as n grows larger the constants and lower orders of the actual running time of the insertion sort start to effect the the running time less and less because the highest order of $n^2$ is so large.
+This is to be expected, because as n grows larger the constants and lower orders of the actual running time of the insertion sort start to affect the running time less and less as the highest order of $n^2$ is so large.
 It is expected for lower values of n to not correlate well with $ \theta(n^2) $ because the system running the algorithm may be executing superfolous commands such as checking for system.
@@
 \section{Conclusions}
+Insertion sort is not the best choice for sorting large amounts of data because as n grows large, it takes insertion sort $ \theta(n^2) $ time to complete on modern day computers.
+Obviously, the faster the system executing insertion sort, the faster the algorithm will complete, however, with large enough values of n, it can take powerful systems quite a long amount of time to complete, as shown in Figure 1.
+
+%% TODO: Add more here
 

### 2015-02-04 Programming Project A/Insertion Sort Analysis in Java.tex
added more issues

@@
 
-Another issue encountered during the development of this program was
+Another issue encountered during the development of this program was executing Java $assert$ statements.
+When assertions were wrapped in parenthesis, they evaluated to a boolean, and the assert would throw an $AssertionError$ with the explanation 'cannot compare to boolean'.
+Removing the parenthesis wrapping the assertion fixed this issue.
 
-%% Also mention generics
+The last issue encountered during the development process of this program was generic programming.
+The program would not compile because not all Objects have a compareTo method.
+Adding the statement $<T$ $extends$ $Comparable<T>>$ fixed this issue.
 
@@
 
+%% TODO: Cite shit
+
 %% END REFERENCES

### 2015-02-04 Programming Project A/Insertion Sort Analysis in Java.tex
added references, but for some reason they are not working.... :-0
got bib working
minor update
added abstract

@@
 %
-% TODO:
+% Note:
 % 1. Edit the used packages and their options (see above).
@@
 \begin{abstract}
-%% TODO: Write abstract
+The present study investigates the insertion sort sorting algorithm and how it performs on data sets of different sizes when implemented in Java.
+Insertion sort is a sorting algorithm that runs in $ \theta(n^2) $.
+For sorting large sets of data, insertion sort does not perform exemplary, however, for simple sorting of data sets under 10,000, insertion sort performs reasonably quickly on modern day computers.
 %% END ABSTRACT
@@
 In order to show how an algorithm might run on a given set of hardware, and how the algorithm will perform when given large amounts of data, algorithms are analysed.
-More specifically, insertion sort is analysed to determined what the run time might be for sorting large amounts of data on any modern computer.
+More specifically, insertion sort is analysed to determine what the run time might be for sorting large amounts of data on any modern computer.
 %% END MOTIVATION
@@
 Obviously, the faster the system executing insertion sort, the faster the algorithm will complete, however, with large enough values of n, it can take powerful systems quite a long amount of time to complete, as shown in Figure 1.
-
-%% TODO: Add more here
-
 %% END CONCLUSION
@@
 %% BEGIN REFERENCES
-\section*{References}
+%%\section*{References}
 
-%% TODO: Cite shit
+\nocite{*}
+\bibliographystyle{IEEEtran}
+\bibliography{bib}
 

### 2015-02-04 Programming Project A/Insertion Sort Analysis in Java.tex
finished first draft

@@
 All arrays used in testing are Java ArrayList<Integer> unless otherwise specified.
-All times are recorded in milliseconds using a stopwatch class borrowed from [NEED CITATION].
-%% TODO: Add Citation %%
+All times are recorded in milliseconds using a stopwatch class borrowed from Robert Sedgewick and Kevin Wayne \cite{stopwatch}.
 It is important to note that the stopwatch class used takes the elapsed real-time between the start of the insertion sort algorithm and the end of the insertion sort algorithm as opposed to taking the elapsed processor-time because these tests were run on a multi-core computer.
@@
 Adding the statement $<T$ $extends$ $Comparable<T>>$ fixed this issue.
+\cite{erik}
 
@@
 The expected growth of insertion sort as the number of elements (n) grows large can be represented as $ \theta(n^2) $ where $ \theta() $ represents the asymptotically tightly bound running time.
-At $n_0$ the algorithm took 0.0003 milliseconds to complete.
+\cite{textbook} At $n_0$ the algorithm took 0.0003 milliseconds to complete.
 At $n_1$ the algorithm also took 0.0003 milliseconds to complete.

### 2015-02-11 Programming Project A/Insertion Sort Analysis in Java.tex
updated with new notes and programming project B

--- 2015-02-04 Programming Project A/Insertion Sort Analysis in Java.tex
@@
 A sorting algorithm is used to sort data with a natural order.
-One such sorting algorithm is insertion sort, which sorts by iterating through a list of data, taking the current position, and repositioning it into a more appropriate place in the list.
+One such sorting algorithm is insertion sort, which sorts by iterating through a list of data, taking the current position and repositioning it into a more appropriate place in the list.
 How will this procedure perform when handling high volumes of data?
@@
 \Procedure{Insertion-Sort}{A}
-\If{$A.length <= 1$}
-\State{\Return{$A$}}
-\EndIf
 \State{$i = 2$}
@@
 \newline
-\textbf{Outer-Loop Maintenance}: The outer-loop invariant holds because A'[1 ... i - 1] and A[1 ... i - 1] both contain the same elements, although they may be in different orders.
+\textbf{Outer-Loop Maintenance}: The outer-loop invariant holds because A'[1 ... i - 1] and A[1 ... i - 1] both contain the same elements, although they maybe in different orders.
 \newline
@@
 \newline
-\textbf{Inner-Loop Maintenance}: At the beginning of each iteration of the loop the inner-loop invariant holds because j counts down from i, and A'[j+1] is swapped with A'[j] only if A'[j+1] is less than A[j].
+\textbf{Inner-Loop Maintenance}: At the beginning of each iteration of the loop the inner-loop invariants holds because j counts down from i and A'[j+1] is swapped with A'[j] only if A'[j+1] is less than A[j].
 \newline
-\textbf{Inner-Loop Termination}: The negation of the guard implies that j = A.length and that A'[1 ...
-j] has been entirely traversed and sorted in strictly non-decreasing order, which maintains the inner-loop invariant.
+\textbf{Inner-Loop Termination}: The negation of the implies that j = A.length and A'[1 ...
+j] has been entirely traversed and sorted in strictly non-decreasing order which maintains the inner-loop invariant.
 \newline
@@
 It is important to note that the stopwatch class used takes the elapsed real-time between the start of the insertion sort algorithm and the end of the insertion sort algorithm as opposed to taking the elapsed processor-time because these tests were run on a multi-core computer.
-In the table below, A denotes Array. Times in the table below are given as averages out of 10 trials.
+In the table below A denotes Array. Times in the table below are given as averages out of 10 trials.
 \newline
@@
 \subsection{Problems Encountered}
-One major issue encountered during the development of this insertion sort was that after completing the sort, A' was sorted properly except for the first element in the array.
+One major issue encountered during the development of this insertion sort was that after completing the sort, A' was completely sorted properly except for the first element in the array.
 No matter what value the first element of A had, it did not change position in A'.
@@
 Changing the guard for the inner for loop (see Algorithm 1 line 6) from \textit{key < A[j] and j downto 1} to \textit{j downto 1 and key < A[j]} corrected this issue.
-
-Another issue encountered during the development of this program was executing Java $assert$ statements.
-When assertions were wrapped in parenthesis, they evaluated to a boolean, and the assert would throw an $AssertionError$ with the explanation 'cannot compare to boolean'.
-Removing the parenthesis wrapping the assertion fixed this issue.
-
-The last issue encountered during the development process of this program was generic programming.
-The program would not compile because not all Objects have a compareTo method.
-Adding the statement $<T$ $extends$ $Comparable<T>>$ fixed this issue.
-
 %% END TESTING
@@
 \begin{center}
-\includegraphics[scale=.75]{insertion_sort_time_analysis.pdf}
+\includegraphics[width=0.7\linewidth]{insertion_sort_time_analysis.pdf}
 \end{center}
@@
 
-The expected growth of insertion sort as the number of elements (n) grows large can be represented as $ \theta(n^2) $ where $ \theta() $ represents the asymptotically tightly bound running time.
-At $n_0$ the algorithm took 0.0003 milliseconds to complete.
-At $n_1$ the algorithm also took 0.0003 milliseconds to complete.
-For both these values of $n$, the algorithm runs at a constant time because no for-loops are executed.
-As $n$ grows larger the time to complete the experimental data correlates quite accurately with the expected growth.
-For $n_{1*10^6} $ the data matches up perfectly.
-This is to be expected, because as n grows larger the constants and lower orders of the actual running time of the insertion sort start to affect the running time less and less as the highest order of $n^2$ is so large.
-It is expected for lower values of n to not correlate well with $ \theta(n^2) $ because the system running the algorithm may be executing superfolous commands such as checking for system.
-For the data graphed above (see Fig.
-1) this was in fact the case.
-To aid this issue, averages of 10 trails were taken and used in Figure 1.
 %% END EXPERIMENTAL ANALYSIS
@@
 \section{Conclusions}
-Insertion sort is not the best choice for sorting large amounts of data because as n grows large, it takes insertion sort $ \theta(n^2) $ time to complete on modern day computers.
-Obviously, the faster the system executing insertion sort, the faster the algorithm will complete, however, with large enough values of n, it can take powerful systems quite a long amount of time to complete, as shown in Figure 1.
-
-%% TODO: Add more here
 
@@
 \section*{References}
-
-%% TODO: Cite shit
 
@@
 ]{Code/DummyData.java}
-
-The class Stopwatch has not been altered from its original form.
-\newline
 \lstinputlisting[caption=Stopwatch,

### 2015-02-16 Programming Project B/paper.tex
added paper and stuff

--- 2015-02-11 Programming Project A/Insertion Sort Analysis in Java.tex
@@
 %% Title Page
-\title{Insertion Sort Analysis in Java}
+\title{Java Sorting Algorithms Run-Time Analysis}
 \author{Preston Stosur-Bassett}
-\date{February 2, 2015}
+\date{February 23, 2015}
 \maketitle
@@
 \pagestyle{fancy}
-\fancyhead[R]{Insertion Sort Analysis in Java, page \thepage}
+\fancyhead[R]{Java Sorting Algorithms Run-Time Analysis, page \thepage}
 \fancyhead[L]{Preston Stosur-Bassett}
@@
 \begin{abstract}
-%% TODO: Write abstract
+
 %% END ABSTRACT
@@
 \section{Motivation}
-In order to show how an algorithm might run on a given set of hardware, and how the algorithm will perform when given large amounts of data, algorithms are analysed.
-More specifically, insertion sort is analysed to determined what the run time might be for sorting large amounts of data on any modern computer.
+
 %% END MOTIVATION
@@
 \section{Background}
-A sorting algorithm is used to sort data with a natural order.
-One such sorting algorithm is insertion sort, which sorts by iterating through a list of data, taking the current position and repositioning it into a more appropriate place in the list.
-How will this procedure perform when handling high volumes of data?
-How will it perform when executed on different machines?
-Because insertion sort is a more basic sorting algorithm, numerous papers and articles have been written answering these two questions.
+
 %% END BACKGROUND
@@
 \section{Procedure}
-An insertion sort can be implemented in a multitude of languages using the pseudocode provided below.
-\newline
-\textbf{Insertion Sort Pre-Condition}: A is a non-empty array of data with a natural order.
-\newline
-\textbf{Insertion Sort Post-Condition}: A' is a permutation of A (containing all the same elements) in strictly non-decreasing order.
-\begin{algorithm}
-\caption {\textsc{Insertion-Sort}(A)}
-\label{algo:insertionsort}
-\begin{algorithmic}[1]
-\Procedure{Insertion-Sort}{A}
-\State{$i = 2$}
-\While{$i$ upto $A.length$}
-	\State{$key = A[i]$}
-	\State{$j = i - 1$}
-	\While{$j$ downto $1$ and $key < A[j]$}
-		\State{$A[j + 1] = A[j]$}
-		\State{$j = j - 1$}
-	\EndWhile
-	\State{$A[j+1] = key$}
-	\State{$i = i + 1$}
-\EndWhile
-\EndProcedure
-\end{algorithmic}
-\end{algorithm}
-\newline
-\textbf{Outer-Loop Invariant}: The subarray A'[1 ... i - 1] contains all the same elements as the subarray A[1 .. i - 1].
-\newline 
-\textbf{Outer-Loop Initialization}: The outer-loop invariant holds because A'[1 ... i - 1] and A[1 ... i - 1] both contain the same one element.
-\newline
-\textbf{Outer-Loop Maintenance}: The outer-loop invariant holds because A'[1 ... i - 1] and A[1 ... i - 1] both contain the same elements, although they maybe in different orders.
-\newline
-\textbf{Outer-Loop Termination}: When the outer-loop terminates, i = A.length, which implies that the entire array has been traversed and the guard has been negated.
-The negation of the guard implies that A'[1 ...
-i - 1] contains all the elements in A[1 ...
-i - 1].
-\newline
-\newline
-\textbf{Inner-Loop Invariant}: A'[1 ... j] is sorted in strictly non-decreasing order.
-\newline
-\textbf{Inner-Loop Initialization}: Before the first iteration of the loop, j = 1, meaning the subarray A'[1 ... j] contains exactly one element, which is already sorted.
-\newline
-\textbf{Inner-Loop Maintenance}: At the beginning of each iteration of the loop the inner-loop invariants holds because j counts down from i and A'[j+1] is swapped with A'[j] only if A'[j+1] is less than A[j].
-\newline
-\textbf{Inner-Loop Termination}: The negation of the implies that j = A.length and A'[1 ...
-j] has been entirely traversed and sorted in strictly non-decreasing order which maintains the inner-loop invariant.
-\newline
-\newline
-\textbf{Conclusion}: The termination of both the inner and outer loops implies that the entire array has been traversed, A' is a permutation of A containing all the same elements in strictly non-decreasing order.
-This satisfies the post condition.
+
 %% END PROCEDURE
@@
 \subsection{Testing Plan and Results}
-All arrays used in testing are Java ArrayList<Integer> unless otherwise specified.
-All times are recorded in milliseconds using a stopwatch class borrowed from [NEED CITATION].
-%% TODO: Add Citation %%
-It is important to note that the stopwatch class used takes the elapsed real-time between the start of the insertion sort algorithm and the end of the insertion sort algorithm as opposed to taking the elapsed processor-time because these tests were run on a multi-core computer.
-In the table below A denotes Array. Times in the table below are given as averages out of 10 trials.
-\newline
-\captionof{table}{Test Results}
-\begin{center}
-\begin{tabular}{|l|l|l|l|}
-\hline Tested Input & Expected Results & Actual Results & Time \\
-\hline Empty A & Empty A & Empty A & 0.0003 \\
-\hline A of 1000 Strings & Sorted A 1000 Strings & Sorted A 1000 Strings & 0.021 \\
-\hline A 1 Element & Original A & Original A & 0.0003 \\
-\hline A 10 Elements & Sorted A 10 Elements & Sorted A 10 Elements & 0.0005 \\
-\hline A 100 Elements & Sorted A 100 Elements & Sorted A 100 Elements & 0.0021 \\
-\hline A 1000 Elements & Sorted A 1000 Elements & Sorted A 1000 Elements & 0.019 \\
-\hline A 10000 Elements & Sorted A 10000 Elements & Sorted A 10000 Elements & 0.129 \\
-\hline A 100000 Elements & Sorted A 100000 Elements & Sorted A 100000 Elements & 6.4923 \\
-\hline A 1000000 Elements & Sorted A 1000000 Elements & Sorted A 1000000 Elements & 2135.5007 \\
-\hline A 10000000 Elements & Sorted A 10000000 Elements & OS Crash & N/A \\
-\hline A 1000 Identical Elements & Original Array & Original Array & 0.0052 \\
-\hline
-\end{tabular}
-\end{center}
 
 \subsection{Problems Encountered}
-One major issue encountered during the development of this insertion sort was that after completing the sort, A' was completely sorted properly except for the first element in the array.
-No matter what value the first element of A had, it did not change position in A'.
-For example, if A[5, 6, 3, 4, 7] was passed to the insertion sort algorithm, the returned array would look like A'[5, 3, 4, 6, 7].
-Changing the guard for the inner for loop (see Algorithm 1 line 6) from \textit{key < A[j] and j downto 1} to \textit{j downto 1 and key < A[j]} corrected this issue.
+
 %% END TESTING
@@
 \section{Experimental Analysis}
- The insertion sort demonstrated in Algorithm 1 was implemented in Java and executed on an HP SprectreXT TouchSmart with 4 Core Intel i7 processor clocked at 1.9GHz running Ubuntu Gnome 14.10 64-bit.
- \begin{figure}[!]
-\begin{center}
-\includegraphics[width=0.7\linewidth]{insertion_sort_time_analysis.pdf}
-\end{center}
-\captionof{figure}{Insertion Sort Time Analysis}
-\label{fig:insertionsorttimeanalysis}
-\end{figure}
 
@@
 %% BEGIN REFERENCES
-\section*{References}
+%% \section*{References}
+
+\nocite{*}
+\bibligraphystyle{IEEEtran}
+\bibliography{}
 
@@
 \section*{Appendix}
-\lstinputlisting[caption=TestDriver,
-label={TestDriver.java},
-breaklines=true,
-]{Code/TestDriver.java}
-\lstinputlisting[caption=Debug,
-label={Debug.java},
-breaklines=true,
-]{Code/Debug.java}
-\lstinputlisting[caption=DummyData,
-label={DummyData.java},
-breaklines=true,
-]{Code/DummyData.java}
-\lstinputlisting[caption=Stopwatch,
-label={Stopwatch.java}
-breaklines=true,
-]{Code/Stopwatch.java}
-\lstinputlisting[caption=Sort,
-label={Sort.java},
-breaklines=true,
-]{Code/Sort.java}
+
 %% END APPENDIX

### 2015-02-16 Programming Project B/Code/DummyData.java
added initial code

@@ -1,0 +1,103 @@
+/*
+*	@Author Preston Stosur-Bassett
+*	@Date Jan 25, 2015
+*	@Class DummyData
+*	@Description This class contains methods to generate dummy data given a set of parameters.
+*/
+
+import java.util.ArrayList;
+import java.util.Random;
+
+public class DummyData {
+
+	/*
+	*	@Description runArrayList<Integer> will take an ArrayList of Integer Objects and add a given amount of values to it
+	*	@param int end the ending value to denote when to stop adding to the array list
+	*	@param int min the minimum value of the randomly generated data.
+	*	@param int max the maximum value of the randomly generated data.
+	*	@param ArrayList<Integer> list the list to add value to and return
+	*	@return ArrayList<Integer> the list after it has been updated with the randomly generated data
+	*/
+	public static ArrayList<Integer> runArrayList(int end, int min, int max, ArrayList<Integer> list) {
+		Random random = new Random();
+		Debug debugger = new Debug();
+		int start = 0;
+		// INVARIANT: A.length >= start
+		// INITIALIZATION: start = 0, A.length can be longer than 0 when initially passed, but not smaller, so our invariant holds
+		debugger.assertGreatEqual(list.size(), start);
+		while(start < end) {
+			// MAINTANANCE: At the beginning of each iteration, one element was added to A and start was increased by one, therefore, our invariant holds true.
+			debugger.assertGreatEquals(list.size(), start);
+			Integer intToAdd = new Integer(random.nextInt((max - min + 1) + min));
+			list.add(intToAdd);
+
+			//Count up on the iterator
+			start++;
+		}
+		/*TERMINATION: The negation of the guard implies that (end - start) number of elements have been added to A, since start is initialized as 0 at the beginning of the method and is
+			incremented by 1 each iteration of the loop, which means that start amount of elements have been added to A, and so our invariant holds true.	*/
+		debugger.assertGreatEquals(list.size(), start);
+
+		return list;
+	}
+
+	/*
+	*	@Description: runArrayList<String> will take an ArrayList of String Objects and add a given amount of String numerical values to it
+	*	@param int end the ending value to denote when to stop adding to the array list
+	*	@param ArrayList<String> list the list to add String values to and return
+	*	@return ArrayList<String> the list after it has been updated with the randomly generated numerical String values
+	*/
+	public static ArrayList<String> runArrayList(int end, ArrayList<String> list) {
+		Random random = new Random();
+		Debug debugger = new Debug();
[…]

### 2015-02-16 Programming Project B/Code/Sort.java

@@ -1,0 +1,72 @@
+/*
+*	@Author: Preston Stosur-Bassett
+*	@Date: Jan 24, 2015
+*	@Class: Sort
+*	@Description: This class will contain many methods that will sort generic data types using common sorting algorithms.
+*/
+
+import java.util.ArrayList;
+import java.util.List;
+
+public class Sort<T extends Comparable<T>> {
+	/*
+	*	@Pre-Condition: ArrayList<T> unsorted is an unsorted ArrayList of a comparable data type that is non-empty
+	*	@Post-Condition: ArrayList<T> will return a permutation of <code>unsorted</code> that will be in increasing order
+	*	@Description: Insertion will sort an ArrayList of generic type T in increasing order using an insertion sort
+	*	@param ArrayList<T> unsorted is a non-empty unsorted array list of T, where T is a comparable type
+	*	@return sorted is a permutation of <code>unsorted</code> where all the elements are sorted in increasing order
+	*/
+	// INVARIANT (Outer-Loop): The pre condition implies that sorted[0 ... i - 1] will contain all the same data as unsorted[0 ... i - 1].
+	// INVARIANT (Inner-Loop): sorted[0 ... j] is sorted in stricly non-decreasing order.
+	public ArrayList<T> insertion(ArrayList<T> unsorted) {
+		Debug debugger = new Debug<List<T>>();
+		debugger.turnOn();
+		ArrayList<T> sorted = unsorted;
+		if(sorted.size() > 1) {
+			int i = 1;
+			/* INITIALIZATION (Outer-Loop): The invariant holds because i = 1, and there is one element in the subarray of sorted[0 ... i - 1] and unsorted[0 ... i - 1], */
+			List<T> subSortedOI = sorted.subList(0, i - 1);
+			List<T> subUnsortedOI = unsorted.subList(0, i - 1);
+			debugger.assertEquals(subUnsortedOI, subSortedOI);
+
+			while(i < sorted.size()) {
+				/* MAINTENANCE (Outer-Loop): At the beginning of each iteration of the loop, the loop invariant is maintained because the subarray of sorted[0 ... i - 1] contains all the same elements as
+					unsorted[0 ... i - 1] */
+				List<T> subSortedOM = sorted.subList(0, i - 1);
+				List<T> subUnsortedOM = unsorted.subList(0, i - 1);
+				debugger.assertEquals(subUnsortedOM, subSortedOM);
+
+				T value = sorted.get(i);
[…]

### 2015-02-16 Programming Project B/Code/Driver.java
added more useless code to use as a backbone or something

@@ -1,0 +1,12 @@
+/*
+* @Author: Preston Stosur-Bassett
+* @Date: Feb, 23, 2015
+* @Class: Driver
+* @Description: This class will test the functionality of the overall program by serving as a driver that runs through and calls all other required classes
+*/
+
+public class Driver{
+  public static void main(String args[]) {
+    //Do stuff hereturn 
+  }
+}

### 2015-02-16 Programming Project B/Code/Sort.java

@@ -10,6 +10,30 @@ import java.util.List;
 
 public class Sort<T extends Comparable<T>> {
 	/*
+	*	@Pre-Condition:
+	*	@Post-Condition:
+	*	@Description:
+	*	@param
+	*	@return
+	*/
+	//INVARIANTS
+	public ArrayList<T> heap(ArrayList<T> unsorted) {
+
+	}
+
+	/*
+	*	@Pre-Condition:
+	*	@Post-Condition:
+	*	@Description:
+	*	@param
+	*	@return
+	*/
+	// INVARIANTS
+	public ArrayList<T> merge(ArrayList<T> unsorted) {
+
+	}
+
+	/*
 	*	@Pre-Condition: ArrayList<T> unsorted is an unsorted ArrayList of a comparable data type that is non-empty
 	*	@Post-Condition: ArrayList<T> will return a permutation of <code>unsorted</code> that will be in increasing order
 	*	@Description: Insertion will sort an ArrayList of generic type T in increasing order using an insertion sort

### 2015-02-18 Programming Project B/paper.tex
modified insertion sort, added partial merge sort

@@
 %% Title Page
-\title{Java Sorting Algorithms Run-Time Analysis}
+\title{Java Sorting Algorithms Comparison}
 \author{Preston Stosur-Bassett}
@@
 \section{Motivation}
-
+In order to show how an algorithm might run on a given set of hardware, and how the algorithm will perform when given large amounts of data, algorithms are analysed.
+Sorting algorithms sort data into a natural order.
+By analysing sorting algorithms, the fastest algorithm for a given problem can be determined.
 %% END MOTIVATION
@@
 \section{Background}
-
+A sorting algorithm is used to sort data with a natural order.
+One such sorting algorithm is insertion sort, which sorts by iterating through a list of data, taking the current position, and repositioning it into a more appropriate place in the list.
+%% Merge sort and Heap Sort
 %% END BACKGROUND
@@
 %% BEGIN REFERENCES
-%% \section*{References}
+ \section*{References}
 
-\nocite{*}
-\bibligraphystyle{IEEEtran}
-\bibliography{}
+%%\nocite{*}
+%%\bibligraphystyle{IEEEtran}
+%%\bibliography{}
 

### 2015-02-18 Programming Project B/Code/Sort.java

@@ -17,8 +17,20 @@ public class Sort<T extends Comparable<T>> {
 	*	@return
 	*/
 	//INVARIANTS
-	public ArrayList<T> heap(ArrayList<T> unsorted) {
+	public ArrayList<T> heapSort(ArrayList<T> unsorted) {
 
+	}
+
+	/*
+	*	@Pre-Condition:
+	*	@Post-Condition:
+	*	@Description:
+	*	@param
+	*	@param
+	*	@return
+	*/
+	private ArrayList<T> mergeTogether(ArrayList<T> left, ArrayList<T> right) {
+		ArrayList<T> returnValue;
 	}
 
 	/*
@@ -29,20 +41,41 @@ public class Sort<T extends Comparable<T>> {
 	*	@return
 	*/
 	// INVARIANTS
-	public ArrayList<T> merge(ArrayList<T> unsorted) {
-
+	public ArrayList<T> mergeSort(ArrayList<T> unsorted) {
+		ArrayList<T> sorted = unsorted;
+		ArrayList<T> left = new ArrayList<T>();
+		ArrayList<T> right = new ArrayList<T>();
+		ArrayList<T> returnValue;
+		if(sorted.size() <= 1) {
+			returnValue = sorted;
+		}
+		else {
+			int mid = (sorted.size() / 2);
+			for(int i = 0; i < mid; i++) {
+				T temp = sorted.get(i);
+				left.add(temp);
+			}
+			for(int i = mid; i < sorted.size(); i++) {
+				T temp = sorted.get(i);
+				right.add(temp);
+			}
+			left = mergeSort(left);
+			right = mergeSort(right);
+			returnValue = mergeTogether(left, right);
+		}
+		return returnValue;
 	}
 
 	/*
 	*	@Pre-Condition: ArrayList<T> unsorted is an unsorted ArrayList of a comparable data type that is non-empty
 	*	@Post-Condition: ArrayList<T> will return a permutation of <code>unsorted</code> that will be in increasing order
-	*	@Description: Insertion will sort an ArrayList of generic type T in increasing order using an insertion sort
+	*	@Description: insertionSort will sort an ArrayList of generic type T in increasing order using an insertion sort
 	*	@param ArrayList<T> unsorted is a non-empty unsorted array list of T, where T is a comparable type
 	*	@return sorted is a permutation of <code>unsorted</code> where all the elements are sorted in increasing order
 	*/
[…]

### 2015-02-18 Programming Project B/Code/Sort.java
added new project, added a few lines of code

@@ -29,8 +29,12 @@ public class Sort<T extends Comparable<T>> {
 	*	@param
 	*	@return
 	*/
+	//INVARIANTS
 	private ArrayList<T> mergeTogether(ArrayList<T> left, ArrayList<T> right) {
-		ArrayList<T> returnValue;
+		ArrayList<T> combined = new ArrayList<T>();
+		while(left.size() > 0 && right.size() > 0) {
+
+		}
 	}
 
 	/*

### 2015-02-18 Programming Project B/Code/Sort.java
added some code

@@ -33,7 +33,10 @@ public class Sort<T extends Comparable<T>> {
 	private ArrayList<T> mergeTogether(ArrayList<T> left, ArrayList<T> right) {
 		ArrayList<T> combined = new ArrayList<T>();
 		while(left.size() > 0 && right.size() > 0) {
-
+			if(left.get(0).compareTo(right.get(0)) < 0) {
+				combined.add(left.get(0));
+				//left = rest.add(left); ==> WHAT THE HELL IS REST?!
+			}
 		}
 	}
 

### 2015-02-19 Programming Project B/Code/Sort.java
added ore code, im unsure about its usefulness though

@@ -35,9 +35,35 @@ public class Sort<T extends Comparable<T>> {
 		while(left.size() > 0 && right.size() > 0) {
 			if(left.get(0).compareTo(right.get(0)) < 0) {
 				combined.add(left.get(0));
-				//left = rest.add(left); ==> WHAT THE HELL IS REST?!
+				//left = rest(left) ==> WHAT THE HELL IS REST?!
+				//TODO: 
+				//I have no idea if this is actually what im supposed to be doing here...
+				left = left.remove(0);
+			}
+			else {
+				combined.add(right.get(0));
+				// right = rest(right) ==> WHAT THE HELL IS REST?!
+				//TODO:
+				//I have no idea if this is actually what im supposed to be doing here...
+				right = right.remove(0);
 			}
 		}
+		if(left.size() > 0) {
+			//TODO:
+			//I have no idea if this is actually what im supposed to be doing here...
+			for(int i = 0; i < left.size(); i++) {
+				combined.add(left.get(i));
+			}
+		}
+		if(right.size() > 0) {
+			//TODO:
+			//I have no idea if this is actually what im supposed to be doing here...
+			for(int i = 0; i < right.size(); i++) {
+				combined.add(right.get(i));
+			}
+		}
+
+		return combined;
 	}
 
 	/*

### 2015-02-19 Programming Project B/Code/Sort.java
added more documentation

@@ -22,41 +22,42 @@ public class Sort<T extends Comparable<T>> {
 	}
 
 	/*
-	*	@Pre-Condition:
-	*	@Post-Condition:
-	*	@Description:
-	*	@param
-	*	@param
-	*	@return
+	*	@Pre-Condition: TODO: Write this shit.
+	*	@Post-Condition: TODO: Write this shit.
+	*	@Description: mergeTogether is used by the mergeSort method to recombine the left and right sections of the ArrayList<T> that is being sorted by merge sort.
+	*	@param ArrayList<T> left a non-empty ArrayList<T> where T is a comparable data type with a natural order.
+	*	@param ArrayList<T> right a non-empty ArrayList<T> where T is a comparable data type with a natural order.
+	*	@return ArrayList<T> combined should contain all the elements of left and right in stricly non-decreasing order
 	*/
-	//INVARIANTS
+	//TODO: Write the invariants for this shit.
+	//INVARIANTS (There should be around 3 invariants for this method)
 	private ArrayList<T> mergeTogether(ArrayList<T> left, ArrayList<T> right) {
 		ArrayList<T> combined = new ArrayList<T>();
 		while(left.size() > 0 && right.size() > 0) {
 			if(left.get(0).compareTo(right.get(0)) < 0) {
 				combined.add(left.get(0));
 				//left = rest(left) ==> WHAT THE HELL IS REST?!
-				//TODO: 
+				//TODO: Figure out what the fuck I am actually doing here.
 				//I have no idea if this is actually what im supposed to be doing here...
 				left = left.remove(0);
 			}
 			else {
 				combined.add(right.get(0));
 				// right = rest(right) ==> WHAT THE HELL IS REST?!
-				//TODO:
+				//TODO: Figure out what the fuck I am actually doing here.
 				//I have no idea if this is actually what im supposed to be doing here...
 				right = right.remove(0);
 			}
 		}
 		if(left.size() > 0) {
-			//TODO:
+			//TODO: Figure out what the fuck I am actually doing here.
 			//I have no idea if this is actually what im supposed to be doing here...
 			for(int i = 0; i < left.size(); i++) {
 				combined.add(left.get(i));
 			}
 		}
 		if(right.size() > 0) {
-			//TODO:
+			//TODO: Figure out what the fuck I am actually doing here.
 			//I have no idea if this is actually what im supposed to be doing here...
[…]

### 2015-02-19 Programming Project B/paper.tex
added more background for the paper

@@
 One such sorting algorithm is insertion sort, which sorts by iterating through a list of data, taking the current position, and repositioning it into a more appropriate place in the list.
-%% Merge sort and Heap Sort
+Heap sort is another sorting algorithm that greatly differs than insertion sort in that it uses a divide, conquer, and combine method; meaning that it breaks the set it is sorting into subsets until the subsets can no longer be broken up and then merge sort combines the subsets together rendering the correct answer.
+%% Write this for heap sort.
 %% END BACKGROUND

### 2015-02-19 Programming Project B/Code/Sort.java

@@ -24,7 +24,7 @@ public class Sort<T extends Comparable<T>> {
 	/*
 	*	@Pre-Condition: TODO: Write this shit.
 	*	@Post-Condition: TODO: Write this shit.
-	*	@Description: mergeTogether is used by the mergeSort method to recombine the left and right sections of the ArrayList<T> that is being sorted by merge sort.
+	*	@Description: mergeTogether is used by the mergeSort method to recombine the left and right sections of the ArrayList<T> that is being sorted by merge sort. Note that this is a helper method for the mergeSort method, and should not be called externally of this class.
 	*	@param ArrayList<T> left a non-empty ArrayList<T> where T is a comparable data type with a natural order.
 	*	@param ArrayList<T> right a non-empty ArrayList<T> where T is a comparable data type with a natural order.
 	*	@return ArrayList<T> combined should contain all the elements of left and right in stricly non-decreasing order

### 2015-02-19 Programming Project B/Code/Sort.java
shoudl have working merge sort at this point

@@ -33,37 +33,34 @@ public class Sort<T extends Comparable<T>> {
 	//INVARIANTS (There should be around 3 invariants for this method)
 	private ArrayList<T> mergeTogether(ArrayList<T> left, ArrayList<T> right) {
 		ArrayList<T> combined = new ArrayList<T>();
-		while(left.size() > 0 && right.size() > 0) {
-			if(left.get(0).compareTo(right.get(0)) < 0) {
-				combined.add(left.get(0));
-				//left = rest(left) ==> WHAT THE HELL IS REST?!
-				//TODO: Figure out what the fuck I am actually doing here.
-				//I have no idea if this is actually what im supposed to be doing here...
-				left = left.remove(0);
+		int i = 0;
+		int y = 0;
+		int x = 0;
+
+		while(left.size() != i && right.size() != y) {
+			if(left.get(i).compareTo(right.get(y)) < 0) {
+				combined.add(x, left.get(i));
+				i++;
+				x++;
 			}
 			else {
-				combined.add(right.get(0));
-				// right = rest(right) ==> WHAT THE HELL IS REST?!
-				//TODO: Figure out what the fuck I am actually doing here.
-				//I have no idea if this is actually what im supposed to be doing here...
-				right = right.remove(0);
-			}
-		}
-		if(left.size() > 0) {
-			//TODO: Figure out what the fuck I am actually doing here.
-			//I have no idea if this is actually what im supposed to be doing here...
-			for(int i = 0; i < left.size(); i++) {
-				combined.add(left.get(i));
-			}
-		}
-		if(right.size() > 0) {
-			//TODO: Figure out what the fuck I am actually doing here.
-			//I have no idea if this is actually what im supposed to be doing here...
-			for(int i = 0; i < right.size(); i++) {
-				combined.add(right.get(i));
+				combined.add(x, right.get(y));
+				y++;
+				x++;
 			}
 		}
 
+		while(left.size() != i) {
+			combined.add(x, left.get(i));
+			i++;
+			x++;
+		}
+
+		while(right.size() != y) {
+			combined.add(x, right.get(y));
+			y++;
+			x++;
+		}
 		return combined;
 	}
 

### 2015-02-19 Programming Project B/Code/Driver.java
added testing to driver

@@ -7,6 +7,15 @@
 
 public class Driver{
   public static void main(String args[]) {
-    //Do stuff hereturn 
+    DummyData testData = new DummyData();
+
+    ArrayList<Integer> testList = new ArrayList<Integer>();
+    testList = testData.runArrayList(10, 1, 25, testList);
+
+    Sort sorter = new Sort();
+
+    testList = sorter.mergeSort(testList);
+
+    System.out.println(testList);
   }
 }

### 2015-02-19 Programming Project B/Code/Debug.java
working merge sort and test

@@ -98,7 +98,7 @@ public class Debug<T> {
 	*	@param int large an integer primative value to assert is greator than or equal to the second arguement
 	*	@param int small an integer primative value to assert the first arguement is greator than or equal to.
 	*/
-	public void assertGreatEqual(int large, int small) {
+	public void assertGreatEquals(int large, int small) {
 		if(debugOn == true) {
 			assert large >= small;
 		}
@@ -109,7 +109,7 @@ public class Debug<T> {
 	*	@param int small an integer primative value to assert is less than or equal to the second arguement
 	*	@param int large an integer primative value to assert the first arguement is less than or equal to.
 	*/
-	public void assertLessEqual(int small, int large) {
+	public void assertLessEquals(int small, int large) {
 		if(debugOn == true) {
 			assert small <= large;
 		}

### 2015-02-19 Programming Project B/Code/Driver.java

@@ -5,6 +5,8 @@
 * @Description: This class will test the functionality of the overall program by serving as a driver that runs through and calls all other required classes
 */
 
+import java.util.ArrayList;
+
 public class Driver{
   public static void main(String args[]) {
     DummyData testData = new DummyData();
@@ -12,10 +14,14 @@ public class Driver{
     ArrayList<Integer> testList = new ArrayList<Integer>();
     testList = testData.runArrayList(10, 1, 25, testList);
 
+    System.out.println("Unsorted list: ");
+    System.out.println(testList);
+
     Sort sorter = new Sort();
 
     testList = sorter.mergeSort(testList);
 
+    System.out.println("Sorted List: ");
     System.out.println(testList);
   }
 }

### 2015-02-19 Programming Project B/Code/DummyData.java

@@ -24,7 +24,7 @@ public class DummyData {
 		int start = 0;
 		// INVARIANT: A.length >= start
 		// INITIALIZATION: start = 0, A.length can be longer than 0 when initially passed, but not smaller, so our invariant holds
-		debugger.assertGreatEqual(list.size(), start);
+		debugger.assertGreatEquals(list.size(), start);
 		while(start < end) {
 			// MAINTANANCE: At the beginning of each iteration, one element was added to A and start was increased by one, therefore, our invariant holds true.
 			debugger.assertGreatEquals(list.size(), start);

### 2015-02-19 Programming Project B/Code/Sort.java

@@ -17,9 +17,9 @@ public class Sort<T extends Comparable<T>> {
 	*	@return
 	*/
 	//INVARIANTS
-	public ArrayList<T> heapSort(ArrayList<T> unsorted) {
+	/*public ArrayList<T> heapSort(ArrayList<T> unsorted) {
 
-	}
+	}*/
 
 	/*
 	*	@Pre-Condition: TODO: Write this shit.

### 2015-02-19 Programming Project B/paper.tex
added insertion sort pseudocode to paper and pre/post conditions to mergeTogether method

@@
 \section{Procedure}
-
+An insertion sort can be implemented in a multitude of languages using the pseudocode provided below.
+\newline
+\textbf{Insertion Sort Pre-Condition}: A is a non-empty array of data with a natural order.
+\newline
+\textbf{Insertion Sort Post-Condition}: A' is a permutation of A (containing all the same elements) in strictly non-decreasing order.
+\begin{algorithm}
+\caption {\textsc{Insertion-Sort}(A)}
+\label{algo:insertionsort}
+\begin{algorithmic}[1]
+\Procedure{Insertion-Sort}{A}
+\If{$A.length <= 1$}
+\State{\Return{$A$}}
+\EndIf
+\State{$i = 2$}
+\While{$i$ upto $A.length$}
+	\State{$key = A[i]$}
+	\State{$j = i - 1$}
+	\While{$j$ downto $1$ and $key < A[j]$}
+		\State{$A[j + 1] = A[j]$}
+		\State{$j = j - 1$}
+	\EndWhile
+	\State{$A[j+1] = key$}
+	\State{$i = i + 1$}
+\EndWhile
+\EndProcedure
+\end{algorithmic}
+\end{algorithm}
+\newline
+\textbf{Insertion Sort Outer-Loop Invariant}: The subarray A'[1 ... i - 1] contains all the same elements as the subarray A[1 .. i - 1].
+\newline 
+\textbf{Insertion Sort Outer-Loop Initialization}: The outer-loop invariant holds because A'[1 ... i - 1] and A[1 ... i - 1] both contain the same one element.
+\newline
+\textbf{Insertion Sort Outer-Loop Maintenance}: The outer-loop invariant holds because A'[1 ... i - 1] and A[1 ... i - 1] both contain the same elements, although they may be in different orders.
+\newline
+\textbf{Insertion Sort Outer-Loop Termination}: When the outer-loop terminates, i = A.length, which implies that the entire array has been traversed and the guard has been negated.
+The negation of the guard implies that A'[1 ...
+i - 1] contains all the elements in A[1 ...
+i - 1].
+\newline
+\newline
+\textbf{Insertion Sort Inner-Loop Invariant}: A'[1 ... j] is sorted in strictly non-decreasing order.
+\newline
+\textbf{Insertion Sort Inner-Loop Initialization}: Before the first iteration of the loop, j = 1, meaning the subarray A'[1 ... j] contains exactly one element, which is already sorted.
+\newline
+\textbf{Insertion Sort Inner-Loop Maintenance}: At the beginning of each iteration of the loop the inner-loop invariant holds because j counts down from i, and A'[j+1] is swapped with A'[j] only if A'[j+1] is less than A[j].
+\newline
+\textbf{Insertion Sort Inner-Loop Termination}: The negation of the guard implies that j = A.length and that A'[1 ...
+j] has been entirely traversed and sorted in strictly non-decreasing order, which maintains the inner-loop invariant.
+\newline
+\newline
+\textbf{Insertion Sort Conclusion}: The termination of both the inner and outer loops implies that the entire array has been traversed, A' is a permutation of A containing all the same elements in strictly non-decreasing order.
+This satisfies the post condition.
 %% END PROCEDURE

### 2015-02-19 Programming Project B/Code/Sort.java

@@ -22,8 +22,8 @@ public class Sort<T extends Comparable<T>> {
 	}*/
 
 	/*
-	*	@Pre-Condition: TODO: Write this shit.
-	*	@Post-Condition: TODO: Write this shit.
+	*	@Pre-Condition: ArrayList<T> left is a non-empty sorted array in stricly non-decreasing order where T is a comparable data type with a natural order
+	*	@Post-Condition: ArrayList<T> right is a non-empty sorted array in strictly non-decreasing order where T is a comparable data type with a natural order
 	*	@Description: mergeTogether is used by the mergeSort method to recombine the left and right sections of the ArrayList<T> that is being sorted by merge sort. Note that this is a helper method for the mergeSort method, and should not be called externally of this class.
 	*	@param ArrayList<T> left a non-empty ArrayList<T> where T is a comparable data type with a natural order.
 	*	@param ArrayList<T> right a non-empty ArrayList<T> where T is a comparable data type with a natural order.

### 2015-02-19 Programming Project B/paper.tex
minor updates in formatting to paper
added merge-sort algorithm to the paper

@@
 \Procedure{Insertion-Sort}{A}
-\If{$A.length <= 1$}
+\If{$A.length < 2$}
 \State{\Return{$A$}}
@@
 \EndWhile
+\State{\Return{$A$}}
 \EndProcedure
@@
 This satisfies the post condition.
+\newline
+\newline
+
+A merge sort can be implemented in a variety of languages using the pseudocode provided below. 
+\newline
+\textbf{Merge Sort Pre-Condition}: A is a non-empty array of comparable data with a natural order.
+\newline
+\textbf{Merge Sort Post-Condition}: A' is a permutation of A (containing all the same elements) in stricly non-decreasing order.
+\begin{algorithm}
+\caption {\textsc{Merge-Sort}(A)}
+\label{algo:mergesort}
+\begin{algorithmic}[1]
+\Procedure{Merge-Sort}{A}
+\If{$A.length < 2$}
+\State{\Return{$A$}}
+\EndIf
+\State{$mid = A.length / 2$}
+\For{$i = 1$ upto $mid$}
+	\State{$left[left.length] = A[i]$}
+\EndFor
+\For{$i = mid$ upto $A.length$}
+	\State{$right[right.length] = A[i]$}
+\EndFor
+\State{$left =$ Merge-Sort($left$)}
+\State{$right =$ Merge-Sort($right$)}
+\State{$A =$ Merge($left$, $right$)}
+\State{\Return{$A$}}
+\EndProcedure
+\end{algorithmic}
+\end{algorithm}
+\newline
 %% END PROCEDURE

### 2015-02-19 Programming Project B/paper.tex
added merge-together method to paper

@@
 \section{Procedure}
-An insertion sort can be implemented in a multitude of languages using the pseudocode provided below.
+An insertion sort can be implemented in a multitude of languages using the pseudocode provided in Algorithm 1.
 \newline
@@
 
-A merge sort can be implemented in a variety of languages using the pseudocode provided below. 
+A merge sort can be implemented in a variety of languages using the pseudocode provided in Algorithm 2.
 \newline
@@
 \State{\Return{$A$}}
+\EndProcedure
+\newline
+\Procedure{Merge}{left, right}
+\State{$var$ $i = 0$}
+\State{$var$ $y = 0$}
+\State{$var$ $x = 0$}
+\While{$left.length$ $!=$ $i$ and $right.length$ $!=$ $y$}
+	\If{$left[i] < right[y]$}
+		\State{$combined[x] = left[i]$}
+		\State{$i = i + 1$}
+		\State{$x = x + 1$}
+	\EndIf
+	\If{$right[y] < left[i]$}
+		\State{$combned[x] = right[y]$}
+		\State{$y = y + 1$}
+		\State{$x = x + 1$}
+	\EndIf
+\EndWhile
+\While{$left.length$ $!=$ $i$}
+	\State{$combined[x] = left[i]$}
+	\State{$i = i + 1$}
+	\State{$x = x + 1$}
+\EndWhile
+\While{$right.length$ $!=$ $y$}
+	\State{$combined[x] = right[y]$}
+	\State{$y = y + 1$}
+	\State{$x = x + 1$}
+\EndWhile
+\State{\Return{$combined$}}
 \EndProcedure

### 2015-02-19 Programming Project B/paper.tex
added merge-together method pre/post conditions

@@
 \newline
-\textbf{Merge Sort Pre-Condition}: A is a non-empty array of comparable data with a natural order.
+\textbf{Merge Sort Pre-Condition}: A is a non-empty array of a comparable data with a natural order.
 \newline
-\textbf{Merge Sort Post-Condition}: A' is a permutation of A (containing all the same elements) in stricly non-decreasing order.
+\textbf{Merge Sort Post-Condition}: A' is a permutation of A (containing all the same elements) in strictly non-decreasing order.
+\newline
+\textbf{Merge Pre-Condition}: Left and right are both non-empty arrays of a comparable data type with a natural order in strictly non-decreasing order.
+\newline
+\textbf{Merge Post-Condition}: Combined has all the elements of both left and right in strictly non-decreasing order.
 \begin{algorithm}

### 2015-02-19 Programming Project B/Code/Sort.java
added first attempt at heap sort

@@ -17,9 +17,50 @@ public class Sort<T extends Comparable<T>> {
 	*	@return
 	*/
 	//INVARIANTS
-	/*public ArrayList<T> heapSort(ArrayList<T> unsorted) {
+	private ArrayList<T> heapify(ArrayList<T> unsorted, int i, int total) {
+		int left = i * 2;
+		int right = left + 1;
+		int originalI = i;
 
-	}*/
+		if(left <= total && unsorted[left].compareTo(unsorted[i]) > 0) {
+			i = left;
+		}
+		if(right <= total && unsorted[right].compareTo(unsorted[i]) > 0) {
+			i = right;
+		}
+		if(i != originalI) {
+			T tmp = unsorted.get(originalI);
+			unsorted.set(originalI, unsorted.get(i));
+			unsorted.set(i, tmp);
+			unsorted = heapify(unsorted, i);
+		}
+
+		return unsorted;
+	}
+
+	/*
+	*	@Pre-Condition:
+	*	@Post-Condition:
+	*	@Description:
+	*	@param
+	*	@return
+	*/
+	//INVARIANTS
+	public ArrayList<T> heapSort(ArrayList<T> unsorted) {
+		int arrSize = unsorted.size();
+		for(int i = arrSize / 2; i >= 0; i--) {
+			heapify(unsorted, i);
+		}
+		for(int i = arrSize; i > 0; i--) {
+			T tmp = unsorted.get(0);
+			unsorted.set(0, unsorted.get(i));
+			unsorted.set(i, tmp);
+			arrSize--;
+			unsorted = heapify(arr, 0, arrSize);
+		}
+
+		return unsorted;
+	}
 
 	/*
 	*	@Pre-Condition: ArrayList<T> left is a non-empty sorted array in stricly non-decreasing order where T is a comparable data type with a natural order

### 2015-02-19 Programming Project B/Code/Driver.java
fixed compile time errors

@@ -19,7 +19,8 @@ public class Driver{
 
     Sort sorter = new Sort();
 
-    testList = sorter.mergeSort(testList);
+    //testList = sorter.mergeSort(testList);
+    testList = sorter.heapSort(testList);
 
     System.out.println("Sorted List: ");
     System.out.println(testList);

### 2015-02-19 Programming Project B/Code/Sort.java

@@ -22,17 +22,17 @@ public class Sort<T extends Comparable<T>> {
 		int right = left + 1;
 		int originalI = i;
 
-		if(left <= total && unsorted[left].compareTo(unsorted[i]) > 0) {
+		if(left <= total && unsorted.get(left).compareTo(unsorted.get(i)) > 0) {
 			i = left;
 		}
-		if(right <= total && unsorted[right].compareTo(unsorted[i]) > 0) {
+		if(right <= total && unsorted.get(right).compareTo(unsorted.get(i)) > 0) {
 			i = right;
 		}
 		if(i != originalI) {
 			T tmp = unsorted.get(originalI);
 			unsorted.set(originalI, unsorted.get(i));
 			unsorted.set(i, tmp);
-			unsorted = heapify(unsorted, i);
+			unsorted = heapify(unsorted, i, total);
 		}
 
 		return unsorted;
@@ -49,14 +49,14 @@ public class Sort<T extends Comparable<T>> {
 	public ArrayList<T> heapSort(ArrayList<T> unsorted) {
 		int arrSize = unsorted.size();
 		for(int i = arrSize / 2; i >= 0; i--) {
-			heapify(unsorted, i);
+			unsorted = heapify(unsorted, i, arrSize);
 		}
 		for(int i = arrSize; i > 0; i--) {
 			T tmp = unsorted.get(0);
 			unsorted.set(0, unsorted.get(i));
 			unsorted.set(i, tmp);
 			arrSize--;
-			unsorted = heapify(arr, 0, arrSize);
+			unsorted = heapify(unsorted, 0, arrSize);
 		}
 
 		return unsorted;

### 2015-02-19 Programming Project B/Code/Sort.java
fixed run-time errors

@@ -47,7 +47,7 @@ public class Sort<T extends Comparable<T>> {
 	*/
 	//INVARIANTS
 	public ArrayList<T> heapSort(ArrayList<T> unsorted) {
-		int arrSize = unsorted.size();
+		int arrSize = unsorted.size() - 1;
 		for(int i = arrSize / 2; i >= 0; i--) {
 			unsorted = heapify(unsorted, i, arrSize);
 		}

### 2015-02-19 Programming Project B/Code/Sort.java
updated documentation'

@@ -10,9 +10,11 @@ import java.util.List;
 
 public class Sort<T extends Comparable<T>> {
 	/*
-	*	@Pre-Condition:
+	*	@Pre-Condition: ArrayList<T> is a non-empty set of data where T is a comparable data type with a natural order
 	*	@Post-Condition:
-	*	@Description:
+	*	@Description: heapify is a helper method for heapSort that keeps the heap in order so that the root node is the largest element in the heap.
+	*	@param ArrayList<T> unsorted is a non-empty set of data where T is a comparable data type with a natural order
+	*	@param
 	*	@param
 	*	@return
 	*/
@@ -39,13 +41,13 @@ public class Sort<T extends Comparable<T>> {
 	}
 
 	/*
-	*	@Pre-Condition:
-	*	@Post-Condition:
-	*	@Description:
-	*	@param
-	*	@return
+	*	@Pre-Condition: ArrayList<T> unsorted is a non-empty ArrayList<T> where T is a comparable data type with a natural order.
+	*	@Post-Condition: ArrayList<T> sorted is a permutation of unsorted (it contains all the same elements) in stricly non-decreasing order
+	*	@Description: heapSort will sort a given set of data in an ArrayList<T> in strictly non-decreasing order using the heap sort method.
+	*	@param ArrayList<T> unsorted is a non-empty ArrayList<T> where T is a comparable data type with a natural order
+	*	@return ArrayList<T> sorted is a permutation of unsorted in strictly non-decreasing order
 	*/
-	//INVARIANTS
+	//INVARIANTS (There should be two of them here)
 	public ArrayList<T> heapSort(ArrayList<T> unsorted) {
 		int arrSize = unsorted.size() - 1;
 		for(int i = arrSize / 2; i >= 0; i--) {
@@ -58,8 +60,8 @@ public class Sort<T extends Comparable<T>> {
 			arrSize--;
 			unsorted = heapify(unsorted, 0, arrSize);
 		}
-
-		return unsorted;
+		ArrayList<T> sorted = unsorted;
+		return sorted;
 	}
 
 	/*

### 2015-02-20 Programming Project B/Code/Sort.java
removed invariants placeholder in heapify

@@ -18,7 +18,7 @@ public class Sort<T extends Comparable<T>> {
 	*	@param
 	*	@return
 	*/
-	//INVARIANTS
+	//NOTE: No Invariants as this has no loops
 	private ArrayList<T> heapify(ArrayList<T> unsorted, int i, int total) {
 		int left = i * 2;
 		int right = left + 1;

### 2015-02-20 Programming Project B/Code/Driver.java
started testing for speed

@@ -12,17 +12,19 @@ public class Driver{
     DummyData testData = new DummyData();
 
     ArrayList<Integer> testList = new ArrayList<Integer>();
-    testList = testData.runArrayList(10, 1, 25, testList);
+    testList = testData.runArrayList(1000000, 1, 10000000, testList);
 
     System.out.println("Unsorted list: ");
     System.out.println(testList);
 
     Sort sorter = new Sort();
 
-    //testList = sorter.mergeSort(testList);
-    testList = sorter.heapSort(testList);
+    Stopwatch watchman = new Stopwatch();
+    testList = sorter.mergeSort(testList);
+    //testList = sorter.heapSort(testList);
 
     System.out.println("Sorted List: ");
     System.out.println(testList);
+    System.out.println("Time To Complete: "+watchman.elapsedTime());
   }
 }

### 2015-02-22 Programming Project B/Code/Sort.java
added heapsort pseudocode to paper

@@ -14,8 +14,8 @@ public class Sort<T extends Comparable<T>> {
 	*	@Post-Condition:
 	*	@Description: heapify is a helper method for heapSort that keeps the heap in order so that the root node is the largest element in the heap.
 	*	@param ArrayList<T> unsorted is a non-empty set of data where T is a comparable data type with a natural order
-	*	@param
-	*	@param
+	*	@param int i
+	*	@param int total
 	*	@return
 	*/
 	//NOTE: No Invariants as this has no loops

### 2015-02-22 Programming Project B/paper.tex
added heapsort pseudocode to paper
gathered more testing results

@@
 \newline
+%% TODO: Add pre and post conditions here
+A heap sort can be implemented in a variety of languages using the pseudocode below in Algorithm 3.
+\newline
+\begin{algorithm}
+\caption {\textsc{Heap-Sort}(A)}
+\label{algo:heapsort}
+\begin{algorithmic}[1]
+\Procedure{Heapify}{A, i, total}
+\State{$var$ $left = i * 2$}
+\State{$var$ $right = left + 1$}
+\State{$var$ $iPrime = i$}
+\If{$left <= total$ and $A[left] > A[i]$}
+	\State{$i = left$}
+\EndIf
+\If{$right <= total$ and $A[right] > A[i]$}
+	\State{$i = right$}
+\EndIf
+\If{$i != iPrime$}
+	\State{$var$ $temp = A[iPrime]$}
+	\State{$A[iPrime] = A[i]$}
+	\State{$A[i] = tmp$}
+	\State{$A = heapify(A, i, total)$}
+\EndIf
+\State{\Return{$A$}}
+\EndProcedure
+\Procedure{Heap-Sort}{A}
+\State{$var$ $size = A.length$}
+\For{$var$ $i = size / 2$ downto $1$}
+	\State{$A = heapify(A, i, size$}
+\EndFor
+\For{$var$ $i = size$ downto $1$}
+	\State{$var$ $tmp = A[1]$}
+	\State{$A[0] = A[i]$}
+	\State{$A[i] = tmp$}
+	\State{$size = size - 1$}
+	\State{$A = heapify(A, 1, size)$}
+\EndFor
+\State{\Return{$A$}}
+\EndProcedure
+\end{algorithmic}
+\end{algorithm}
 %% END PROCEDURE
@@
 \subsection{Testing Plan and Results}
-
+All arrays used in testing are Java ArrayList<Integer> unless otherwise spcified.
+All times are recorded in milliseconds using a stopwatch class borrowed from Robert Sedgewock and Kevin Wayne %%TODO: Cite this shit.
+It is important to note that the soptwatch class used takes the elapsed real-time between the start of the sort algorithm and the end of the sort algorithim as opposed to taking the elapsed processor-time becasue these tests were run on a multi-core computer.
+In the table below, A denotes Array. Times in the table below are given as averages out of 10 trials.
+\newline
+\captionof{table}{Insertion Sort Test Results}
+\begin{center}
+\begin{tabular}{|l|l|l|l|}
+\hline Tested Input & Expected Results & Actual Results & Time \\
+\hline Empty A & Empty A & Empty A & 0.0003 \\
+\hline A of 1000 Strings & Sorted A 1000 Strings & Sorted A 1000 Strings & 0.021 \\
+\hline A 1 Element & Original A & Original A & 0.0003 \\
+\hline A 10 Elements & Sorted A 10 Elements & Sorted A 10 Elements & 0.0005 \\
+\hline A 100 Elements & Sorted A 100 Elements & Sorted A 100 Elements & 0.0021 \\
+\hline A 1000 Elements & Sorted A 1000 Elements & Sorted A 1000 Elements & 0.019 \\
+\hline A 10000 Elements & Sorted A 10000 Elements & Sorted A 10000 Elements & 0.129 \\
+\hline A 100000 Elements & Sorted A 100000 Elements & Sorted A 100000 Elements & 6.4923 \\
+\hline A 1000000 Elements & Sorted A 1000000 Elements & Sorted A 1000000 Elements & 2135.5007 \\
+\hline A 10000000 Elements & Sorted A 10000000 Elements & OS Crash & N/A \\
+\hline A 1000 Identical Elements & Original Array & Original Array & 0.0052 \\
+\hline
+\end{tabular}
+\end{center}
+\newline
+\captionof{table}{Merge Sort Test Results}
+\begin{center}
+\begin{tabular}{|l|l|l|l|}
+\hline Tested Input & Expected Results & Actual Results & Time \\
+\hline Empty A & Empty A & Empty A & 0.0002 \\
+\hline A 1 Element & Original A & Original A & 0.0001 \\
+\hline A 10 Elements & Sorted A 10 Elements & Sorted A 10 Elements & 0.0004 \\
+\hline A 100 Elements & Sorted A 100 Elements & Sorted A 100 Elements & 0.0028 \\
+\hline A 1000 Elements & Sorted A 1000 Elements & Sorted A 1000 Elements & 0.0294 \\
+\hline A 10000 Elements & Sorted A 10000 Elements & Sorted A 10000 Elements & 0.2098 \\
+\hline A 100000 Elements & Sorted A 100000 Elements & Sorted A 100000 Elements & 2.1904 \\
+\hline A 1000000 Elements & Sorted A 1000000 Elements & Sorted A 1000000 Elements & 22.9314 \\
+\hline 
+\hline
+\end{tabular}
+\end{center}
 \subsection{Problems Encountered}

### 2015-02-22 Programming Project B/Code/Driver.java

@@ -11,8 +11,9 @@ public class Driver{
   public static void main(String args[]) {
     DummyData testData = new DummyData();
 
-    ArrayList<Integer> testList = new ArrayList<Integer>();
-    testList = testData.runArrayList(1000000, 1, 10000000, testList);
+    ArrayList<String> testList = new ArrayList<String>();
+    //testList = testData.runArrayList(1000, 0, 0, testList);
+    testList = testData.runArrayList(1000, testList);
 
     System.out.println("Unsorted list: ");
     System.out.println(testList);

### 2015-02-22 Programming Project B/Code/Sort.java

@@ -16,7 +16,7 @@ public class Sort<T extends Comparable<T>> {
 	*	@param ArrayList<T> unsorted is a non-empty set of data where T is a comparable data type with a natural order
 	*	@param int i
 	*	@param int total
-	*	@return
+	*	@return ArrayList<T> unsorted 
 	*/
 	//NOTE: No Invariants as this has no loops
 	private ArrayList<T> heapify(ArrayList<T> unsorted, int i, int total) {

### 2015-02-22 Programming Project B/Code/Driver.java
updated shit

@@ -11,18 +11,18 @@ public class Driver{
   public static void main(String args[]) {
     DummyData testData = new DummyData();
 
-    ArrayList<String> testList = new ArrayList<String>();
-    //testList = testData.runArrayList(1000, 0, 0, testList);
-    testList = testData.runArrayList(1000, testList);
+    ArrayList<Integer> testList = new ArrayList<Integer>();
+    testList = testData.runArrayList(10000000, 0, 10000000, testList);
+    //testList = testData.runArrayList(1000, testList);
 
     System.out.println("Unsorted list: ");
-    System.out.println(testList);
+    //System.out.println(testList);
 
     Sort sorter = new Sort();
 
     Stopwatch watchman = new Stopwatch();
-    testList = sorter.mergeSort(testList);
-    //testList = sorter.heapSort(testList);
+    //testList = sorter.mergeSort(testList);
+    testList = sorter.heapSort(testList);
 
     System.out.println("Sorted List: ");
     System.out.println(testList);

### 2015-02-22 Programming Project B/paper.tex
updated shit
added problems encountered section

@@
 \end{center}
-\newline
 \captionof{table}{Merge Sort Test Results}
@@
 \hline Empty A & Empty A & Empty A & 0.0002 \\
+\hline A of 1000 Strings & Sorted A 1000 Strings & Sorted A 1000 Strings & 0.0297 \\
 \hline A 1 Element & Original A & Original A & 0.0001 \\
@@
 \hline A 1000000 Elements & Sorted A 1000000 Elements & Sorted A 1000000 Elements & 22.9314 \\
+\hline A 10000000 Elements & Sorted A 10000000 Elements & Sorted A 10000000 Elements & 241.0322 \\
+\hline A 1000 Identical Elements & Original A & Original A & 0.0298 \\
 \hline
+\end{tabular}
+\end{center}
+\captionof{table}{Heap Sort Test Results}
+\begin{center}
+\begin{tabular}{|l|l|l|l|}
+\hline Tested Input & Expected Results & Actual Results & Time \\
+\hline Empty A & Empty A & Empty A & 0.0002 \\
+\hline A of 1000 Strings & Sorted A 1000 Strings & Sorted A 1000 Strings & .0188 \\
+\hline A 1 Element & Original A & Original A & 0.0003 \\
+\hline A 10 Elements & Sorted A 10 Elements & Sorted A 10 Elements & 0.0006 \\
+\hline A 100 Elements & Sorted A 100 Elements & Sorted A 100 Elements & 0.002 \\
+\hline A 1000 Elements & Sorted A 1000 Elements & Sorted A 1000 Elements & 0.0162 \\
+\hline A 10000 Elements & Sorted A 10000 Elements & Sorted A 10000 Elements & 0.0468 \\
+\hline A 100000 Elements & Sorted A 100000 Elements & Sorted A 100000 Elements & 0.889 \\
+\hline A 1000000 Elements & Sorted A 1000000 Elements & Sorted A 1000000 Elements & 23.0635 \\
+\hline A 10000000 Elements & Sorted A 10000000 Elements & Sorted A 10000000 Elements & 244.1952 \\
+\hline A 1000 Identical Elements & Original A & Original A & 0.0156 \\
 \hline
@@
 \subsection{Problems Encountered}
-
+No problems were encountered.
 %% END TESTING
@@
 \section*{Appendix}
+\lstinputlisting[caption=Driver,
+label={Driver.java},
+breaklines=true,
+]{Code/Driver.java}
+\lstinputlisting[caption=Debug,
+label={Debug.java},
+breaklines=true,
+]{Code/Debug.java}
+\lstinputlisting[caption=DummyData,
+label={DummyData.java},
+breaklines=true,
+]{Code/DummyData.java}
 
+The class Stopwatch has not been altered from its original form.
+\newline
+\lstinputlisting[caption=Stopwatch,
+label={Stopwatch},
+breaklines=true,
+]{Code/Stopwatch.java}
+\lstinputlisting[caption=Sort,
+label={Sort.java},
+breaklines=true,
+]{Code/Sort.java}
 %% END APPENDIX

### 2015-02-23 Programming Project B/Code/Sort.java
added the invariant outline

@@ -16,9 +16,8 @@ public class Sort<T extends Comparable<T>> {
 	*	@param ArrayList<T> unsorted is a non-empty set of data where T is a comparable data type with a natural order
 	*	@param int i
 	*	@param int total
-	*	@return ArrayList<T> unsorted 
+	*	@return ArrayList<T> unsorted
 	*/
-	//NOTE: No Invariants as this has no loops
 	private ArrayList<T> heapify(ArrayList<T> unsorted, int i, int total) {
 		int left = i * 2;
 		int right = left + 1;
@@ -47,18 +46,23 @@ public class Sort<T extends Comparable<T>> {
 	*	@param ArrayList<T> unsorted is a non-empty ArrayList<T> where T is a comparable data type with a natural order
 	*	@return ArrayList<T> sorted is a permutation of unsorted in strictly non-decreasing order
 	*/
-	//INVARIANTS (There should be two of them here)
+	//Invariant for First For Loop: TODO: write this
+	//Invariant for Second For Loop: TODO: write this
 	public ArrayList<T> heapSort(ArrayList<T> unsorted) {
 		int arrSize = unsorted.size() - 1;
-		for(int i = arrSize / 2; i >= 0; i--) {
+		for(int i = arrSize / 2; i >= 0; i--) { //Initialization: TODO: write this
+			//Maintanance: TODO: write this
 			unsorted = heapify(unsorted, i, arrSize);
+			//Termination: TODO: write this
 		}
-		for(int i = arrSize; i > 0; i--) {
+		for(int i = arrSize; i > 0; i--) { //Initialization: TODO: write this
+			//Maintanance: TODO: write this
 			T tmp = unsorted.get(0);
 			unsorted.set(0, unsorted.get(i));
 			unsorted.set(i, tmp);
 			arrSize--;
 			unsorted = heapify(unsorted, 0, arrSize);
+			//Termination: TODO: write this
 		}
 		ArrayList<T> sorted = unsorted;
 		return sorted;
@@ -72,15 +76,17 @@ public class Sort<T extends Comparable<T>> {
 	*	@param ArrayList<T> right a non-empty ArrayList<T> where T is a comparable data type with a natural order.
 	*	@return ArrayList<T> combined should contain all the elements of left and right in stricly non-decreasing order
 	*/
-	//TODO: Write the invariants for this shit.
-	//INVARIANTS (There should be around 3 invariants for this method)
[…]

### 2015-02-23 Programming Project B/paper.tex
added references
added the invariant outline
added first invariant

@@
 \subsection{Problems Encountered}
-No problems were encountered.
+One major issue encountered during the development of this insertion sort was that after completing the sort, A' was sorted properly except for the first element in the array.
+No matter what value the first element of A had, it did not change position in A'.
+For example, if A[5, 6, 3, 4, 7] was passed to the insertion sort algorithm, the returned array would look like A'[5, 3, 4, 6, 7].
+Changing the guard for the inner for loop (see Algorithm 1 line 6) from \textit{key < A[j] and j downto 1} to \textit{j downto 1 and key < A[j]} corrected this issue.
+
+Another issue encountered during the development of this program was executing Java $assert$ statements.
+When assertions were wrapped in parenthesis, they evaluated to a boolean, and the assert would throw an $AssertionError$ with the explanation 'cannot compare to boolean'.
+Removing the parenthesis wrapping the assertion fixed this issue.
+
+The last issue encountered during the development process of this program was generic programming.
+The program would not compile because not all Objects have a compareTo method.
+Adding the statement $<T$ $extends$ $Comparable<T>>$ fixed this issue.
+\cite{erik}
 %% END TESTING
@@
 %% BEGIN REFERENCES
- \section*{References}
+%%\section*{References}
 
-%%\nocite{*}
-%%\bibligraphystyle{IEEEtran}
-%%\bibliography{}
+\nocite{*}
+\bibliographystyle{IEEEtran}
+\bibliography{bib}
 

### 2015-02-23 Programming Project B/Code/Debug.java

@@ -6,6 +6,7 @@
 */
 
 import java.util.List;
+import java.util.ArrayList;
 
 public class Debug<T> {
 	boolean debugOn; //Variable to keep track of whether or not debug is on
@@ -67,6 +68,21 @@ public class Debug<T> {
 				assert actual.get(i - 1).compareTo(actual.get(i - 2)) >= 0;
 
 				i--;
+			}
+		}
+	}
+
+	/*
+	*	@Pre-Condition: actual and expected both contain Integer Objects
+	*	@Post-Condition: If all the elements inside of the actual arraylist are also contained in the expected arraylist, then the assertion holds true
+	*	@Description: Tests to ensure a given ArrayList of Integer Objects contains all the elements of another given ArrayList of Integer Objects
+	*	@param ArrayList<Integer> expected the list to check contains against
+	*	@param ArrayList<Integer> actual the list to check to make sure all its elements are contained in the other arraylist
+	*/
+	public void assertContains(ArrayList<Integer> expected, ArrayList<Integer> actual) {
+		if(debugOn == true) {
+			for(int i = 0; i < actual.size(); i++) {
+				assert expected.contains(actual.get(i));
 			}
 		}
 	}

### 2015-02-23 Programming Project B/Code/Driver.java

@@ -12,11 +12,11 @@ public class Driver{
     DummyData testData = new DummyData();
 
     ArrayList<Integer> testList = new ArrayList<Integer>();
-    testList = testData.runArrayList(10000000, 0, 10000000, testList);
+    testList = testData.runArrayList(10, 0, 100, testList);
     //testList = testData.runArrayList(1000, testList);
 
     System.out.println("Unsorted list: ");
-    //System.out.println(testList);
+    System.out.println(testList);
 
     Sort sorter = new Sort();
 

### 2015-02-23 Programming Project B/Code/Sort.java

@@ -50,7 +50,8 @@ public class Sort<T extends Comparable<T>> {
 	//Invariant for Second For Loop: TODO: write this
 	public ArrayList<T> heapSort(ArrayList<T> unsorted) {
 		int arrSize = unsorted.size() - 1;
-		for(int i = arrSize / 2; i >= 0; i--) { //Initialization: TODO: write this
+		//Initialization: TODO: write this
+		for(int i = arrSize / 2; i >= 0; i--) {
 			//Maintanance: TODO: write this
 			unsorted = heapify(unsorted, i, arrSize);
 			//Termination: TODO: write this
@@ -125,25 +126,42 @@ public class Sort<T extends Comparable<T>> {
 	*	@param a non-empty ArrayList<T> unsorted where T is a Comparable data type with a natural order
 	*	@return ArrayList<T> returnValue which is a permutation of unsorted, in strictly non-decreasing order,
 	*/
-	//Invariant for First For Loop: TODO: write this
+	//Invariant for First While Loop: left contains i elements, all of which can be found in sorted
 	//Invariant for Second For Loop: TODO: write this
 	public ArrayList<T> mergeSort(ArrayList<T> unsorted) {
 		ArrayList<T> sorted = unsorted;
 		ArrayList<T> left = new ArrayList<T>();
 		ArrayList<T> right = new ArrayList<T>();
 		ArrayList<T> returnValue;
+
+		//Debug
+		Debug debugger = new Debug();
+		debugger.turnOn();
+
 		if(sorted.size() <= 1) {
 			returnValue = sorted;
 		}
 		else {
 			int mid = (sorted.size() / 2);
-			for(int i = 0; i < mid; i++) { //Initialization: TODO: write this
-				//Maintanance: TODO: write this
+			int i = 0;
+			//Initialization: Our invariant holds true bcause i is zero and left contains 0 elements.
+			debugger.assertEquals(i, left.size());
+
+			while(i < mid) {
+				//Maintanance: Our invariant holds true because i is increased at the same rate elements are added to it from the same i index in sorted
+				debugger.assertEquals(i, left.size());
+				debugger.assertContains(sorted, left);
+
 				T temp = sorted.get(i);
 				left.add(temp);
-				//Termination: TODO: write this
+
+				i++;
 			}
-			for(int i = mid; i < sorted.size(); i++) { //Initialization: TODO: write this
[…]

### 2015-02-23 Programming Project B/Code/Sort.java
added second invariant

@@ -127,7 +127,7 @@ public class Sort<T extends Comparable<T>> {
 	*	@return ArrayList<T> returnValue which is a permutation of unsorted, in strictly non-decreasing order,
 	*/
 	//Invariant for First While Loop: left contains i elements, all of which can be found in sorted
-	//Invariant for Second For Loop: TODO: write this
+	//Invariant for Second For Loop: right contains y elements, all of which can be found in sorted
 	public ArrayList<T> mergeSort(ArrayList<T> unsorted) {
 		ArrayList<T> sorted = unsorted;
 		ArrayList<T> left = new ArrayList<T>();
@@ -144,11 +144,11 @@ public class Sort<T extends Comparable<T>> {
 		else {
 			int mid = (sorted.size() / 2);
 			int i = 0;
-			//Initialization: Our invariant holds true bcause i is zero and left contains 0 elements.
+			//Initialization: Our invariant holds true bcause i is zero and left contains 0 elements before the first iteration of the loop.
 			debugger.assertEquals(i, left.size());
 
 			while(i < mid) {
-				//Maintanance: Our invariant holds true because i is increased at the same rate elements are added to it from the same i index in sorted
+				//Maintanance: Our invariant holds true because i is increased at the same rate elements are added to left from the same i index in sorted
 				debugger.assertEquals(i, left.size());
 				debugger.assertContains(sorted, left);
 
@@ -161,12 +161,24 @@ public class Sort<T extends Comparable<T>> {
 			debugger.assertEquals(i, left.size());
 			debugger.assertContains(sorted, left);
 
-			for(int y = mid; y < sorted.size(); y++) { //Initialization: TODO: write this
-				//Maintanance: TODO: write this
-				T temp = sorted.get(i);
+			int y = mid;
+			//Initialization: Our invariant holds true because i is zero and right contains 0 elements before the first iteration of the loop.
+			debugger.assertEquals(y, right.size());
+
+			while(y < sorted.size()) {
+				//Maintanance: Our invariant holds true because y is increased at the same rate elements are added to right from the same y index in sorted.
+				debugger.assertEquals(y, right.size());
+				debugger.assertContains(sorted, right);
+
+				T temp = sorted.get(y);
 				right.add(temp);
-				//Termination: TODO: write this
+
+				y++;
 			}
+			//Termination: Our invariant holds true because y has been incremented at the same rate elements are added to left from the same index y in sorted
[…]

### 2015-02-23 Programming Project B/paper.tex
added first two loop invariants to the paper

@@
 \newline
+\textbf{Merge Sort First While Loop Invariant}: Left contains i elements, all of which can be found in A[0 ... i upto mid].
+\newline
+\textbf{Merge Sort First While Loop Invariant Initialization}: The invariant holds true vacuously because i is 0 and left contains no elements.
+\newline
+\textbf{Merge Sort First While Loop Invariant Maintenance}: The invariant holds true because i is incremented at the same rate that elements are added to left from the same i index value in A.
+\newline
+\textbf{Merge Sort First While Loop Invariant Termination}: The invariant holds true because i has been incremented at the same rate that elements have been added to left from the same i index value in A upto mid.
+\newline
+\newline
+\textbf{Merge Sort Second While Loop Invariant}: Right contains i elements, all of which can be found in A[mid ... A.length].
+\newline
+\textbf{Merge Sort Second While Loop Invariant Initialization}: The invariant holds true vacuously because i is 0 and right contains no elements.
+\newline
+\textbf{Merge Sort Second While Loop Invariant Maintenance}: The invariant holds true because i is incremented at the same rate that elements are added to right from the same i index value in A.
+\newline
+\textbf{Merge Sort Second While Loop Invariant Termination}: The invariant holds true because i has been incremented at the same rate that elements have been added to right from the same i index value in A upto to A.length.
+\newline
+
 %% TODO: Add pre and post conditions here

### 2015-02-23 Programming Project B/Code/Sort.java

@@ -127,7 +127,7 @@ public class Sort<T extends Comparable<T>> {
 	*	@return ArrayList<T> returnValue which is a permutation of unsorted, in strictly non-decreasing order,
 	*/
 	//Invariant for First While Loop: left contains i elements, all of which can be found in sorted
-	//Invariant for Second For Loop: right contains y elements, all of which can be found in sorted
+	//Invariant for Second While Loop: right contains y elements, all of which can be found in sorted
 	public ArrayList<T> mergeSort(ArrayList<T> unsorted) {
 		ArrayList<T> sorted = unsorted;
 		ArrayList<T> left = new ArrayList<T>();

### 2015-02-23 Programming Project B/Code/Debug.java
added first loop invariant for merge together

@@ -73,6 +73,23 @@ public class Debug<T> {
 	}
 
 	/*
+	*	@Pre-Condition: <code>ArrayList<Integer> actual</code> is an ArrayList of Integer Objects
+	*	@Post-Condition: If the ArrayList of Integer Objects is in stricly non-decreasing order, the program moves on normally, if not, the pgoram halts with an <code>AssertionError</code>
+	*	@Description: runs an assertion statement against an ArrayList of Integer Objects to ensure that for <code>k = actual.size(); A[k-2] <= A[k-1];</code>
+	*	@param ArrayList<Integer> actual the ArrayList to assert is in stricly non-decreasing order
+	*/
+	public void assertOrder(ArrayList<Integer> actual) {
+		if(debugOn == true) {
+			int i = actual.size();
+			while(i > 1) {
+				assert actual.get(i - 1).compareTo(actual.get(i - 2)) >= 0;
+
+				i--;
+			}
+		}
+	}
+
+	/*
 	*	@Pre-Condition: actual and expected both contain Integer Objects
 	*	@Post-Condition: If all the elements inside of the actual arraylist are also contained in the expected arraylist, then the assertion holds true
 	*	@Description: Tests to ensure a given ArrayList of Integer Objects contains all the elements of another given ArrayList of Integer Objects
@@ -83,6 +100,22 @@ public class Debug<T> {
 		if(debugOn == true) {
 			for(int i = 0; i < actual.size(); i++) {
 				assert expected.contains(actual.get(i));
+			}
+		}
+	}
+
+	/*
+	*	@Pre-Condition: expectedOne, expectedTwo, and actual all contain Integer Objects
+	*	@Post-Condition: If all the elemts inside of the actual ArrayList are also contined in either the expectedOne ArrayList or the expectedTwo ArrayList, then the assertion holds true
+	*	@Description: Tests to ensure a given ArrayList of Integer Objects contains all the elements of another given ArrayList of Integer Objects
+	*	@param: ArrayList<Integer> expectedOne one of the lists to check to see if the given ArrayList actual's elements are contained in
+	*	@param: ArrayList<Integer> expectedTwo one of the lists to check to see if the given ArrayList actual's elements are contained in
+	* @param: ArrayList<Integer> actual the list to check ot make sure all its elements are contained in either expectedOne or expectedTwo
+	*/
[…]

### 2015-02-23 Programming Project B/Code/Sort.java

@@ -77,7 +77,7 @@ public class Sort<T extends Comparable<T>> {
 	*	@param ArrayList<T> right a non-empty ArrayList<T> where T is a comparable data type with a natural order.
 	*	@return ArrayList<T> combined should contain all the elements of left and right in stricly non-decreasing order
 	*/
-	//Invariant for First While Loop: TODO: write this
+	//Invariant for First While Loop: combined contains x number of elements where x is the sum of i and y and those elements are contained in left[0 ... i] or right[0 ... y] in stricly non-decreasing order
 	//Invaraint for Second While Loop: TODO: write this
 	//Invariant for Third While Loop: TODO: write this
 	private ArrayList<T> mergeTogether(ArrayList<T> left, ArrayList<T> right) {
@@ -86,8 +86,22 @@ public class Sort<T extends Comparable<T>> {
 		int y = 0;
 		int x = 0;
 
-		while(left.size() != i && right.size() != y) { //Initialization: TODO: write this
-			//Maintanance: TODO: write this
+		//Debug
+		Debug debugger = new Debug();
+
+		//Initialization: Our invariant holds true vacuously before the first execution of the loop because x, i, and y are all equal to zero, combined is empty and therefore in order
+		debugger.assertEquals(0, i);
+		debugger.assertEquals(0, y);
+		debugger.assertEquals(0, x);
+		debugger.assertEquals(i, combined.size());
+
+		while(left.size() != i && right.size() != y) {
+			//Maintanance: Our invariant holds true at the beginning of each iteration of the loop because x is incremented whenever i or y is incremented and elements are added to combined from left and right in order
+			debugger.assertEquals(i+y, x);
+			debugger.assertEquals(x, combined.size());
+			debugger.assertOrder(combined);
+			debugger.assertContains(right, left, combined);
+
 			if(left.get(i).compareTo(right.get(y)) < 0) {
 				combined.add(x, left.get(i));
 				i++;
@@ -98,10 +112,15 @@ public class Sort<T extends Comparable<T>> {
 				y++;
 				x++;
 			}
-			//Termination: TODO: write this
 		}
+		//Termination: Our invariant holds tur at the termination of the loop because x has been incremented whenever i or y has been incremented, and elements are added to combined from left and right in order
+		debugger.assertEquals(i+y, x);
+		debugger.assertEquals(x, combined.size());
+		debugger.assertOrder(combined);
[…]

### 2015-02-23 Programming Project B/Code/Sort.java
added last invariant for merge together method

@@ -78,8 +78,8 @@ public class Sort<T extends Comparable<T>> {
 	*	@return ArrayList<T> combined should contain all the elements of left and right in stricly non-decreasing order
 	*/
 	//Invariant for First While Loop: combined contains x number of elements where x is the sum of i and y and those elements are contained in left[0 ... i] or right[0 ... y] in stricly non-decreasing order
-	//Invaraint for Second While Loop: TODO: write this
-	//Invariant for Third While Loop: TODO: write this
+	//Invaraint for Second While Loop: combined contains x number of elements where x is greater than or equal to i and those elements are contined in left[0 ... i] in stricly non-decreasing order
+	//Invariant for Third While Loop: combined contains x number of elements where x is greater than or equal to y and those elements are contained in right[0 ... y] in stricly non-decreasing order
 	private ArrayList<T> mergeTogether(ArrayList<T> left, ArrayList<T> right) {
 		ArrayList<T> combined = new ArrayList<T>();
 		int i = 0;
@@ -119,23 +119,52 @@ public class Sort<T extends Comparable<T>> {
 		debugger.assertOrder(combined);
 		debugger.assertContains(right, left, combined);
 
-		//Initialization: TODO: write this
+		//Initialization: Our invariant holds true before the first execution of the loop because x has been incremented whenever i has been incremented and elements have been added to combined from left in order
+		debugger.assertGreatEquals(x, i);
+		debugger.assertEquals(x, combined.size());
+		debugger.assertOrder(combined);
+		debugger.assertContains(left, combined);
+
 		while(left.size() != i) {
-			//Maintanance: TODO: write this
+			//Maintanance: Our invariant holds true at the beginning of each iteration of the loop because x has been incremented whenever i is incremented and elements have been added to combined from left in order
+			debugger.assertGreatEquals(x, i);
+			debugger.assertEquals(x, combined.size());
+			debugger.assertOrder(combined);
+			debugger.assertContains(left, combined);
+
 			combined.add(x, left.get(i));
 			i++;
 			x++;
-			//Termination: TODO: write this
 		}
+		//Termination: Our invariant holds true at the termination of the loop because x has been incremented whenever i was incremented and elements have been added to combined from left in order
+		debugger.assertGreatEquals(x, i);
+		debugger.assertEquals(x, combined.size());
+		debugger.assertOrder(combined);
+		debugger.assertContains(left, combined);
 
-		//Initialization: TODO: write this
[…]

### 2015-02-23 Programming Project B/paper.tex
added first merge together loop ivnariant to paper

@@
 \newline
+\newline
+\textbf{Merge Together First While Loop Invariant}: Combined contains x number of elements where x is the sum of i and y and those elements are contained in left[0 ...
+i] or right[0 ...
+y] in stricly non-decreasing order.
+\newline
+\textbf{Merge Together First While Loop Invariant Initialization}: Our invariant holds true vacuously before the first iteration of the loop because $i = 0$, $y = 0$, $i + y = 0$, $x = 0$ therefore $x = i + y$ and combined contains no elements and is vacuously in order.
+\newline
+\textbf{Merge Together First While Loop Invariant Maintenance}: Our invariant holds true at the beginning of each iteration of the loop because $x$ is incremented whenever $i$ and $y$ are incremented, and elements are added to combined from left only when $left[i] < right[y]$ and added from right only when $left[i] >= right[y]$.
+\newline
+\textbf{Merge Together First While Loop Invariant Termination}: Our invariant holds true when the loop terminates because $x$ has been incremented whenever $i$ and $y$ are incremented, and elements have only been added to combined from left when $left[i] < right[y]$ and from right only when $left[i] >= right[y]$.
+\newline
+\newline
+\textbf{Merge Together Second While Loop Invariant}:
 

### 2015-02-23 Programming Project B/paper.tex
added available invarariants to the paper up to a point

@@
 \newline
+\textbf{Merge Sort Conclusion}: %%TODO: Write this
+\newline
+\newline
 \textbf{Merge Together First While Loop Invariant}: Combined contains x number of elements where x is the sum of i and y and those elements are contained in left[0 ...
@@
 \newline
-\textbf{Merge Together Second While Loop Invariant}:
+\textbf{Merge Together Second While Loop Invariant}: Combined contains x number of elements where $x >= i$ and those elements are contained in left[0 ... i] in stricly non-decreasing order.
+\newline
+\textbf{Merge Together Second While Loop Invariant Initialization}: The invariant holds true before the first iteration of the loop because x has been incremented when i has been incremented and elements have been added to combined from left if and only if $left[i] < right[y]$ and from right if and only if $left[i] >= right[y]$.
+\newline
+\textbf{Merge Together Second While Loop Invariant Maintenance}: The invariant holds true at the beginning of each iteration of the loop because x has been incremented when i has been incremented and elements have been added to combined from left in order.
+\newline
+\textbf{Merge Together Second While Loop Invariant Termination}: The invariant holds true at the termination of the loop because x has been incremented when i has been incremented and elements have been added to combined from left in order.
+\newline
+\newline
+\textbf{Merge Together Third While Loop Invariant}: Combined contains x number of elements where x is greather than or equal to y and those elements are contained in right[0 ...
+y] in stricly non-decreasing order.
+\newline
+\textbf{Merge Together Third While Loop Invariant Initialization}: The invariant holds true before the first initialization of the loop because x has been incremented when y has been incremented and elements have been added to combined from left if and only if $left[i] < right[y]$ and from right if and only if $left[i] >= right[y]$.
+\newline
+\textbf{Merge Together Third While Loop Invariant Maintenance}: The invariant holds true at the beginning of each iteration of the loop because x has been incremented when y has been incremented and elements have been added to combined from right in order.
+\newline
+\textbf{Merge Together Third While Loop Invariant Termination}: The invariant holds true at the termination of the loop because x has been incremented when y has been incremented and elements have been added to combined from right in order.
+\newline
+\newline
+\textbf{Merge Together Conclusion}: %%TODO: Write this.
+\newline
+\newline
 
-%% TODO: Add pre and post conditions here
 A heap sort can be implemented in a variety of languages using the pseudocode below in Algorithm 3.
+\newline
+\textbf{Heap Sort Pre-Condition}: A is a non-empty array with a natural order.
+\newline
+\textbf{Heap Sort Post-Condition}: A' is a permutation of A in strictly non-decreasing order.
+\newline
+\textbf{Heapify Pre-Condition}: A is a non-empty array with a comparable data type in a natural order.
+\newline
+\textbf{Heapify Post-Condition}: %%TODO: Write this.
 \newline

### 2015-02-23 Programming Project B/Code/Sort.java

@@ -11,7 +11,7 @@ import java.util.List;
 public class Sort<T extends Comparable<T>> {
 	/*
 	*	@Pre-Condition: ArrayList<T> is a non-empty set of data where T is a comparable data type with a natural order
-	*	@Post-Condition:
+	*	@Post-Condition: TODO: Write this
 	*	@Description: heapify is a helper method for heapSort that keeps the heap in order so that the root node is the largest element in the heap.
 	*	@param ArrayList<T> unsorted is a non-empty set of data where T is a comparable data type with a natural order
 	*	@param int i

### 2015-02-23 Programming Project B/paper.tex
copy and pasted shit from old paper

@@
 \section{Experimental Analysis}
+The insertion sort demonstrated in Algorithm 1, 2, and 3 was implemented in Java and executed on an HP SpectreXT TouchSmart with a 4 core Intel i7 processor clocked at 1.9GHz running Ubuntu Gnome 14.10 64-bit.
+%% TODO: Include Graphs here
 
+The expected growth of insertion sort as the number of elements (n) grows large can be represented as $ \theta(n^2) $ where $ \theta() $ represents the asymptotically tightly bound running time.
+\cite{textbook} At $n_0$ the algorithm took 0.0003 milliseconds to complete.
+At $n_1$ the algorithm also took 0.0003 milliseconds to complete.
+For both these values of $n$, the algorithm runs at $ \theta(1) $, which is a constant value because the algorithm completes before a loop is run.
+As $n$ grows larger the time to complete the experimental data correlates quite accurately with the expected growth.
+For $n_{1*10^6} $ the data matches up perfectly.
+This is to be expected, because as n grows larger the constants and lower orders of the actual running time of the insertion sort start to affect the running time less and less as the highest order of $n^2$ is so large.
+It is expected for lower values of n to not correlate well with $ \theta(n^2) $ because the system running the algorithm may be executing superfolous commands such as checking for system.
+For the data graphed above (see Fig.
+1) this was in fact the case.
+To aid this issue, averages of 10 trails were taken and used in Figure 1.
+
+%%TODO: Add shit about merge sort here
+%%TODO: Add shit about heap sort here
 %% END EXPERIMENTAL ANALYSIS

### 2015-02-23 Programming Project B/paper.tex
added a graph and other shit too

@@
 \newline
-\textbf{Insertion Sort Outer-Loop Invariant}: The subarray A'[1 ... i - 1] contains all the same elements as the subarray A[1 .. i - 1].
+\textbf{Insertion Sort Outer-Loop Invariant}: The subarray A'[1 ... i - 1] contains all the same elements as the subarray A[1 ... i - 1].
 \newline 
@@
 \newline
-\textbf{Merge Sort First While Loop Invariant}: Left contains i elements, all of which can be found in A[0 ... i upto mid].
+\textbf{Merge Sort First While Loop Invariant}: Left contains i elements, all of which can be found in A[1 ... i upto mid].
 \newline
@@
 \newline
-\textbf{Merge Together First While Loop Invariant}: Combined contains x number of elements where x is the sum of i and y and those elements are contained in left[0 ...
-i] or right[0 ...
+\textbf{Merge Together First While Loop Invariant}: Combined contains x number of elements where x is the sum of i and y and those elements are contained in left[1 ...
+i] or right[1 ...
 y] in stricly non-decreasing order.
@@
 \newline
-\textbf{Merge Together Second While Loop Invariant}: Combined contains x number of elements where $x >= i$ and those elements are contained in left[0 ... i] in stricly non-decreasing order.
+\textbf{Merge Together Second While Loop Invariant}: Combined contains x number of elements where $x >= i$ and those elements are contained in left[1 ... i] in stricly non-decreasing order.
 \newline
@@
 \newline
-\textbf{Merge Together Third While Loop Invariant}: Combined contains x number of elements where x is greather than or equal to y and those elements are contained in right[0 ...
+\textbf{Merge Together Third While Loop Invariant}: Combined contains x number of elements where x is greather than or equal to y and those elements are contained in right[1 ...
 y] in stricly non-decreasing order.
@@
 The insertion sort demonstrated in Algorithm 1, 2, and 3 was implemented in Java and executed on an HP SpectreXT TouchSmart with a 4 core Intel i7 processor clocked at 1.9GHz running Ubuntu Gnome 14.10 64-bit.
-%% TODO: Include Graphs here
+ \begin{figure}[!]
+\begin{center}
+\includegraphics[scale=.70]{test-results.pdf}
+\end{center}
+\captionof{figure}{Merge, Heap and Insertion Sort Run-Time Comparison}
+\label{fig:comparisonfig}
+\end{figure}
 
@@
 
-%%TODO: Add shit about merge sort here
-%%TODO: Add shit about heap sort here
+The expected growth of merge sort as the number of elements (n) grows large can be represented as $ \theta(n*\lg(n)) $.
+At $n_0$ the sort took 0.0002 milliseconds to complete, however, at $n_1$ the algorithm took 0.0001 milliseconds to complete.
+The discrepency can be attributed to superfolous commands being executed by the system.
+As $n$ grows larger the time to complete correlates with the expected running time.
+It is expected for lower values of $n$ to not correlate as nicely with the expected running time because the system constant will have a larger impact on the smaller values of $n$.
+It is important to note that merge sort ran much faster than insertion sort at large values of n such as $n_{1*10^6} $ and was able to process $n_{1*10^7} $ where insertion sort crashed the system.
+The best case running time for this implementation of merge sort is $ \theta(1) $ and is elicited only when $ n <= 1 $ because if $n <= 1$ the array is vacuously sorted and can be returned immediately without executing any loop, and run at a constant time.
+
+Heap sort is very similar to merge sort in that its expected running time is $ \theta(n*\lg(n)) $.
+At lower values of $n$ heap sort ran marginally quicker than merge sort, however, as $n$ grew larger merge sort began to run marginally faster than heap sort.
+This difference is neglagable because of superfolous commands being executed by the system and other running processes.
+For all intensive purposes merge sort and heap sort have the same run time and the major difference between the two is the system constant.
+For $n_0$ heap sort completed in 0.0002 milliseconds and for $n_1$ heap sort completed in 0.0003 milliseconds.
+In both merge sort and insertion sort the sort completes before any loop is executed if $n <= 1$ however in this implementation of heap sort the algorithm runs through the entire sort.
+We can tell from this that the constant time is 0.0002 milliseconds, which correlates nicely with both merge and insertion sort.
+$ \theta(n*\lg(n)) $ is the both the best and worst case run-time for heap sort.
 %% END EXPERIMENTAL ANALYSIS
@@
 \section{Conclusions}
-
+With a running time of $ \theta(n*\lg(n)) $ both heap sort and merge sort perform much faster than insertion sort at a running time of $ \theta(n^2) $, especially when $n$ grows large.
+While faster comptuers will be able to execute insertion sort faster, as $n$ grows large, even on a slower machine, both merge sort and heap sort will perform faster.
+Because it is easier to setup and maintain, insertion sort is a good choice when dealing with small sets of data that don't go beyond $n_{1*10^5}$, however, if data is expected to grow larger than that, either heap or merge sort is the best choice.
 %% END CONCLUSION

### 2015-02-23 Programming Project B/Code/Debug.java
added first loop invariant for heap sort to code

@@ -90,6 +90,19 @@ public class Debug<T> {
 	}
 
 	/*
+	*	@Pre-Condition: ArrayList is an ArrayList of Integers and i is less than or equal to half of the size of actual
+	*	@Post-Condition: If elements exist past i the assertion holds
+	*	@Description: runs an assertion statement against an ArrayList of Integer Objects to ensure that there are children nodes of actual[i].
+	*	@param ArrayList<Integer> actual the array to test against
+	*	@param int i the index value to check has children nodes.
+	*/
+	public void assertChildren(ArrayList<Integer> actual, int i) {
+		if(debugOn == true) {
+			assert actual.size() > i;
+		}
+	}
+
+	/*
 	*	@Pre-Condition: actual and expected both contain Integer Objects
 	*	@Post-Condition: If all the elements inside of the actual arraylist are also contained in the expected arraylist, then the assertion holds true
 	*	@Description: Tests to ensure a given ArrayList of Integer Objects contains all the elements of another given ArrayList of Integer Objects

### 2015-02-23 Programming Project B/Code/Sort.java

@@ -11,8 +11,8 @@ import java.util.List;
 public class Sort<T extends Comparable<T>> {
 	/*
 	*	@Pre-Condition: ArrayList<T> is a non-empty set of data where T is a comparable data type with a natural order
-	*	@Post-Condition: TODO: Write this
-	*	@Description: heapify is a helper method for heapSort that keeps the heap in order so that the root node is the largest element in the heap.
+	*	@Post-Condition: Each parent node is more extreme than its child node.
+	*	@Description: heapify is a helper method for heapSort that keeps the heap in order so that the root node is the most extreme element in the heap.
 	*	@param ArrayList<T> unsorted is a non-empty set of data where T is a comparable data type with a natural order
 	*	@param int i
 	*	@param int total
@@ -46,25 +46,41 @@ public class Sort<T extends Comparable<T>> {
 	*	@param ArrayList<T> unsorted is a non-empty ArrayList<T> where T is a comparable data type with a natural order
 	*	@return ArrayList<T> sorted is a permutation of unsorted in strictly non-decreasing order
 	*/
-	//Invariant for First For Loop: TODO: write this
-	//Invariant for Second For Loop: TODO: write this
+	//Invariant for First While Loop: unsorted[i] is the parent element of a heap
+	//Invariant for Second While Loop: TODO: write this
 	public ArrayList<T> heapSort(ArrayList<T> unsorted) {
+		//Debug
+		Debug debugger = new Debug();
+
 		int arrSize = unsorted.size() - 1;
-		//Initialization: TODO: write this
-		for(int i = arrSize / 2; i >= 0; i--) {
-			//Maintanance: TODO: write this
+		int i = arrSize / 2;
+		//Initialization: Our invariant holds true before the first iteration of the loop because unsorted[i] must have child elements
+		debugger.assertChildren(unsorted, i);
+		while(i >= 0) {
+			//Maintanance: Our invariant holds true at the beginning of each iteration of the loop because unsorted[i] must have children elements
+			debugger.assertChildren(unsorted, i);
+
 			unsorted = heapify(unsorted, i, arrSize);
-			//Termination: TODO: write this
+
+			i--;
 		}
-		for(int i = arrSize; i > 0; i--) { //Initialization: TODO: write this
[…]

### 2015-02-23 Programming Project B/Code/Sort.java
added second loop invariant for heap sort in the code

@@ -46,8 +46,8 @@ public class Sort<T extends Comparable<T>> {
 	*	@param ArrayList<T> unsorted is a non-empty ArrayList<T> where T is a comparable data type with a natural order
 	*	@return ArrayList<T> sorted is a permutation of unsorted in strictly non-decreasing order
 	*/
-	//Invariant for First While Loop: unsorted[i] is the parent element of a heap
-	//Invariant for Second While Loop: TODO: write this
+	//Invariant for First While Loop: unsorted[i] is the parent element in a heap
+	//Invariant for Second While Loop: All elements in unsorted greater than the index value of y are in stricly non-decreasing order
 	public ArrayList<T> heapSort(ArrayList<T> unsorted) {
 		//Debug
 		Debug debugger = new Debug();
@@ -68,9 +68,12 @@ public class Sort<T extends Comparable<T>> {
 		debugger.assertChildren(unsorted, i);
 
 		int y = arrSize;
-		//Initialization: TODO: write this
+		//Initialization: Our invariant holds vacuously true before the first iteration of the loop because there are no elements in unsorted that are at an index value greater than y.
+		debugger.assertStrictLess(arrSize, y+1);
+
 		while(y > 0) {
-			//Maintanance: TODO: write this
+			//Maintanance: Our invariant holds true at the beginning of each iteration of the loop because all elements greater than y are in strictly non-decreasing order
+
 			T tmp = unsorted.get(0);
 			unsorted.set(0, unsorted.get(y));
 			unsorted.set(y, tmp);
@@ -79,7 +82,8 @@ public class Sort<T extends Comparable<T>> {
 
 			y--;
 		}
-		//Termination: TODO: write this
+		//Termination: Our invariant holds true at the termination of the loop because y decreases as each largest element is moved to the end of the list until the entire array has been traversed, so that all elements greater than y are in stricly non-decreasing order
+		debugger.assertOrder(unsorted);
 
 		ArrayList<T> sorted = unsorted;
 		return sorted;

### 2015-02-23 Programming Project B/paper.tex
added last invariants to the paper

@@
 \newline
-\textbf{Merge Sort First While Loop Invariant}: Left contains i elements, all of which can be found in A[1 ... i upto mid].
+\textbf{Merge Sort First For Loop Invariant}: Left contains i elements, all of which can be found in A[1 ... i upto mid].
 \newline
-\textbf{Merge Sort First While Loop Invariant Initialization}: The invariant holds true vacuously because i is 0 and left contains no elements.
+\textbf{Merge Sort First For Loop Invariant Initialization}: The invariant holds true vacuously because i is 0 and left contains no elements.
 \newline
-\textbf{Merge Sort First While Loop Invariant Maintenance}: The invariant holds true because i is incremented at the same rate that elements are added to left from the same i index value in A.
+\textbf{Merge Sort First For Loop Invariant Maintenance}: The invariant holds true because i is incremented at the same rate that elements are added to left from the same i index value in A.
 \newline
-\textbf{Merge Sort First While Loop Invariant Termination}: The invariant holds true because i has been incremented at the same rate that elements have been added to left from the same i index value in A upto mid.
+\textbf{Merge Sort First For Loop Invariant Termination}: The invariant holds true because i has been incremented at the same rate that elements have been added to left from the same i index value in A upto mid.
 \newline
 \newline
-\textbf{Merge Sort Second While Loop Invariant}: Right contains i elements, all of which can be found in A[mid ... A.length].
+\textbf{Merge Sort Second For Loop Invariant}: Right contains i elements, all of which can be found in A[mid ... A.length].
 \newline
-\textbf{Merge Sort Second While Loop Invariant Initialization}: The invariant holds true vacuously because i is 0 and right contains no elements.
+\textbf{Merge Sort Second For Loop Invariant Initialization}: The invariant holds true vacuously because i is 0 and right contains no elements.
 \newline
-\textbf{Merge Sort Second While Loop Invariant Maintenance}: The invariant holds true because i is incremented at the same rate that elements are added to right from the same i index value in A.
+\textbf{Merge Sort Second For Loop Invariant Maintenance}: The invariant holds true because i is incremented at the same rate that elements are added to right from the same i index value in A.
 \newline
-\textbf{Merge Sort Second While Loop Invariant Termination}: The invariant holds true because i has been incremented at the same rate that elements have been added to right from the same i index value in A upto to A.length.
+\textbf{Merge Sort Second For Loop Invariant Termination}: The invariant holds true because i has been incremented at the same rate that elements have been added to right from the same i index value in A upto to A.length.
 \newline
@@
 \end{algorithm}
+\newline
+\textbf{Heap Sort First Loop Invariant}: A[i] is the parent element in a heap.
+\newline
+\textbf{Heap Sort First For Loop Invariant Initialization}: The invariant holds true before the first iteration of the loop because i is less than the size of the array, and therefore has a child node.
+\newline
+\textbf{Heap Sort First For Loop Invariant Maintenance}: The invariant holds true at the beginning of each iteration of the loop because A[i] will always be smaller than the size of the array, and therefore have children nodes.
+\newline
+\textbf{Heap Sort First For Loop Invariant Termination}: The invariant holds true at the termination of the loop because i will be the smallest index value of A and therefore will have children nodes.
+\newline
+\newline
+%% TODO: something about a conclusion
+\textbf{Heap Sort Second For Loop Invariant}: All elements in A at the index value greater than i are in stricly non-decreasing order.
+\newline
+\textbf{Heap Sort Second For Loop Invariant Initialization}: The invariant holds vacuously true before the first iteration of the loop because there are no elements in A that are at an index value greater than i.
+\newline
+\textbf{Heap Sort Second For Loop Invariant Maintenance}: The invariant holds true at the beginning of each iteration of the loop because each most extreme element is moved to the very end of the array at an index value greater than i.
+\newline
+\textbf{Heap Sort Second For Loop Invariant Termination}: The invariant holds true at the termination of the loop because i decreases as ecah largest element is moved to the end of A until the entire array has been traversed, so that all elements greater than i are in stirctly non-decreasing order.
+\newline
+\newline
+%% TODO: Something about a conclusion here or some shit like that
 %% END PROCEDURE

### 2015-02-23 Programming Project B/paper.tex
'finished ' paper

@@
 \begin{abstract}
-
+The present study investiages the insertion sort, merge sort, and heap sort sorting algorithms and how they perform on data sets of different sizes and types whe implemented in Java.
+Insertion sort is a sorting algorithm that runs in $ \theta(n^2) $.
+For sorting large sets of data, insertion sort does not perform exemplary, however, for simple sorting of data sets under 10,000, insertion sort performs resonably quickly on modern day computers.
+Merge sort and heap sort both perform at $ \theta(n*\lg(n)) $.
+Both merge and heap sort perform exceptionally well when sorting both small and large sets of data, even upto 10,000,000 elements.
 %% END ABSTRACT
@@
 \newline
-\textbf{Merge Sort Conclusion}: %%TODO: Write this
-\newline
-\newline
 \textbf{Merge Together First While Loop Invariant}: Combined contains x number of elements where x is the sum of i and y and those elements are contained in left[1 ...
@@
 \newline
-\textbf{Merge Together Conclusion}: %%TODO: Write this.
-\newline
-\newline
 
@@
 \newline
-\textbf{Heapify Post-Condition}: %%TODO: Write this.
+\textbf{Heapify Post-Condition}: Each parent node is more extreme than its child  node.
 \newline
@@
 \newline
-%% TODO: something about a conclusion
 \textbf{Heap Sort Second For Loop Invariant}: All elements in A at the index value greater than i are in stricly non-decreasing order.
@@
 \newline
-%% TODO: Something about a conclusion here or some shit like that
 %% END PROCEDURE
@@
 All arrays used in testing are Java ArrayList<Integer> unless otherwise spcified.
-All times are recorded in milliseconds using a stopwatch class borrowed from Robert Sedgewock and Kevin Wayne %%TODO: Cite this shit.
+All times are recorded in milliseconds using a stopwatch class borrowed from Robert Sedgewock and Kevin Wayne \cite{stopwatch}
 It is important to note that the soptwatch class used takes the elapsed real-time between the start of the sort algorithm and the end of the sort algorithim as opposed to taking the elapsed processor-time becasue these tests were run on a multi-core computer.

### 2015-02-23 Programming Project B/paper.tex
final paper turned in

@@
 \begin{abstract}
-The present study investiages the insertion sort, merge sort, and heap sort sorting algorithms and how they perform on data sets of different sizes and types whe implemented in Java.
+The present study investigates the insertion sort, merge sort, and heap sort sorting algorithms and how they perform on data sets of different sizes and types when implemented in Java.
 Insertion sort is a sorting algorithm that runs in $ \theta(n^2) $.
-For sorting large sets of data, insertion sort does not perform exemplary, however, for simple sorting of data sets under 10,000, insertion sort performs resonably quickly on modern day computers.
+For sorting large sets of data, insertion sort does not perform exemplary, however, for simple sorting of data sets under 10,000, insertion sort performs reasonably quickly on modern day computers.
 Merge sort and heap sort both perform at $ \theta(n*\lg(n)) $.
@@
 One such sorting algorithm is insertion sort, which sorts by iterating through a list of data, taking the current position, and repositioning it into a more appropriate place in the list.
-Heap sort is another sorting algorithm that greatly differs than insertion sort in that it uses a divide, conquer, and combine method; meaning that it breaks the set it is sorting into subsets until the subsets can no longer be broken up and then merge sort combines the subsets together rendering the correct answer.
-%% Write this for heap sort.
+Merge sort is another sorting algorithm that greatly differs than insertion sort in that it uses a divide, conquer, and combine method; meaning that it breaks the set it is sorting into subsets until the subsets can no longer be broken up and then merge sort combines the subsets together rendering the correct answer.
+Heap sort creates a heap data structure and then swaps the first index with the last index value and re-heaps the data structure until the entire set is sorted.
 %% END BACKGROUND
@@
 \newline
-\textbf{Merge Sort Pre-Condition}: A is a non-empty array of a comparable data with a natural order.
+\textbf{Merge Sort Pre-Condition}: A is a non-empty array of a comparable data type with a natural order.
 \newline
@@
 \newline
-\textbf{Merge Together Third While Loop Invariant}: Combined contains x number of elements where x is greather than or equal to y and those elements are contained in right[1 ...
-y] in stricly non-decreasing order.
+\textbf{Merge Together Third While Loop Invariant}: Combined contains x number of elements where x is greater than or equal to y and those elements are contained in right[1 ...
+y] in strictly non-decreasing order.
 \newline
@@
 \newline
-\textbf{Heap Sort Second For Loop Invariant Termination}: The invariant holds true at the termination of the loop because i decreases as ecah largest element is moved to the end of A until the entire array has been traversed, so that all elements greater than i are in stirctly non-decreasing order.
+\textbf{Heap Sort Second For Loop Invariant Termination}: The invariant holds true at the termination of the loop because i decreases as each largest element is moved to the end of A until the entire array has been traversed, so that all elements greater than i are in stirctly non-decreasing order.
 \newline
@@
 \subsection{Testing Plan and Results}
-All arrays used in testing are Java ArrayList<Integer> unless otherwise spcified.
-All times are recorded in milliseconds using a stopwatch class borrowed from Robert Sedgewock and Kevin Wayne \cite{stopwatch}
-It is important to note that the soptwatch class used takes the elapsed real-time between the start of the sort algorithm and the end of the sort algorithim as opposed to taking the elapsed processor-time becasue these tests were run on a multi-core computer.
-In the table below, A denotes Array. Times in the table below are given as averages out of 10 trials.
+All arrays used in testing are Java ArrayList<Integer> unless otherwise specified.
+All times are recorded in milliseconds using a stopwatch class borrowed from Robert Sedgwick and Kevin Wayne \cite{stopwatch}
+It is important to note that the stopwatch class used takes the elapsed real-time between the start of the sort algorithm and the end of the sort algorithm as opposed to taking the elapsed processor-time because these tests were run on a multi-core computer.
+In the table below, A denotes Array. Times in the table below are given as averages out of 10 trials in milliseconds.
 \newline

### 2015-02-25 Mini Project A/Driver.java
lcs project

@@ -1,0 +1,12 @@
+public class Driver {
+	public static void main(String args[]) {
+		String first = "ABCDAC";
+		String second = "AADOKC";
+
+		LCS lcs = new LCS();
+		String result = lcs.runLCS(first, second);
+
+		System.out.println("Subsequence: "+result);
+		System.out.println("Length: "+result.length());
+	}
+}

### 2015-02-25 Mini Project A/LCS.java

@@ -1,0 +1,44 @@
+/*
+*	@Author: Preston Stosur-Bassett
+*	@Description: LCS will find the longest common substring of two strings. 
+*	@Date: 25.2.15
+*	@Class: LCS
+*/
+
+public class LCS {
+	int length;
+	String sequence;
+
+	public String runLCS(String one, String two) {
+		if(one.length() > 0 && two.length() > 0) {
+			String oneSub = one.substring(0, one.length()-1);
+			String twoSub = two.substring(0, two.length()-1);
+
+			if(one.length() == 0 || two.length() == 0) {
+				length = 0;
+				sequence = "";
+				return sequence;
+			}
+			else if(one.charAt(one.length()-1) == two.charAt(two.length()-1)) {
+				sequence = runLCS(oneSub, twoSub) + one.charAt(one.length()-1);
+				return sequence;
+			}
+			else {
+				String a = runLCS(one, twoSub);
+				String b = runLCS(two, oneSub);
+				
+				if(a.length() > b.length()) {
+					 sequence = a;
+					 return sequence;
+				}			
+				else {
+					sequence = b;
+					return sequence;
+				}
+			}
+		}
+		else {
+			return sequence;
+		}
+	}
+}

### 2015-02-25 Mini Project A/lcs.js

@@ -1,0 +1,24 @@
+/*
+*	@Author: Preston Stosur-Bassett
+*	@Date: 25.2.15
+*	@Language: Javascript
+*	@Description: The javascript function will find the longest common substring of two strings.
+*/
+
+function lcs(one, two) {
+	var oneSub = one.substr(0, one.length-1);
+	var twoSub = two.substr(0, two.length-1);
+
+	if(one.length == 0 || two.length == 0) {
+		return "";
+	}
+	else if(one.charAt(one.length-1) == two.charAt(two.length-1)) {
+		return lcs(oneSub, twoSub) + one.charAt(one.length - 1);
+	}
+	else {
+		var a = lcs(one, twoSub);
+		var b = lcs(oneSub, two);
+		return (a.length > b.length) ? a : b;
+	}
+}
+

### 2015-02-25 Mini Project A/lcs.js
added a more understandable if statement

@@ -18,7 +18,12 @@ function lcs(one, two) {
 	else {
 		var a = lcs(one, twoSub);
 		var b = lcs(oneSub, two);
-		return (a.length > b.length) ? a : b;
+		if(a.length > b.length) {
+			return a;
+		}
+		else {
+			return b;
+		}
 	}
 }
 

### 2015-03-10 Programming Project C/Code/DummyData.java
fixed greedy algorithm, and may i say i did a great job, it runs really well

@@ -29,10 +29,11 @@ public class DummyData {
 			// MAINTANANCE: At the beginning of each iteration, one element was added to A and start was increased by one, therefore, our invariant holds true.
 			debugger.assertGreatEquals(list.size(), start);
 			Integer intToAdd = new Integer(random.nextInt((max - min + 1) + min));
-			list.add(intToAdd);
+			if(intToAdd != 0) {
+				list.add(intToAdd);
 
-			//Count up on the iterator
-			start++;
+				start++;
+			} 
 		}
 		/*TERMINATION: The negation of the guard implies that (end - start) number of elements have been added to A, since start is initialized as 0 at the beginning of the method and is
 			incremented by 1 each iteration of the loop, which means that start amount of elements have been added to A, and so our invariant holds true.	*/

### 2015-03-10 Programming Project C/Code/Knapsack.java

@@ -31,36 +31,38 @@ public class Knapsack<T extends Comparable<T>> {
       returnValue = 0;
     }
     else {
-      ArrayList<Integer> tab = new ArrayList<Integer>(elems);
-      int sizePrime = backpackSize;
       int profit = 0;
+
+      ArrayList<ArrayList<Integer>> ratioListings = new ArrayList<ArrayList<Integer>>(elems);
 
       //Initialization: TODO: write this
       for(int i = 0; i < elems; i++) {
         //Maintenance: TODO: write this
-        tab.add(new Integer(0));
+        int ratio = prices.get(i).intValue() / weights.get(i).intValue();
+        ArrayList<Integer> innerRatioListing = new ArrayList<Integer>(3);
+        innerRatioListing.add(new Integer(ratio));
+        innerRatioListing.add(new Integer(prices.get(i).intValue()));
+        innerRatioListing.add(new Integer(weights.get(i).intValue()));
+        ratioListings.add(innerRatioListing);
       }
       //Termination: TODO: write this
 
-      //Initialization: TODO: write this
+      Sort sorter = new Sort();
+
+      ratioListings = sorter.insertionSortNestedArray(ratioListings, 0);
+
       int i = 0;
-      while(i < elems - 1) {
+      //Initialization: TODO: write this
+      while(backpackSize > 0 && i < elems) {
         //Maintenance: TODO: write this
-        if(!(weights.get(i).intValue() > sizePrime)) {
-          tab.set(i, new Integer(1));
-          profit = profit + prices.get(i).intValue();
-          sizePrime = sizePrime - weights.get(i).intValue();
+        if(backpackSize - ratioListings.get(i).get(2).intValue() > 0) {
+          profit = profit + ratioListings.get(i).get(1).intValue();
+          backpackSize = backpackSize - ratioListings.get(i).get(2).intValue();
         }
 
         i++;
       }
       //Termination: TODO: write this
-
-      if(i < elems) {
-        tab.set(i, new Integer(sizePrime / weights.get(i).intValue()));
-      }
-
-      profit = profit + (tab.get(i).intValue() * prices.get(i).intValue());
 
       returnValue = profit;
     }
@@ -108,8 +110,12 @@ public class Knapsack<T extends Comparable<T>> {
       }
       // First Outer While Loop Termination: TODO: write this
 
+      // Second Outer For Loop Initialization: TODO: write this
       for(int i = 0; i < elems - 1; i++) {
[…]

### 2015-03-10 Programming Project C/Code/Sort.java

@@ -315,4 +315,42 @@ public class Sort<T extends Comparable<T>> {
 		}
 		return sorted;
 	}
+
+
+	/*
+	*	@Pre-Condition: TODO: write this
+	*	@Post-Condition: TODO: write this
+	*	@Description: TODO: write this
+	* @param ArrayList<ArrayList<T>> unsorted TODO: write this
+	*	@param int sortingIndex TODO: write this
+	*	@return ArrayList<ArrayList<T>> sorted TODO: write this
+	*/
+	//There should be two loop invariants, I should be able to take them straight from above
+	public ArrayList<ArrayList<T>> insertionSortNestedArray(ArrayList<ArrayList<T>> list, int sortingIndex) {
+		if(list.size() > 1) {
+			int i = 1;
+
+			//Outer While Loop Initialization: TODO: write this
+			while(i < list.size()) {
+				//Outer While Loop Maintenance: TODO: write this
+				ArrayList<T> currentElement = list.get(i);
+				T value = list.get(i).get(sortingIndex);
+				int j = i - 1;
+
+				//Inner While Loop Initialization: TODO: write this
+				while(j >= 0 && (value.compareTo(list.get(j).get(sortingIndex)) > 0)) {
+					//Inner While Loop Maintenance: TODO: write this
+					list.set(j+1, list.get(j));
+					j--;
+				}
+				//Inner While Loop Termination: TODO: write this
+				list.set(j+1, currentElement);
+
+				i++;
+			}
+			//Outer While Loop Termination: TODO: write this
+		}
+
+		return list;
+	}
 }

### 2015-03-10 Programming Project C/Code/Knapsack.java
had some testing out the code, and oh yeah, greedy is much, much better

@@ -55,7 +55,7 @@ public class Knapsack<T extends Comparable<T>> {
       //Initialization: TODO: write this
       while(backpackSize > 0 && i < elems) {
         //Maintenance: TODO: write this
-        if(backpackSize - ratioListings.get(i).get(2).intValue() > 0) {
+        if(backpackSize - ratioListings.get(i).get(2).intValue() >= 0) {
           profit = profit + ratioListings.get(i).get(1).intValue();
           backpackSize = backpackSize - ratioListings.get(i).get(2).intValue();
         }

### 2015-03-10 Programming Project C/test.cpp
testing

@@ -1,74 +1,0 @@
-# include<stdio.h>
-
-void knapsack(int n, float weight[], float profit[], float capacity) {
-   float x[20], tp = 0;
-   int i, j, u;
-   u = capacity;
-
-   for (i = 0; i < n; i++)
-      x[i] = 0.0;
-
-   for (i = 0; i < n; i++) {
-      if (weight[i] > u)
-         break;
-      else {
-         x[i] = 1.0;
-         tp = tp + profit[i];
-         u = u - weight[i];
-      }
-   }
-
-   if (i < n)
-      x[i] = u / weight[i];
-
-   tp = tp + (x[i] * profit[i]);
-
-   printf("\nThe result vector is:- ");
-   for (i = 0; i < n; i++)
-      printf("%f\t", x[i]);
-
-   printf("\nMaximum profit is:- %f", tp);
-
-}
-
-int main() {
-   float weight[20], profit[20], capacity;
-   int num, i, j;
-   float ratio[20], temp;
-
-   printf("\nEnter the no. of objects:- ");
-   scanf("%d", &num);
-
-   printf("\nEnter the wts and profits of each object:- ");
-   for (i = 0; i < num; i++) {
-      scanf("%f %f", &weight[i], &profit[i]);
-   }
-
-   printf("\nEnter the capacityacity of knapsack:- ");
-   scanf("%f", &capacity);
-
-   for (i = 0; i < num; i++) {
-      ratio[i] = profit[i] / weight[i];
-   }
-
-   for (i = 0; i < num; i++) {
-      for (j = i + 1; j < num; j++) {
-         if (ratio[i] < ratio[j]) {
-            temp = ratio[j];
-            ratio[j] = ratio[i];
-            ratio[i] = temp;
-
-            temp = weight[j];
-            weight[j] = weight[i];
-            weight[i] = temp;
-
-            temp = profit[j];
-            profit[j] = profit[i];
-            profit[i] = temp;
-         }
-      }
-   }
-
-   knapsack(num, weight, profit, capacity);
-   return(0);
-}

### 2015-03-11 Programming Project C/01knapsack.tex
updated test results, started framework for paper

--- 2015-02-23 Programming Project B/paper.tex
@@
 %% Title Page
-\title{Java Sorting Algorithms Comparison}
+\title{0/1 Knapsack}
 \author{Preston Stosur-Bassett}
-\date{February 23, 2015}
+\date{March 11, 2015}
 \maketitle
@@
 \pagestyle{fancy}
-\fancyhead[R]{Java Sorting Algorithms Run-Time Analysis, page \thepage}
+\fancyhead[R]{0/1 Knapsack in Java, page \thepage}
 \fancyhead[L]{Preston Stosur-Bassett}
@@
 \begin{abstract}
+%%TODO: write this
 
@@
 \section{Motivation}
-In order to show how an algorithm might run on a given set of hardware, and how the algorithm will perform when given large amounts of data, algorithms are analysed.
-Sorting algorithms sort data into a natural order.
-By analysing sorting algorithms, the fastest algorithm for a given problem can be determined.
+%%TODO: write this
+
 %% END MOTIVATION
@@
 \section{Background}
-A sorting algorithm is used to sort data with a natural order.
-One such sorting algorithm is insertion sort, which sorts by iterating through a list of data, taking the current position, and repositioning it into a more appropriate place in the list.
-Heap sort is another sorting algorithm that greatly differs than insertion sort in that it uses a divide, conquer, and combine method; meaning that it breaks the set it is sorting into subsets until the subsets can no longer be broken up and then merge sort combines the subsets together rendering the correct answer.
-%% Write this for heap sort.
+%%TODO: write this
+
 %% END BACKGROUND
@@
 \section{Procedure}
-An insertion sort can be implemented in a multitude of languages using the pseudocode provided in Algorithm 1.
-\newline
-\textbf{Insertion Sort Pre-Condition}: A is a non-empty array of data with a natural order.
-\newline
-\textbf{Insertion Sort Post-Condition}: A' is a permutation of A (containing all the same elements) in strictly non-decreasing order.
-\begin{algorithm}
-\caption {\textsc{Insertion-Sort}(A)}
-\label{algo:insertionsort}
-\begin{algorithmic}[1]
-\Procedure{Insertion-Sort}{A}
-\If{$A.length < 2$}
-\State{\Return{$A$}}
-\EndIf
-\State{$i = 2$}
-\While{$i$ upto $A.length$}
-	\State{$key = A[i]$}
-	\State{$j = i - 1$}
-	\While{$j$ downto $1$ and $key < A[j]$}
-		\State{$A[j + 1] = A[j]$}
-		\State{$j = j - 1$}
-	\EndWhile
-	\State{$A[j+1] = key$}
-	\State{$i = i + 1$}
-\EndWhile
-\State{\Return{$A$}}
-\EndProcedure
-\end{algorithmic}
-\end{algorithm}
-\newline
-\textbf{Insertion Sort Outer-Loop Invariant}: The subarray A'[1 ... i - 1] contains all the same elements as the subarray A[1 ... i - 1].
-\newline 
-\textbf{Insertion Sort Outer-Loop Initialization}: The outer-loop invariant holds because A'[1 ... i - 1] and A[1 ... i - 1] both contain the same one element.
-\newline
-\textbf{Insertion Sort Outer-Loop Maintenance}: The outer-loop invariant holds because A'[1 ... i - 1] and A[1 ... i - 1] both contain the same elements, although they may be in different orders.
-\newline
-\textbf{Insertion Sort Outer-Loop Termination}: When the outer-loop terminates, i = A.length, which implies that the entire array has been traversed and the guard has been negated.
-The negation of the guard implies that A'[1 ...
-i - 1] contains all the elements in A[1 ...
-i - 1].
-\newline
-\newline
-\textbf{Insertion Sort Inner-Loop Invariant}: A'[1 ... j] is sorted in strictly non-decreasing order.
-\newline
-\textbf{Insertion Sort Inner-Loop Initialization}: Before the first iteration of the loop, j = 1, meaning the subarray A'[1 ... j] contains exactly one element, which is already sorted.
-\newline
-\textbf{Insertion Sort Inner-Loop Maintenance}: At the beginning of each iteration of the loop the inner-loop invariant holds because j counts down from i, and A'[j+1] is swapped with A'[j] only if A'[j+1] is less than A[j].
-\newline
-\textbf{Insertion Sort Inner-Loop Termination}: The negation of the guard implies that j = A.length and that A'[1 ...
-j] has been entirely traversed and sorted in strictly non-decreasing order, which maintains the inner-loop invariant.
-\newline
-\newline
-\textbf{Insertion Sort Conclusion}: The termination of both the inner and outer loops implies that the entire array has been traversed, A' is a permutation of A containing all the same elements in strictly non-decreasing order.
-This satisfies the post condition.
-\newline
-\newline
+%%TODO: write this
 
-A merge sort can be implemented in a variety of languages using the pseudocode provided in Algorithm 2.
-\newline
-\textbf{Merge Sort Pre-Condition}: A is a non-empty array of a comparable data with a natural order.
-\newline
-\textbf{Merge Sort Post-Condition}: A' is a permutation of A (containing all the same elements) in strictly non-decreasing order.
-\newline
-\textbf{Merge Pre-Condition}: Left and right are both non-empty arrays of a comparable data type with a natural order in strictly non-decreasing order.
-\newline
-\textbf{Merge Post-Condition}: Combined has all the elements of both left and right in strictly non-decreasing order.
-\begin{algorithm}
-\caption {\textsc{Merge-Sort}(A)}
-\label{algo:mergesort}
-\begin{algorithmic}[1]
-\Procedure{Merge-Sort}{A}
-\If{$A.length < 2$}
-\State{\Return{$A$}}
-\EndIf
-\State{$mid = A.length / 2$}
-\For{$i = 1$ upto $mid$}
-	\State{$left[left.length] = A[i]$}
-\EndFor
-\For{$i = mid$ upto $A.length$}
-	\State{$right[right.length] = A[i]$}
-\EndFor
-\State{$left =$ Merge-Sort($left$)}
-\State{$right =$ Merge-Sort($right$)}
-\State{$A =$ Merge($left$, $right$)}
-\State{\Return{$A$}}
-\EndProcedure
-\newline
-\Procedure{Merge}{left, right}
-\State{$var$ $i = 0$}
-\State{$var$ $y = 0$}
-\State{$var$ $x = 0$}
-\While{$left.length$ $!=$ $i$ and $right.length$ $!=$ $y$}
-	\If{$left[i] < right[y]$}
-		\State{$combined[x] = left[i]$}
-		\State{$i = i + 1$}
-		\State{$x = x + 1$}
-	\EndIf
-	\If{$right[y] < left[i]$}
-		\State{$combned[x] = right[y]$}
-		\State{$y = y + 1$}
-		\State{$x = x + 1$}
-	\EndIf
-\EndWhile
-\While{$left.length$ $!=$ $i$}
-	\State{$combined[x] = left[i]$}
-	\State{$i = i + 1$}
-	\State{$x = x + 1$}
-\EndWhile
-\While{$right.length$ $!=$ $y$}
-	\State{$combined[x] = right[y]$}
-	\State{$y = y + 1$}
-	\State{$x = x + 1$}
-\EndWhile
-\State{\Return{$combined$}}
-\EndProcedure
-\end{algorithmic}
-\end{algorithm}
-\newline
-\textbf{Merge Sort First For Loop Invariant}: Left contains i elements, all of which can be found in A[1 ... i upto mid].
-\newline
-\textbf{Merge Sort First For Loop Invariant Initialization}: The invariant holds true vacuously because i is 0 and left contains no elements.
-\newline
-\textbf{Merge Sort First For Loop Invariant Maintenance}: The invariant holds true because i is incremented at the same rate that elements are added to left from the same i index value in A.
-\newline
-\textbf{Merge Sort First For Loop Invariant Termination}: The invariant holds true because i has been incremented at the same rate that elements have been added to left from the same i index value in A upto mid.
-\newline
-\newline
-\textbf{Merge Sort Second For Loop Invariant}: Right contains i elements, all of which can be found in A[mid ... A.length].
-\newline
-\textbf{Merge Sort Second For Loop Invariant Initialization}: The invariant holds true vacuously because i is 0 and right contains no elements.
-\newline
-\textbf{Merge Sort Second For Loop Invariant Maintenance}: The invariant holds true because i is incremented at the same rate that elements are added to right from the same i index value in A.
-\newline
-\textbf{Merge Sort Second For Loop Invariant Termination}: The invariant holds true because i has been incremented at the same rate that elements have been added to right from the same i index value in A upto to A.length.
-\newline
-\newline
-\textbf{Merge Sort Conclusion}: %%TODO: Write this
-\newline
-\newline
-\textbf{Merge Together First While Loop Invariant}: Combined contains x number of elements where x is the sum of i and y and those elements are contained in left[1 ...
-i] or right[1 ...
-y] in stricly non-decreasing order.
-\newline
-\textbf{Merge Together First While Loop Invariant Initialization}: Our invariant holds true vacuously before the first iteration of the loop because $i = 0$, $y = 0$, $i + y = 0$, $x = 0$ therefore $x = i + y$ and combined contains no elements and is vacuously in order.
-\newline
-\textbf{Merge Together First While Loop Invariant Maintenance}: Our invariant holds true at the beginning of each iteration of the loop because $x$ is incremented whenever $i$ and $y$ are incremented, and elements are added to combined from left only when $left[i] < right[y]$ and added from right only when $left[i] >= right[y]$.
-\newline
-\textbf{Merge Together First While Loop Invariant Termination}: Our invariant holds true when the loop terminates because $x$ has been incremented whenever $i$ and $y$ are incremented, and elements have only been added to combined from left when $left[i] < right[y]$ and from right only when $left[i] >= right[y]$.
-\newline
-\newline
-\textbf{Merge Together Second While Loop Invariant}: Combined contains x number of elements where $x >= i$ and those elements are contained in left[1 ... i] in stricly non-decreasing order.
-\newline
-\textbf{Merge Together Second While Loop Invariant Initialization}: The invariant holds true before the first iteration of the loop because x has been incremented when i has been incremented and elements have been added to combined from left if and only if $left[i] < right[y]$ and from right if and only if $left[i] >= right[y]$.
-\newline
-\textbf{Merge Together Second While Loop Invariant Maintenance}: The invariant holds true at the beginning of each iteration of the loop because x has been incremented when i has been incremented and elements have been added to combined from left in order.
-\newline
-\textbf{Merge Together Second While Loop Invariant Termination}: The invariant holds true at the termination of the loop because x has been incremented when i has been incremented and elements have been added to combined from left in order.
-\newline
-\newline
-\textbf{Merge Together Third While Loop Invariant}: Combined contains x number of elements where x is greather than or equal to y and those elements are contained in right[1 ...
-y] in stricly non-decreasing order.
-\newline
-\textbf{Merge Together Third While Loop Invariant Initialization}: The invariant holds true before the first initialization of the loop because x has been incremented when y has been incremented and elements have been added to combined from left if and only if $left[i] < right[y]$ and from right if and only if $left[i] >= right[y]$.
-\newline
-\textbf{Merge Together Third While Loop Invariant Maintenance}: The invariant holds true at the beginning of each iteration of the loop because x has been incremented when y has been incremented and elements have been added to combined from right in order.
-\newline
-\textbf{Merge Together Third While Loop Invariant Termination}: The invariant holds true at the termination of the loop because x has been incremented when y has been incremented and elements have been added to combined from right in order.
-\newline
-\newline
-\textbf{Merge Together Conclusion}: %%TODO: Write this.
-\newline
-\newline
-
-A heap sort can be implemented in a variety of languages using the pseudocode below in Algorithm 3.
-\newline
-\textbf{Heap Sort Pre-Condition}: A is a non-empty array with a natural order.
-\newline
-\textbf{Heap Sort Post-Condition}: A' is a permutation of A in strictly non-decreasing order.
-\newline
-\textbf{Heapify Pre-Condition}: A is a non-empty array with a comparable data type in a natural order.
-\newline
-\textbf{Heapify Post-Condition}: %%TODO: Write this.
-\newline
-\begin{algorithm}
-\caption {\textsc{Heap-Sort}(A)}
-\label{algo:heapsort}
-\begin{algorithmic}[1]
-\Procedure{Heapify}{A, i, total}
-\State{$var$ $left = i * 2$}
-\State{$var$ $right = left + 1$}
-\State{$var$ $iPrime = i$}
-\If{$left <= total$ and $A[left] > A[i]$}
-	\State{$i = left$}
-\EndIf
-\If{$right <= total$ and $A[right] > A[i]$}
-	\State{$i = right$}
-\EndIf
-\If{$i != iPrime$}
-	\State{$var$ $temp = A[iPrime]$}
-	\State{$A[iPrime] = A[i]$}
-	\State{$A[i] = tmp$}
-	\State{$A = heapify(A, i, total)$}
-\EndIf
-\State{\Return{$A$}}
-\EndProcedure
-\Procedure{Heap-Sort}{A}
-\State{$var$ $size = A.length$}
-\For{$var$ $i = size / 2$ downto $1$}
-	\State{$A = heapify(A, i, size$}
-\EndFor
-\For{$var$ $i = size$ downto $1$}
-	\State{$var$ $tmp = A[1]$}
-	\State{$A[0] = A[i]$}
-	\State{$A[i] = tmp$}
-	\State{$size = size - 1$}
-	\State{$A = heapify(A, 1, size)$}
-\EndFor
-\State{\Return{$A$}}
-\EndProcedure
-\end{algorithmic}
-\end{algorithm}
-\newline
-\textbf{Heap Sort First Loop Invariant}: A[i] is the parent element in a heap.
-\newline
-\textbf{Heap Sort First For Loop Invariant Initialization}: The invariant holds true before the first iteration of the loop because i is less than the size of the array, and therefore has a child node.
-\newline
-\textbf{Heap Sort First For Loop Invariant Maintenance}: The invariant holds true at the beginning of each iteration of the loop because A[i] will always be smaller than the size of the array, and therefore have children nodes.
-\newline
-\textbf{Heap Sort First For Loop Invariant Termination}: The invariant holds true at the termination of the loop because i will be the smallest index value of A and therefore will have children nodes.
-\newline
-\newline
-%% TODO: something about a conclusion
-\textbf{Heap Sort Second For Loop Invariant}: All elements in A at the index value greater than i are in stricly non-decreasing order.
-\newline
-\textbf{Heap Sort Second For Loop Invariant Initialization}: The invariant holds vacuously true before the first iteration of the loop because there are no elements in A that are at an index value greater than i.
-\newline
-\textbf{Heap Sort Second For Loop Invariant Maintenance}: The invariant holds true at the beginning of each iteration of the loop because each most extreme element is moved to the very end of the array at an index value greater than i.
-\newline
-\textbf{Heap Sort Second For Loop Invariant Termination}: The invariant holds true at the termination of the loop because i decreases as ecah largest element is moved to the end of A until the entire array has been traversed, so that all elements greater than i are in stirctly non-decreasing order.
-\newline
-\newline
-%% TODO: Something about a conclusion here or some shit like that
 %% END PROCEDURE
@@
 \subsection{Testing Plan and Results}
-All arrays used in testing are Java ArrayList<Integer> unless otherwise spcified.
-All times are recorded in milliseconds using a stopwatch class borrowed from Robert Sedgewock and Kevin Wayne %%TODO: Cite this shit.
-It is important to note that the soptwatch class used takes the elapsed real-time between the start of the sort algorithm and the end of the sort algorithim as opposed to taking the elapsed processor-time becasue these tests were run on a multi-core computer.
-In the table below, A denotes Array. Times in the table below are given as averages out of 10 trials.
-\newline
-\captionof{table}{Insertion Sort Test Results}
-\begin{center}
-\begin{tabular}{|l|l|l|l|}
-\hline Tested Input & Expected Results & Actual Results & Time \\
-\hline Empty A & Empty A & Empty A & 0.0003 \\
-\hline A of 1000 Strings & Sorted A 1000 Strings & Sorted A 1000 Strings & 0.021 \\
-\hline A 1 Element & Original A & Original A & 0.0003 \\
-\hline A 10 Elements & Sorted A 10 Elements & Sorted A 10 Elements & 0.0005 \\
-\hline A 100 Elements & Sorted A 100 Elements & Sorted A 100 Elements & 0.0021 \\
-\hline A 1000 Elements & Sorted A 1000 Elements & Sorted A 1000 Elements & 0.019 \\
-\hline A 10000 Elements & Sorted A 10000 Elements & Sorted A 10000 Elements & 0.129 \\
-\hline A 100000 Elements & Sorted A 100000 Elements & Sorted A 100000 Elements & 6.4923 \\
-\hline A 1000000 Elements & Sorted A 1000000 Elements & Sorted A 1000000 Elements & 2135.5007 \\
-\hline A 10000000 Elements & Sorted A 10000000 Elements & OS Crash & N/A \\
-\hline A 1000 Identical Elements & Original Array & Original Array & 0.0052 \\
-\hline
-\end{tabular}
-\end{center}
-\captionof{table}{Merge Sort Test Results}
-\begin{center}
-\begin{tabular}{|l|l|l|l|}
-\hline Tested Input & Expected Results & Actual Results & Time \\
-\hline Empty A & Empty A & Empty A & 0.0002 \\
-\hline A of 1000 Strings & Sorted A 1000 Strings & Sorted A 1000 Strings & 0.0297 \\
-\hline A 1 Element & Original A & Original A & 0.0001 \\
-\hline A 10 Elements & Sorted A 10 Elements & Sorted A 10 Elements & 0.0004 \\
-\hline A 100 Elements & Sorted A 100 Elements & Sorted A 100 Elements & 0.0028 \\
-\hline A 1000 Elements & Sorted A 1000 Elements & Sorted A 1000 Elements & 0.0294 \\
-\hline A 10000 Elements & Sorted A 10000 Elements & Sorted A 10000 Elements & 0.2098 \\
-\hline A 100000 Elements & Sorted A 100000 Elements & Sorted A 100000 Elements & 2.1904 \\
-\hline A 1000000 Elements & Sorted A 1000000 Elements & Sorted A 1000000 Elements & 22.9314 \\
-\hline A 10000000 Elements & Sorted A 10000000 Elements & Sorted A 10000000 Elements & 241.0322 \\
-\hline A 1000 Identical Elements & Original A & Original A & 0.0298 \\
-\hline
-\end{tabular}
-\end{center}
-\captionof{table}{Heap Sort Test Results}
-\begin{center}
-\begin{tabular}{|l|l|l|l|}
-\hline Tested Input & Expected Results & Actual Results & Time \\
-\hline Empty A & Empty A & Empty A & 0.0002 \\
-\hline A of 1000 Strings & Sorted A 1000 Strings & Sorted A 1000 Strings & .0188 \\
-\hline A 1 Element & Original A & Original A & 0.0003 \\
-\hline A 10 Elements & Sorted A 10 Elements & Sorted A 10 Elements & 0.0006 \\
-\hline A 100 Elements & Sorted A 100 Elements & Sorted A 100 Elements & 0.002 \\
-\hline A 1000 Elements & Sorted A 1000 Elements & Sorted A 1000 Elements & 0.0162 \\
-\hline A 10000 Elements & Sorted A 10000 Elements & Sorted A 10000 Elements & 0.0468 \\
-\hline A 100000 Elements & Sorted A 100000 Elements & Sorted A 100000 Elements & 0.889 \\
-\hline A 1000000 Elements & Sorted A 1000000 Elements & Sorted A 1000000 Elements & 23.0635 \\
-\hline A 10000000 Elements & Sorted A 10000000 Elements & Sorted A 10000000 Elements & 244.1952 \\
-\hline A 1000 Identical Elements & Original A & Original A & 0.0156 \\
-\hline
-\end{tabular}
-\end{center}
+%%TODO: write this
+
 \subsection{Problems Encountered}
-One major issue encountered during the development of this insertion sort was that after completing the sort, A' was sorted properly except for the first element in the array.
-No matter what value the first element of A had, it did not change position in A'.
-For example, if A[5, 6, 3, 4, 7] was passed to the insertion sort algorithm, the returned array would look like A'[5, 3, 4, 6, 7].
-Changing the guard for the inner for loop (see Algorithm 1 line 6) from \textit{key < A[j] and j downto 1} to \textit{j downto 1 and key < A[j]} corrected this issue.
+%%TODO: write this
 
-Another issue encountered during the development of this program was executing Java $assert$ statements.
-When assertions were wrapped in parenthesis, they evaluated to a boolean, and the assert would throw an $AssertionError$ with the explanation 'cannot compare to boolean'.
-Removing the parenthesis wrapping the assertion fixed this issue.
-
-The last issue encountered during the development process of this program was generic programming.
-The program would not compile because not all Objects have a compareTo method.
-Adding the statement $<T$ $extends$ $Comparable<T>>$ fixed this issue.
-\cite{erik}
 %% END TESTING
-
-%% BEGIN EXPERIMENTAL ANALYSIS
-\section{Experimental Analysis}
-The insertion sort demonstrated in Algorithm 1, 2, and 3 was implemented in Java and executed on an HP SpectreXT TouchSmart with a 4 core Intel i7 processor clocked at 1.9GHz running Ubuntu Gnome 14.10 64-bit.
- \begin{figure}[!]
-\begin{center}
-\includegraphics[scale=.70]{test-results.pdf}
-\end{center}
-\captionof{figure}{Merge, Heap and Insertion Sort Run-Time Comparison}
-\label{fig:comparisonfig}
-\end{figure}
-
-The expected growth of insertion sort as the number of elements (n) grows large can be represented as $ \theta(n^2) $ where $ \theta() $ represents the asymptotically tightly bound running time.
-\cite{textbook} At $n_0$ the algorithm took 0.0003 milliseconds to complete.
-At $n_1$ the algorithm also took 0.0003 milliseconds to complete.
-For both these values of $n$, the algorithm runs at $ \theta(1) $, which is a constant value because the algorithm completes before a loop is run.
-As $n$ grows larger the time to complete the experimental data correlates quite accurately with the expected growth.
-For $n_{1*10^6} $ the data matches up perfectly.
-This is to be expected, because as n grows larger the constants and lower orders of the actual running time of the insertion sort start to affect the running time less and less as the highest order of $n^2$ is so large.
-It is expected for lower values of n to not correlate well with $ \theta(n^2) $ because the system running the algorithm may be executing superfolous commands such as checking for system.
-For the data graphed above (see Fig.
-1) this was in fact the case.
-To aid this issue, averages of 10 trails were taken and used in Figure 1.
-
-The expected growth of merge sort as the number of elements (n) grows large can be represented as $ \theta(n*\lg(n)) $.
-At $n_0$ the sort took 0.0002 milliseconds to complete, however, at $n_1$ the algorithm took 0.0001 milliseconds to complete.
-The discrepency can be attributed to superfolous commands being executed by the system.
-As $n$ grows larger the time to complete correlates with the expected running time.
-It is expected for lower values of $n$ to not correlate as nicely with the expected running time because the system constant will have a larger impact on the smaller values of $n$.
-It is important to note that merge sort ran much faster than insertion sort at large values of n such as $n_{1*10^6} $ and was able to process $n_{1*10^7} $ where insertion sort crashed the system.
-The best case running time for this implementation of merge sort is $ \theta(1) $ and is elicited only when $ n <= 1 $ because if $n <= 1$ the array is vacuously sorted and can be returned immediately without executing any loop, and run at a constant time.
-
-Heap sort is very similar to merge sort in that its expected running time is $ \theta(n*\lg(n)) $.
-At lower values of $n$ heap sort ran marginally quicker than merge sort, however, as $n$ grew larger merge sort began to run marginally faster than heap sort.
-This difference is neglagable because of superfolous commands being executed by the system and other running processes.
-For all intensive purposes merge sort and heap sort have the same run time and the major difference between the two is the system constant.
-For $n_0$ heap sort completed in 0.0002 milliseconds and for $n_1$ heap sort completed in 0.0003 milliseconds.
-In both merge sort and insertion sort the sort completes before any loop is executed if $n <= 1$ however in this implementation of heap sort the algorithm runs through the entire sort.
-We can tell from this that the constant time is 0.0002 milliseconds, which correlates nicely with both merge and insertion sort.
-$ \theta(n*\lg(n)) $ is the both the best and worst case run-time for heap sort.
-%% END EXPERIMENTAL ANALYSIS
-
-%% BEGIN CONCLUSION
-\section{Conclusions}
-With a running time of $ \theta(n*\lg(n)) $ both heap sort and merge sort perform much faster than insertion sort at a running time of $ \theta(n^2) $, especially when $n$ grows large.
-While faster comptuers will be able to execute insertion sort faster, as $n$ grows large, even on a slower machine, both merge sort and heap sort will perform faster.
-Because it is easier to setup and maintain, insertion sort is a good choice when dealing with small sets of data that don't go beyond $n_{1*10^5}$, however, if data is expected to grow larger than that, either heap or merge sort is the best choice.
-%% END CONCLUSION
-
-\newpage
-
-%% BEGIN REFERENCES
-%%\section*{References}
-
-\nocite{*}
-\bibliographystyle{IEEEtran}
-\bibliography{bib}
-
-%% END REFERENCES
-
-\newpage
-
-%% BEGIN APPENDIX
-\section*{Appendix}
-\lstinputlisting[caption=Driver,
-label={Driver.java},
-breaklines=true,
-]{Code/Driver.java}
-\lstinputlisting[caption=Debug,
-label={Debug.java},
-breaklines=true,
-]{Code/Debug.java}
-\lstinputlisting[caption=DummyData,
-label={DummyData.java},
-breaklines=true,
-]{Code/DummyData.java}
-
-The class Stopwatch has not been altered from its original form.
-\newline
-\lstinputlisting[caption=Stopwatch,
-label={Stopwatch},
-breaklines=true,
-]{Code/Stopwatch.java}
-\lstinputlisting[caption=Sort,
-label={Sort.java},
-breaklines=true,
-]{Code/Sort.java}
-%% END APPENDIX
-
-%% END DOCUMENT
-\end{document}

### 2015-03-11 Programming Project C/01knapsack.tex
finished scaffolding for paper

@@
 %% END TESTING
+
+%% BEGIN EXPERIMENTAL ANALYSIS
+\section{Experimental Analysis}
+%%TODO: write this
+
+%% END EXPERIMENTAL ANALYSIS
+
+%% BEGIN CONCLUSION
+\section{Conclusion}
+%%TODO: write this
+
+%% END CONCLUSION
+
+\newpage
+
+%% BEGIN REFERNCES
+\section*{References}
+
+%% \nocite{*}
+%% \bibliographystyle{IEEEtran}
+%% \bibliography{bib}
+
+%% END REFERENCES
+
+\newpage
+
+%% BEGIN APPENDIX
+\section*{Appendix}
+\lstinputlisting[caption=Driver,
+label={Driver.java},
+breaklines=true,
+]{Code/Driver.java}
+\lstinputlisting[caption=Knapsack,
+label={Knapsack.java},
+breaklines=true,
+]{Code/Knapsack.java}
+\lstinputlisting[caption=Sort,
+label={Sort.java},
+breaklines=true,
+]{Code/Sort.java}
+
+The class Stopwatch has been altered from its original form.
+\newline
+\lstinputlisting[caption=Stopwatch,
+label={Stopwatch.java},
+breaklines=true,
+]{Code/Stopwatch.java}
+\lstinputlisting[caption=Debug,
+label={Debug.java},
+breaklines=true,
+]{Code/Debug.java}
+\lstinputlisting[caption=DummyData,
+label={DummyData.java},
+breaklines=true,
+]{Code/DummyData.java}
+%% END APPENDIX
+
+%% END DOCUMENT
+\end{document}

### 2015-03-11 Programming Project C/Code/Sort.java
wrote invariants for new sort, woo!

@@ -318,37 +318,40 @@ public class Sort<T extends Comparable<T>> {
 
 
 	/*
-	*	@Pre-Condition: TODO: write this
-	*	@Post-Condition: TODO: write this
-	*	@Description: TODO: write this
-	* @param ArrayList<ArrayList<T>> unsorted TODO: write this
-	*	@param int sortingIndex TODO: write this
-	*	@return ArrayList<ArrayList<T>> sorted TODO: write this
+	*	@Pre-Condition: ArrayList<ArrayList<T>> unsorted is an unsorted a nested non-empty ArrayList of a non-empty ArrayList (in tabular format) of a comparable data type with a natural order where sortingIndex is an index value of the nest ArrayList.
+	*	@Post-Condition: ArrayList<ArrayList<T>> will return a permutation of <code>unsorted</code> that will be in stricly non-increasing order.
+	*	@Description: insertionSortNestedArray will sort a nested ArrayList in tabular format of a comparable data type and given a specific index value of the inner ArrayList will sort the inner ArrayLists into stricly non-increasing order within the outer ArrayList
+	* @param ArrayList<ArrayList<T>> list is a non-empty unsorted nested ArrayList of ArrayList of data type T, where T is a comparable data type with a natural order.
+	*	@param int sortingIndex is an index value of the inner ArrayList to use for sorting comparisons
+	*	@return ArrayList<ArrayList<T>> list is a permutation of <code>unsorted</code> where all the elements in the outer ArrayList are sorted in stricly non-increasing order by inner ArrayLists index value of sortingIndex
 	*/
-	//There should be two loop invariants, I should be able to take them straight from above
-	public ArrayList<ArrayList<T>> insertionSortNestedArray(ArrayList<ArrayList<T>> list, int sortingIndex) {
+	//INVARIANT (Outer-Loop): The Pre-Condition implies that A[0 ... i - 1] will contain all the same data as A'[0 ... i - 1]
+	//INVARIANT (Inner-Loop): A[0 ... j] is sorted in stricly non-increasing order
+	public ArrayList<ArrayList<T>> insertionSortNestedArray(ArrayList<ArrayList<T>> unsorted, int sortingIndex) {
+		ArrayList<ArrayList<T>> list = unsorted;
 		if(list.size() > 1) {
 			int i = 1;
 
-			//Outer While Loop Initialization: TODO: write this
[…]

### 2015-03-11 Programming Project C/Code/Sort.java
added appropriate assertions

@@ -328,30 +328,52 @@ public class Sort<T extends Comparable<T>> {
 	//INVARIANT (Outer-Loop): The Pre-Condition implies that A[0 ... i - 1] will contain all the same data as A'[0 ... i - 1]
 	//INVARIANT (Inner-Loop): A[0 ... j] is sorted in stricly non-increasing order
 	public ArrayList<ArrayList<T>> insertionSortNestedArray(ArrayList<ArrayList<T>> unsorted, int sortingIndex) {
+		Debug debugger = new Debug<List<T>>();
 		ArrayList<ArrayList<T>> list = unsorted;
+
 		if(list.size() > 1) {
 			int i = 1;
 
 			/*INITIALIZATION (Outer-Loop): Before the first iteration of the loop the invariant holds beause i = 1, and there is one elment in the subarray of A[0 ... i - 1] and A'[0 ... i - 1] */
+			List<ArrayList<T>> subSortedOI = list.subList(0, i - 1);
+			List<ArrayList<T>> subUnsortedOI = unsorted.subList(0, i - 1);
+			debugger.assertEquals(subUnsortedOI, subSortedOI);
+
 			while(i < list.size()) {
 				/*MAINTENANCE (Outer-Loop): At the beginning of each iteration of the loop, the loop invariant is maintained because the subarray of A'[0 ... i - 1] contains all the same elements as A[0 ... i - 1] */
+				List<ArrayList<T>> subSortedOM = list.subList(0, i - 1);
+				List<ArrayList<T>> subUnsortedOM = unsorted.subList(0, i - 1);
+				debugger.assertEquals(subUnsortedOM, subSortedOM);
+
 				ArrayList<T> currentElement = list.get(i);
 				T value = list.get(i).get(sortingIndex);
 				int j = i - 1;
 
 				/*INITIALIZATION (Inner-Loop): Before the first iteration of the loop, j = 0, the subarray of sorted[0 ... 0] contains one element and therefore the invariants hold vacuously. */
+				List subSortedII = list.subList(0, j);
+				//debugger.assertOrder(subSortedII); //TODO: fix this?
 				while(j >= 0 && (value.compareTo(list.get(j).get(sortingIndex)) > 0)) {
 					/*MAINTENANCE (Inner-Loop): At the beginning of each iteration A[0 ... j] is sorted in stricly non-increasing order */
+					List subSortedIM = list.subList(0, j);
+					//debugger.assertOrder(subSortedIM); //TODO: fix this?
+
 					list.set(j+1, list.get(j));
[…]

### 2015-03-11 Programming Project C/Code/Debug.java
updated more documentation, added fixes to those things that needed it

@@ -72,6 +72,25 @@ public class Debug<T> {
 		}
 	}
 
+
+	/*
+	*	@Pre-Condition: <code>List<ArrayList<Integer>> actual</code> is an iterable list of ArrayList of Integer objects where sortingIndex is an index value of the ArrayList
+	*	@Post-Condition: If the LIst of ArrayList of Integer objects is in stricly non-increasing order, the program moves on normally, if not, the program halts with an <code>AssertionError</code>
+	*	@Description: runs an assertion statement against a list of ArrayList of Integer objects to ensure that for <code>k = actual.size(); A[k-2] >= A[k-1];</code>
+	*	@param List<ArrayList<Integer>> actual the list to assert is in stricly non-decreasing order
+	*	@param int sortingIndex the index value to make sorting comparisons from
+	*/
+	public void assertOrder(List<ArrayList<Integer>> actual, int sortingIndex) {
+		if(debugOn == true) {
+			int i = actual.size();
+			while(i > 1) {
+				assert actual.get(i - 1).get(sortingIndex).compareTo(actual.get(i - 2).get(sortingIndex)) <= 0;
+
+				i--;
+			}
+		}
+	}
+
 	/*
 	*	@Pre-Condition: <code>ArrayList<Integer> actual</code> is an ArrayList of Integer Objects
 	*	@Post-Condition: If the ArrayList of Integer Objects is in stricly non-decreasing order, the program moves on normally, if not, the pgoram halts with an <code>AssertionError</code>

### 2015-03-11 Programming Project C/Code/Sort.java

@@ -351,18 +351,18 @@ public class Sort<T extends Comparable<T>> {
 
 				/*INITIALIZATION (Inner-Loop): Before the first iteration of the loop, j = 0, the subarray of sorted[0 ... 0] contains one element and therefore the invariants hold vacuously. */
 				List subSortedII = list.subList(0, j);
-				//debugger.assertOrder(subSortedII); //TODO: fix this?
+				debugger.assertOrder(subSortedII, sortingIndex);
 				while(j >= 0 && (value.compareTo(list.get(j).get(sortingIndex)) > 0)) {
 					/*MAINTENANCE (Inner-Loop): At the beginning of each iteration A[0 ... j] is sorted in stricly non-increasing order */
 					List subSortedIM = list.subList(0, j);
-					//debugger.assertOrder(subSortedIM); //TODO: fix this?
+					debugger.assertOrder(subSortedIM, sortingIndex);
 
 					list.set(j+1, list.get(j));
 					j--;
 				}
 				/*TERMINATION (Inner-Loop): The negation of the gaurd implies that A'[0 ... j] has been entirely traversed and is in stricly non-increasing order. */
 				List subSortedIT = list.subList(0, j+1);
-				//debugger.assertOrder(subSortedIT); //TODO: fix this?
+				debugger.assertOrder(subSortedIT, sortingIndex);
 
 				list.set(j+1, currentElement);
 
@@ -371,7 +371,7 @@ public class Sort<T extends Comparable<T>> {
 			/*TERMINATION (Outer-Loop): When the loop terminates, i is equal to A'.length meaning the entire array has been traversed and that the guard has been negated.
 			  													The negation of the guard implies that A'[0 ... i - 1] contains all the elements of A[0 ... i - 1] */
 			List subSortedOT = list.subList(0, i - 1);
-			//debugger.assertOrder(subSortedOT); //TODO: fix this?
+			debugger.assertOrder(subSortedOT, sortingIndex);
 			debugger.assertEquals(new Integer(list.size()), new Integer(i));
 			debugger.assertEquals(unsorted, list);
 		}

### 2015-03-11 Programming Project C/Code/Knapsack.java
updated some documentation but so far most things are going to shit right now

@@ -10,18 +10,17 @@ import java.lang.Math;
 
 public class Knapsack<T extends Comparable<T>> {
   /*
-  * @Pre-Condition: TODO: write this
-  * @Post-Condition: TODO: write thsi
-  * @Description: TODO: write this
-  * @param int elems TODO: write this
-  * @param ArrayList<Integer> weights TODO: write this
-  * @param ArrayList<Integer> prices  TODO: write this
-  * @param int backpackSize TODO: write this
-  * @return int profit TODO: write this
+  * @Pre-Condition: <code>weights</code> and <code>prices</code> both have in them <code>elems</code> amount of elements
+  * @Post-Condition: The returned value will be a reasonable solution for the largest value in price where the aggregate of the corresponding weight does not excede the <code>backpackSize</code>
+  * @Description: greedyKnapsack implements a greedy algorithm to find a reasonable solution for a 0/1 Knapsack problem
+  * @param int elems is the amount of elements in both <code>ArrayList<Integer> weights</code> and <code>ArrayList<Integer> prices</code>
+  * @param ArrayList<Integer> weights is a non-empty ArrayList of Integer objects with exactly <code>elems</code> amount of elements in it and contains absolutely no zeros
+  * @param ArrayList<Integer> prices is a non-empty ArrayList of Integer object with exactly <code>elems</code> amount of elements in it
+  * @param int backpackSize is the maximum value of weights that the knapsack can hold
+  * @return int profit is a reasonable solution to the given 0/1 Knapsack problem
   */
-  //Invariants (There should be 1 or 2)
-  //For Loop Invariant: TODO: write this
-  //While Loop Invariant: TODO: write this
+  //INVARIANT (First Loop): TODO: write this
+  //INVARIANT (Second Loop): TODO: write this
   public int greedyKnapsack(int elems, ArrayList<Integer> weights,  ArrayList<Integer> prices, int backpackSize) {
     int returnValue;
     if(elems == 1) {
@@ -35,26 +34,29 @@ public class Knapsack<T extends Comparable<T>> {
 
       ArrayList<ArrayList<Integer>> ratioListings = new ArrayList<ArrayList<Integer>>(elems);
 
-      //Initialization: TODO: write this
-      for(int i = 0; i < elems; i++) {
-        //Maintenance: TODO: write this
-        int ratio = prices.get(i).intValue() / weights.get(i).intValue();
+      int v = 0;
+      /*INITIALIZATION (First Loop): */
[…]

### 2015-03-11 Programming Project C/01knapsack.tex
added pseudocode to paper

@@
 \section{Motivation}
-%%TODO: write this
+A 0/1 Knapsack problem is most often described as how does a theif steal the most valuable items with the amount of space he has available to carry them.
+However, the applications of a 0/1 Knapsack far beyond just this.
+When solved, a 0/1 Knapsack problem will render the optimal solution, whether it be how one should construct their schedule so they can get the most done in one day, or how to pack a truck such that the fewest amount of trips are taken.
 
@@
 \section{Background}
-%%TODO: write this
+Unfortunately, solving a 0/1 Knapsack problem often times proves difficult because in order to find the best solution, all possible solutions must be attempted, a method known as brute force.
+It is not pratical to brute force a problem of this kind because it would take $ \theta{n!} $ to solve, which is unacceptable for large sets of numbers.
+However, there are other methods beside brute force that allow for the correct or 'good-enought' solution to be found.
+One of these methods is known as dynamic programming, which trades off speed for space.
+With dynamic programming the correct solution can be found in $ \theta{n*m} $ time, where n is the amount of items and m is the size of the knapsack.
+Another method for solving this type of problem is called a greedy algorithm, which works by making the best decision on which item to take locally.
+A greedy solution to the 0/1 Knapsack problem traditionally runs in $ \theta{n} $ time.
+Unfortunately a greedy solution is not always correct and therefore highly unreliable.
 
@@
 \section{Procedure}
-%%TODO: write this
+A multidimentional array insertion sort can be implemented in a multitude of languages using the pseudocode provided in Algorithm 1.
+\newline
+\textbf{Insertion Sort Pre-Condition}: A is an unsorted non-empty array of non-empty arrays containing a comparable data type with a natural order such that v is an index value of the inner array.
+\newline
+\textbf{Insertion Sort Post-Condition}: A' is a permutation of A that is in strictly non-increasing order.
+\begin{algorithm}
+\caption {\textsc{Insertion-Sort}(A, v)}
+\label{algo:insertionsort}
+\begin{algorithmic}[1]
+\Procedure{Insertion-Sort}{A, v}
+\If{$A.length < 2$}
+\State{\Return{$A$}}
+\EndIf
+\State{$i = 2$}
+\While{$i$ upto $A.length$}
+	\State{$key = A[i][v]$}
+	\State{$a = A[i]$}
+	\State{$j = i - 2$}
+	\While{$j$ downto $1$ and $key > A[j][v]$}
+		\State{$A[j+1] = A[j]$}
+		\State{$j = j - 1$}
+	\EndWhile
+	\State{$A[j+1] = a$}
+	\StateP$i = i + 1$}
+\EndWhile
+\State{\Return{$A$}}
+\EndProcedure
+\end{algorithmic}
+\end{algorithm}
+\newline
+\textbf{Insertion Sort Outer-Loop Invariant}: The subarray A'[1 ... i - 1] contains all the same elements as the subarray A[1 ... i - 1].
+\newline 
+\textbf{Insertion Sort Outer-Loop Initialization}: The outer-loop invariant holds because A'[1 ... i - 1] and A[1 ... i - 1] both contain the same one element.
+\newline
+\textbf{Insertion Sort Outer-Loop Maintenance}: The outer-loop invariant holds because A'[1 ... i - 1] and A[1 ... i - 1] both contain the same elements, although they may be in different orders.
+\newline
+\textbf{Insertion Sort Outer-Loop Termination}: When the outer-loop terminates, i = A.length, which implies that the entire array has been traversed and the guard has been negated.
+The negation of the guard implies that A'[1 ...
+i - 1] contains all the elements in A[1 ...
+i - 1].
+\newline
+\newline
+\textbf{Insertion Sort Inner-Loop Invariant}: A'[1 ... j] is sorted in strictly non-decreasing order.
+\newline
+\textbf{Insertion Sort Inner-Loop Initialization}: Before the first iteration of the loop, j = 1, meaning the subarray A'[1 ... j] contains exactly one element, which is already sorted.
+\newline
+\textbf{Insertion Sort Inner-Loop Maintenance}: At the beginning of each iteration of the loop the inner-loop invariant holds because j counts down from i, and A'[j+1] is swapped with A'[j] only if A'[j+1] is less than A[j].
+\newline
+\textbf{Insertion Sort Inner-Loop Termination}: The negation of the guard implies that j = A.length and that A'[1 ...
+j] has been entirely traversed and sorted in strictly non-decreasing order, which maintains the inner-loop invariant.
+\newline
+\newline
+\textbf{Insertion Sort Conclusion}: The termination of both the inner and outer loops implies that the entire array has been traversed, A' is a permutation of A containing all the same elements in strictly non-decreasing order.
+This satisfies the post condition.
+\newline
+\newline
+
+A greedy solution to a 0/1 knapsack problem can be implemented in a variety of languages using the pseudocode in Algorithm 2.
+\newline
+\textbf{Greedy Knapsack Pre-Condition}: $Weights$ and $Prices$ both have in them $n$ number of elements
+\newline
+\textbf{Greedy Knapsack Post-Condition}: The returned value will be a resonable solution for the largest value of price combinations such that the aggregate of the corresponding weights does not excede knapsack capacity $c$.
+\begin{algorithm}
+\caption {\textsc{Greedy-Knapsack}(n, Weights, Prices, c)}
+\label{algo:greedy}
+\begin{algorithmic}[1]
+\Procedure{Greedy-Knapsack}{n, Weights, Prices, c}
+\If{$n == 1$}
+	\State{\Return{$Prices[1]$}}
+\EndIf
+\If{$n <= 0$}
+	\State{\Return{$0$}}
+\EndIf
+\State{$profit = 0$}
+\State{$ratio = new Array$}
+\For{$v = 1$ upto $n$}
+	\State{$a = Prices[v] / Weights[v]$}
+	\State{$ratio[v] = a$}
+\EndFor
+\State{$ratio = Insertion-Sort(ratio, 1)$}
+\State{$i = 1$}
+\While{$c > 0$ and $i < n$}
+	\If{$c - ratio[i][3] >= 0$}
+		\State{$profit = profit + ratio[i][2]$}
+		\State{$c = c - ratio[i][3]$}
+	\EndIf
+	\State{$i = i + 1$}
+\EndWhile
+\State{\Return{$profit$}}
+\EndProcedure
+\end{algorithmic}
+\end{algorithm}
+\newline
+%% TODO: Write invariants for greedy algorithm here
+\newline
+\newline
+
+A dynamic solution to 0/1 knapsack can be implemented in a variety of languages using the pseudocode provided in Algorithm 3.
+\newline
+\textbf{Dynamic Knapsack Pre-Condition}: $Weights$ and $Prices$ both have $n$ number of elements
+\newline
+\textbf{Dynamic Knapsack Post-Condition}: The returned value will be the correct solution for the largest value of price combinations such that the aggregate of the corresponding weights does not excede knapsack capacity $c$.
+\begin{algorithmic}
+\caption {\textsc{Dynamic-Knapsack}(n, Weights, Prices, c)}
+\label{algo:dynamic}
+\begin{algorithm}[1]
+\Procedure{Dynamic-Knapsack}{n, Weights, Prices, c}
+\If{$n == 1$}
+	\State{\Return{$Prices[1]$}}
+\EndIf
+\If{$n <= 0$}
+	\State{\Return{$0$}}
+\EndIf
+\State{$tab[n][c] = new Nested Array$}
+\For{$x = 1$ upto $n$}
+	\For{$y = 1$ upto $c$}
+		\State{$tab[x][y] = 0$}
+	\EndFor
+\EndFor
+\For{$i = 1$ upto $n$}
+	\For{$j = 0$ upto $c$}
+		\If{$Weights[i] <= j$ and $Prices[i] + tab[i][j - Weights[i]] > tab[i][j]$}
+			\State{$tab[i+1][j] = Price[i] + tab[i][j - weights[i]]$}
+		\EndIf
+		\Else
+			\State{$tab[i+1][j] = tab[i][j]$}
+		\EndElse
+	\EndFor
+\EndFor
+\State{\Return{$tab[n][c]$}}
+\EndProcedure
+\end{algorithm}
+\end{algorithmic}
+\newline
+%% TODO: Write pseudocode invariants here for Dynamic solution
+\newline
+\newline
 

### 2015-03-11 Programming Project C/01knapsack.tex
wrote more of the paper, but the important stuff is not there, its a bit out of order, how the fuck will i get this done on time?

@@
 	\State{$A[j+1] = a$}
-	\StateP$i = i + 1$}
+	\State{$i = i + 1$}
 \EndWhile
@@
 \newline
-\textbf{Greedy Knapsack Post-Condition}: The returned value will be a resonable solution for the largest value of price combinations such that the aggregate of the corresponding weights does not excede knapsack capacity $c$.
+\textbf{Greedy Knapsack Post-Condition}: The returned value will be a reasonable solution for the largest value of price combinations such that the aggregate of the corresponding weights does not exceed knapsack capacity $c$.
 \begin{algorithm}
@@
 \newline
-\textbf{Dynamic Knapsack Post-Condition}: The returned value will be the correct solution for the largest value of price combinations such that the aggregate of the corresponding weights does not excede knapsack capacity $c$.
-\begin{algorithmic}
+\textbf{Dynamic Knapsack Post-Condition}: The returned value will be the correct solution for the largest value of price combinations such that the aggregate of the corresponding weights does not exceed knapsack capacity $c$.
+\begin{algorithm}
 \caption {\textsc{Dynamic-Knapsack}(n, Weights, Prices, c)}
 \label{algo:dynamic}
-\begin{algorithm}[1]
+\begin{algorithmic}[1]
 \Procedure{Dynamic-Knapsack}{n, Weights, Prices, c}
@@
 		\EndIf
-		\Else
+		\If{$!$$(Weights[i] <= j)$ and $!$$(Prices[i] + tab[i][j - Weights[i]] > tab[i][j])$}
 			\State{$tab[i+1][j] = tab[i][j]$}
-		\EndElse
+		\EndIf
 	\EndFor
@@
 \EndProcedure
+\end{algorithmic}
 \end{algorithm}
-\end{algorithmic}
 \newline
@@
 \subsection{Testing Plan and Results}
-%%TODO: write this
+All arrays used in testing are Java ArrayList<Integer> except for $tab$ in the dynamic solution, which is a primitive Java nested array.
+All times are recorded in milliseconds using a stopwatch class borrowed from Robert Sedgwick and Kevin Wayne \cite{stopwatch}
+It is important to note that the stopwatch lass used takes the elapsed real-time between the start of the call to the knapsack solution and the end of that call as opposed to taking the elapse processor-time because these tests were run on a multi-core computer.
+In the table below, times are given as averages out of 10 trials in milliseconds.
+%% TODO: You need to implement a testing plan testing the correctness of the two algorithms!
 
 \subsection{Problems Encountered}
-%%TODO: write this
+One major issue encountered during the development of the dynamic programming solution was that when the second pair of loops would make calls to any of ArrayLists (including $tab$, which at the time was also an ArrayList implementation) there would be an out of bounds error because of a negative index value.
+The out of bounds error was eminating from the conditional in the second nested loop.
+The cause for this error is still unknown.
+To aid the issue $tab$ was changed from a nested ArrayList to a primitive nested Java array.
+
+Another issue encountered during the development of this program was executing Java $assert$ statements.
+When assertions were wrapped in parenthesis, they evaluated to a boolean, and the assert would throw an $AssertionError$ with the explanation 'cannot compare to boolean'.
+Removing the parenthesis wrapping the assertion fixed this issue.
+
+When the dynamic solution was executed with values of anything above $n = 100000$ and $c = 200000$ Java gave a "Out Of Memory Error: Java heap space".
+This can be attributed the extreme number of elements being stored in $tab$.
+The computer's system monitor reported RAM usage at 95\%.
+Unfortunately, this issue cannot be aided.
+
+The last issue encountered during the development process of this program was generic programming.
+The program would not compile because not all Objects have a compareTo method.
+Adding the statement $<T$ $extends$ $Comparable<T>>$ fixed this issue.
+\cite{erik}
 
@@
 %%TODO: write this
+\begin{figure}[!]
+\begin{center}
+\includegraphics[scale=.55]{test-results.pdf}
+\end{center}
+\captionof{figure}{Java Implementation of Greedy and Dynamic Solutions to 0/1 Knapsack Run-Time Analysis}
+\label{fig:runtime}
+\end{figure}
 

### 2015-03-11 Programming Project C/01knapsack.tex
added more to the paper

@@
 It is important to note that the stopwatch lass used takes the elapsed real-time between the start of the call to the knapsack solution and the end of that call as opposed to taking the elapse processor-time because these tests were run on a multi-core computer.
+When the number of elements was below 100, the size of the knapsack was 50; when values of 100 and greater were used the size of the knapsack was $2n$ where $n$ is the number of elements.
+For example, if the number of elements was 100, the size of the knapsack was 200, in order so the knapsack was sure to be able to make reasonable comparisons.
 In the table below, times are given as averages out of 10 trials in milliseconds.
+\newline
+\captionof{table}{Greedy Solution Real-Time Run-Time}
+\begin{center}
+\begin{tabular}{|l|l|l|}
+\hline Number of Elements & Size of Knapsack & Time \\
+\hline 0 & 50 & 0.0 \\
+\hline 1 & 50 & 0.0 \\
+\hline 10 & 50 & 0.0006 \\
+\hline %% TODO: Finish table
+\end{tabular}
+\end{center}
+%% TODO: Make table for dynamic solution
 %% TODO: You need to implement a testing plan testing the correctness of the two algorithms!

### 2015-03-11 Programming Project C/01knapsack.tex
fixed off by one issue in dynamic solution, added table in paper

@@
 \hline 10 & 50 & 0.0006 \\
-\hline %% TODO: Finish table
+\hline 100 & 200 & 0.0043 \\
+\hline 1000 & 2000 & 0.0231 \\
+\hline 10000 & 20000 & 0.252 \\
+\hline 100000 & 200000 & 59.1176 \\
+\hline
 \end{tabular}
 \end{center}
-%% TODO: Make table for dynamic solution
+\captionof{table}{Dynamic Solution Real-Time Run-Tim}
+\begin{center}
+\begin{tabular}{|l|l|l|}
+\hline 0 & 50 & 0.0001 \\
+\hline 1 & 50 & 0.0 \\
+\hline 10 & 50 & 0.0002 \\
+\hline 100 & 200 & 0.0119 \\
+\hline 1000 & 2000 & 0.0356 \\
+\hline 10000 & 20000 & 1.9243 \\
+\hline
+\end{tabular}
+\end{center}
 %% TODO: You need to implement a testing plan testing the correctness of the two algorithms!
@@
 
-When the dynamic solution was executed with values of anything above $n = 100000$ and $c = 200000$ Java gave a "Out Of Memory Error: Java heap space".
+When the dynamic solution was executed with values of anything above $n = 10000$ and $c = 20000$ Java gave a "Out Of Memory Error: Java heap space".
 This can be attributed the extreme number of elements being stored in $tab$.

### 2015-03-11 Programming Project C/Code/Driver.java

@@ -24,8 +24,21 @@ public class Driver {
 
       System.out.println("Max Knapsack Capacity: "+knapsackSize);
 
-      prices = testData.runArrayList(numberOfElements, 1, 1000, prices);
-      weights = testData.runArrayList(numberOfElements, 1, 50, weights);
+      //prices = testData.runArrayList(numberOfElements, 1, 1000, prices);
+      //weights = testData.runArrayList(numberOfElements, 1, 50, weights);
+
+      prices.add(new Integer(10));
+      prices.add(new Integer(100));
+      prices.add(new Integer(55));
+      prices.add(new Integer(12));
+      prices.add(new Integer(63));
+
+      weights.add(new Integer(45));
+      weights.add(new Integer(20));
+      weights.add(new Integer(40));
+      weights.add(new Integer(9));
+      weights.add(new Integer(50));
+
 
       System.out.println("Set P:"+prices);
       System.out.println("Set W:"+weights);
@@ -47,8 +60,20 @@ public class Driver {
 
       System.out.println("Max Knapsack Capacity: "+knapsackSize);
 
-      prices = testData.runArrayList(numberOfElements, 1, 1000, prices);
-      weights = testData.runArrayList(numberOfElements, 1, 50, weights);
+      //prices = testData.runArrayList(numberOfElements, 1, 1000, prices);
+      //weights = testData.runArrayList(numberOfElements, 1, 50, weights);
+
+      prices.add(new Integer(10));
+      prices.add(new Integer(100));
+      prices.add(new Integer(55));
+      prices.add(new Integer(12));
+      prices.add(new Integer(63));
+
+      weights.add(new Integer(45));
+      weights.add(new Integer(20));
+      weights.add(new Integer(40));
+      weights.add(new Integer(9));
+      weights.add(new Integer(50));
 
       System.out.println("Set P:"+prices);
       System.out.println("Set W:"+weights);

### 2015-03-11 Programming Project C/Code/Knapsack.java

@@ -88,6 +88,7 @@ public class Knapsack<T extends Comparable<T>> {
   //INVARIANT (Second Inner-Loop): TODO: write this
   public int dynamicKnapsack(int elems, ArrayList<Integer> weights, ArrayList<Integer> prices, int backpackSize) {
     int returnValue = 0;
+    backpackSize = backpackSize + 1;
     if(elems == 1) {
       returnValue = prices.get(0).intValue();
     }

### 2015-03-11 Programming Project C/Code/Driver.java
added a compare argument

@@ -24,21 +24,8 @@ public class Driver {
 
       System.out.println("Max Knapsack Capacity: "+knapsackSize);
 
-      //prices = testData.runArrayList(numberOfElements, 1, 1000, prices);
-      //weights = testData.runArrayList(numberOfElements, 1, 50, weights);
-
-      prices.add(new Integer(10));
-      prices.add(new Integer(100));
-      prices.add(new Integer(55));
-      prices.add(new Integer(12));
-      prices.add(new Integer(63));
-
-      weights.add(new Integer(45));
-      weights.add(new Integer(20));
-      weights.add(new Integer(40));
-      weights.add(new Integer(9));
-      weights.add(new Integer(50));
-
+      prices = testData.runArrayList(numberOfElements, 1, 1000, prices);
+      weights = testData.runArrayList(numberOfElements, 1, 50, weights);
 
       System.out.println("Set P:"+prices);
       System.out.println("Set W:"+weights);
@@ -60,20 +47,8 @@ public class Driver {
 
       System.out.println("Max Knapsack Capacity: "+knapsackSize);
 
-      //prices = testData.runArrayList(numberOfElements, 1, 1000, prices);
-      //weights = testData.runArrayList(numberOfElements, 1, 50, weights);
-
-      prices.add(new Integer(10));
-      prices.add(new Integer(100));
-      prices.add(new Integer(55));
-      prices.add(new Integer(12));
-      prices.add(new Integer(63));
-
-      weights.add(new Integer(45));
-      weights.add(new Integer(20));
-      weights.add(new Integer(40));
-      weights.add(new Integer(9));
-      weights.add(new Integer(50));
+      prices = testData.runArrayList(numberOfElements, 1, 1000, prices);
+      weights = testData.runArrayList(numberOfElements, 1, 50, weights);
 
       System.out.println("Set P:"+prices);
       System.out.println("Set W:"+weights);
@@ -86,6 +61,48 @@ public class Driver {
       Integer totalProfitObject = new Integer(totalProfit);
       System.out.println("The total profit according to this dynamic algorithm is: "+totalProfitObject);
       System.out.println("Time to Complete: "+elapsedTime);
+    }
+    else if(args[0].equals("compare") == true && args[1] != null && args[2] != null) {
+      System.out.println("Comparing greedy solution to dynamic solution...");
+
+      Stopwatch dynamicWatch = new Stopwatch();
[…]

### 2015-03-11 Programming Project C/Code/Driver.java
fixed rounding error

@@ -71,6 +71,11 @@ public class Driver {
       int numberOfElements = Integer.parseInt(args[1]);
       int knapsackSize = Integer.parseInt(args[2]);
 
+      if(numberOfElements <= 0) {
+        System.out.println("Cannot have 0 or negative number for number of elements. Changing to 1.");
+        numberOfElements = 1;
+      }
+
       System.out.println("Max Knapsack Capacity: "+knapsackSize);
 
       prices = testData.runArrayList(numberOfElements, 1, 1000, prices);
@@ -84,22 +89,22 @@ public class Driver {
       int greedyProfit = theif.greedyKnapsack(numberOfElements, weights, prices, knapsackSize);
       double greedyTime = greedyWatch.elapsedTime();
 
-      System.out.println("Greedy Solution: $"+greedyProfit);
+      System.out.println("Greedy Solution: $"+greedyProfit+". Completed in "+greedyTime+" milliseconds.");
 
       dynamicWatch.startTime();
       int dynamicProfit = theif.dynamicKnapsack(numberOfElements, weights, prices, knapsackSize);
       double dynamicTime = dynamicWatch.elapsedTime();
 
-      System.out.println("Dynamic Solution: $"+dynamicProfit);
+      System.out.println("Dynamic Solution: $"+dynamicProfit+". Completed in "+dynamicTime+" milliseconds.");
+
       System.out.println("");
 
-      int percentError = 0;
-      if(dynamicProfit != 0) {
-        int top = Math.abs(greedyProfit - dynamicProfit);
-        int bottom = Math.abs(dynamicProfit);
-        int error = top / bottom;
-        percentError = error * 100;
-      }
+      int difference = greedyProfit - dynamicProfit;
+      int top = Math.abs(difference);
+      int bottom = Math.abs(dynamicProfit);
+      double error = new Integer(top).doubleValue() / new Integer(bottom).doubleValue();
+      double exactPercentError = error * 100;
+      int percentError = new Double(exactPercentError).intValue();
 
       System.out.println("Percent Error: "+percentError+"%");
 

### 2015-03-11 Programming Project C/01knapsack.tex
added more things

@@
 All times are recorded in milliseconds using a stopwatch class borrowed from Robert Sedgwick and Kevin Wayne \cite{stopwatch}
-It is important to note that the stopwatch lass used takes the elapsed real-time between the start of the call to the knapsack solution and the end of that call as opposed to taking the elapse processor-time because these tests were run on a multi-core computer.
+It is important to note that the stopwatch class used takes the elapsed real-time between the start of the call to the knapsack solution and the end of that call as opposed to taking the elapse processor-time because these tests were run on a multi-core computer.
+%% TODO: rephrase the below paragraph
 When the number of elements was below 100, the size of the knapsack was 50; when values of 100 and greater were used the size of the knapsack was $2n$ where $n$ is the number of elements.
 For example, if the number of elements was 100, the size of the knapsack was 200, in order so the knapsack was sure to be able to make reasonable comparisons.
-In the table below, times are given as averages out of 10 trials in milliseconds.
 \newline
@@
 \begin{tabular}{|l|l|l|}
+\hline Number of Elements & Size of Knapsack & Time \\
 \hline 0 & 50 & 0.0001 \\
@@
 %% TODO: You need to implement a testing plan testing the correctness of the two algorithms!
+\newline
+\captionof{table}{Dynamic vs Greedy Solutions Accuracy Comparison}
+\begin{center}
+\begin{tabular}{|l|l|l|l}
+\hline Number of Elements & Size of Knapsack & Percent Error \\
+\hline 0 & 50 & 0 \\
+\hline 1 & 50 & 0 \\
+\hline 10 & 50 & 5.3 \\
+\hline 100 & 200 & 0.032887 \\
+\hline 1000 & 2000 & 0.032887 \\
+\hline 10000 & 20000 & 0.00555725 \\
+\end{tabular}
+\end{center}
 

### 2015-03-11 Programming Project C/Code/Driver.java

@@ -103,8 +103,7 @@ public class Driver {
       int top = Math.abs(difference);
       int bottom = Math.abs(dynamicProfit);
       double error = new Integer(top).doubleValue() / new Integer(bottom).doubleValue();
-      double exactPercentError = error * 100;
-      int percentError = new Double(exactPercentError).intValue();
+      double percentError = error * 100;
 
       System.out.println("Percent Error: "+percentError+"%");
 

### 2015-03-13 Programming Project C/Code/Knapsack.java
fixed several unwanted results, added invariants to documentation

@@ -7,6 +7,7 @@
 
 import java.util.ArrayList;
 import java.lang.Math;
+import java.util.List;
 
 public class Knapsack<T extends Comparable<T>> {
   /*
@@ -19,11 +20,12 @@ public class Knapsack<T extends Comparable<T>> {
   * @param int backpackSize is the maximum value of weights that the knapsack can hold
   * @return int profit is a reasonable solution to the given 0/1 Knapsack problem
   */
-  //INVARIANT (First Loop): TODO: write this
-  //INVARIANT (Second Loop): TODO: write this
+  //INVARIANT (First Loop): ratioListings[0 ... v-1] has the same number of non-null elements in it as both prices[0 ... v-1] and weights[0 ... v-1]
+  //INVARIANT (Second Loop): The aggregate of the weights corresponding to the prices included in the value of profit are less than or equal to the value of backpackSize
   public int greedyKnapsack(int elems, ArrayList<Integer> weights,  ArrayList<Integer> prices, int backpackSize) {
+    Debug debugger = new Debug();
     int returnValue;
-    if(elems == 1) {
+    if(elems == 1 && weights.get(0).intValue() <= backpackSize) {
       returnValue = prices.get(0).intValue();
     }
     else if(elems < 1) {
@@ -35,9 +37,15 @@ public class Knapsack<T extends Comparable<T>> {
       ArrayList<ArrayList<Integer>> ratioListings = new ArrayList<ArrayList<Integer>>(elems);
 
       int v = 0;
-      /*INITIALIZATION (First Loop): */
+      /*INITIALIZATION (First Loop): Our invariant holds before the first iteration of the loop because v = 0, and ratioListings[v-1], prices[v-1] and weights[v-1] all do not exist, and therefore is vacously true.*/
+
       while(v < elems) {
-        /*MAINTANENCE (First-Loop): */
+        /*MAINTANENCE (First-Loop): At the beginning of each iteration of the loop our invariant holds true because each time the loop runs, v is increased by one and one element is added to raioListings, therefore ratioListings[0 ... v-1] will have exactly the same number of elements as both prices[0 ... v-1] and weights[0 ... v-1].*/
+        if(v > 0) {
+          debugger.assertEquals(new Integer(weights.subList(0, v-1).size()), new Integer(prices.subList(0, v-1).size()));
+          debugger.assertEquals(new Integer(prices.subList(0, v-1).size()), new Integer(ratioListings.subList(0, v-1).size()));
+        }
+
[…]

### 2015-03-13 Programming Project C/01knapsack.tex
updated pdf
added invariants for greedy algorithm

@@
 \Procedure{Greedy-Knapsack}{n, Weights, Prices, c}
-\If{$n == 1$}
+\If{$n == 1$ and $Weights[1] <= c$}
 	\State{\Return{$Prices[1]$}}
@@
 \newline
-%% TODO: Write invariants for greedy algorithm here
+\textbf{Greedy Solution First-Loop Invariant}: ratio[1 ... v-1] has the same number of non-null elements in it as both Prices[1 ... v-1] and Weights[1 ... v-1].
+\newline
+\textbf{Greedy Solution First-Loop Invariant Initialization}: The invariant holds true before the first iteration of the loop because $v=1$, and ratio[1 ...
+v-1], Prices[1 ...
+v-1] and Weights[0 ...
+v-1] all do not exist, and is therefore vacuously true.
+\newline
+\textbf{Greedy Solution First-Loop Invariant Maintenance}: At the beginning of each iteration of the loop the invariant holds true because each time the loop runs, v is incremented by 1, and exactly 1 element is added to ratio, therefore ratio[0 ...
+v-1] will have the exact same number of elements as both Prices[1 ...
+v-1] and Weights[1 ...
+v-1].
+\newline
+\textbf{Greedy Solution First-Loop Invariant Termination}: After the termination of the loop the invariant holds true because v has been incremented by 1 during each iteration of the loop, and ratio has gained 1 element during each iteration of the loop, therefore ratio[1 ...
+v-1] contains the exact same number of elements as both Prices[1 ...
+v-1] and Weights[1 ...
+v-1].
+\newline
+\newline
+\textbf{Greedy Solution Second-Loop Invaraint}: The aggregate of the weights corresponding to the prices included in $profit$ are less than or equal to $c$.
+\newline
+\textbf{Greedy Solution Second-Loop Invariant Initialization}: Before the first iteration of the loop the invariant holds true vacuously because no prices have been included in $profit$.
+\newline
+\textbf{Greedy Solution Second-Loop Invariant Maintenance}: At the beginning of each iteration of the loop the invariant holds true because only items whose weight is less than or equal to the size of $c$ are added to $profit$.
+\newline
+\textbf{Greedy Solution Second-Loop Invariant Termination}: After the termination of the loop, the invariant holds true because only items whose weight was less than or equal to the value of $c$ have been added to $profit$.
 \newline
@@
 \Procedure{Dynamic-Knapsack}{n, Weights, Prices, c}
-\If{$n == 1$}
+\If{$n == 1$ and $Weights[1] <= c$}
 	\State{\Return{$Prices[1]$}}
@@
 It is important to note that the stopwatch class used takes the elapsed real-time between the start of the call to the knapsack solution and the end of that call as opposed to taking the elapse processor-time because these tests were run on a multi-core computer.
-%% TODO: rephrase the below paragraph
-When the number of elements was below 100, the size of the knapsack was 50; when values of 100 and greater were used the size of the knapsack was $2n$ where $n$ is the number of elements.
-For example, if the number of elements was 100, the size of the knapsack was 200, in order so the knapsack was sure to be able to make reasonable comparisons.
+The size of elements used in testing are as follows: For $n < 100$, $c = 50$. For $n >= 100$, $c = 2n$ where $n$ represents the number of items, and $c$ represents the capacity of the knapsack.
 \newline
@@
 \end{center}
-\captionof{table}{Dynamic Solution Real-Time Run-Tim}
+\captionof{table}{Dynamic Solution Real-Time Run-Time}
 \begin{center}
@@
 \end{center}
-%% TODO: You need to implement a testing plan testing the correctness of the two algorithms!
-\newline
+The percent error of greedy solutions is shown in Table 3.
+Percent error was calculated by the percent error formula: $\frac{ | Greedy Solution - Dynamic Solution | }{ | Dynamic Solution | } * 100$.
+The percentages shown in Table 3 are average percent errors taken from trials of 10.
 \captionof{table}{Dynamic vs Greedy Solutions Accuracy Comparison}
@@
 \hline Number of Elements & Size of Knapsack & Percent Error \\
-\hline 0 & 50 & 0 \\
-\hline 1 & 50 & 0 \\
-\hline 10 & 50 & 5.3 \\
-\hline 100 & 200 & 0.032887 \\
-\hline 1000 & 2000 & 0.032887 \\
-\hline 10000 & 20000 & 0.00555725 \\
+\hline 0 & 50 & 0\% \\
+\hline 1 & 50 & 0\% \\
+\hline 10 & 50 & 5.3\% \\
+\hline 100 & 200 & 0.9\% \\
+\hline 1000 & 2000 & 0.032887\% \\
+\hline 10000 & 20000 & 0.00555725\% \\
+\hline
 \end{tabular}
@@
 \section{Experimental Analysis}
-%%TODO: write this
+%%TODO: write this (make sure to mension the thing about the negative correlation between n and the answer, which can be attribute to the difference in size when numbers are larger
 \begin{figure}[!]

### 2015-03-13 Programming Project C/Code/Knapsack.java
added first set of loop invariants for dynamic solution

@@ -99,11 +99,12 @@ public class Knapsack<T extends Comparable<T>> {
   * @param int backpackSize is the maximum value of the weights that the knapsack can hold
   * @return int returnValue is the correct value of the maximum amount of values you can get from prices where their correcsponding weights do not excede the backpackSize
   */
-  //INVARIANT (First Outer-Loop): TODO: write this
-  //INVARIANT (First Inner-Loop): TODO: write this
+  //INVARIANT (First Outer-Loop): tab[0 ... x-1] has the same number of non-null elements in it as both weights[0 ... x-1] and prices[0 ... x-1]
+  //INVARIANT (First Inner-Loop): tab[x] has y number of non-null elements in it.
   //INVARIANT (Second Outer-Loop): TODO: write this
   //INVARIANT (Second Inner-Loop): TODO: write this
   public int dynamicKnapsack(int elems, ArrayList<Integer> weights, ArrayList<Integer> prices, int backpackSize) {
+    Debug debugger = new Debug();
     int returnValue = 0;
     backpackSize = backpackSize + 1;
     if(elems == 1 && weights.get(0).intValue() <= backpackSize) {
@@ -117,21 +118,51 @@ public class Knapsack<T extends Comparable<T>> {
 
       int x = 0;
       int y = 0;
-      /*INITIALIZATION (First Outer-Loop): TODO: write this */
+      int nonNullElements = 0;
+      /*INITIALIZATION (First Outer-Loop): Before the first iteration of the loop the invariant holds vacuously because tab[0 ... x-1], weights[0 ... x-1] and prices[0 ... x-1] all do not exist.*/
       while(x < elems) {
-        /*MAINTENANCE (First Outer-Loop): TODO: write this*/
-        /*INITIALIZATION (First Inner-Loop): TODO: write this */
-        while(y < backpackSize) {
-          /*MAINTENANCE (First Inner-Loop): TODO: write this */
-          tab[x][y] = 0;
+        /*MAINTENANCE (First Outer-Loop): At the beginning of each iteration of the loop the invariant holds because x is incremented by one at the same rate that one element is added to tab[].*/
+        if(x > 0) {
+          int[][] subTabM = Array.copyOfRange(tab, 0, x-1);
+          debugger.assertEquals(new Integer(weights.subList(0, x-1).intValue()), new Integer(prices.subList(0, x-1).intValue()));
+          debugger.assertEquals(new Integer(prices.subList(0, x-1).intValue()), new Integer(subTabM.length));
+        }
 
[…]

### 2015-03-13 Programming Project C/Code/Knapsack.java
changed thingy a little bit n shit

@@ -140,7 +140,7 @@ public class Knapsack<T extends Comparable<T>> {
         while(y < backpackSize) {
           /*MAINTENANCE (First Inner-Loop): At the beginning of each iteration of the loop the invariant holds true because at the same rate y is incremented an element is added to tab[x]*/
           nonNullElements = 0;
-          for(int = 0; i < tab[x].length; i++) {
+          for(int i = 0; i < tab[x].length; i++) {
             if(tab[x][y] != null) {
               nonNullElements++;
             }
@@ -164,11 +164,13 @@ public class Knapsack<T extends Comparable<T>> {
       debugger.assertEquals(new Integer(weights.subList(0, x-1).intValue()), new Integer(prices.subList(0, x-1).intValue()));
       debugger.assertEquals(new Integer(prices.subList(0, x-1).intValue()), new Integer(subTabM.length));
 
+      int i = 0;
+      int j = 0;
       // Second Outer For Loop Initialization: TODO: write this
-      for(int i = 0; i < elems - 1; i++) {
+      while(i < elems - 1) {
         // Second Outer For Loop Maintenance: TODO: write this
         // Second Inner For Loop Initialization: TODO: write this
-        for(int j = 0; j < backpackSize; j++) {
+        while(j < backpackSize) {
           // Second Inner For Loop Maintenance: TODO: write this
           if(weights.get(i).intValue() <= j && prices.get(i).intValue() + tab[i][j - weights.get(i).intValue()] > tab[i][j]) {
             tab[i+1][j] = prices.get(i).intValue() + tab[i][j - weights.get(i).intValue()];
@@ -176,8 +178,12 @@ public class Knapsack<T extends Comparable<T>> {
           else {
             tab[i+1][j] = tab[i][j];
           }
+
+          j++;
         }
         // Second Inner For Loop Termination: TODO: write this
+
+        i++;
       }
       // Second Outer For Loop Termination: TODO: write this
 

### 2015-03-13 Programming Project C/Code/Knapsack.java
finished invariants in code

@@ -101,8 +101,8 @@ public class Knapsack<T extends Comparable<T>> {
   */
   //INVARIANT (First Outer-Loop): tab[0 ... x-1] has the same number of non-null elements in it as both weights[0 ... x-1] and prices[0 ... x-1]
   //INVARIANT (First Inner-Loop): tab[x] has y number of non-null elements in it.
-  //INVARIANT (Second Outer-Loop): TODO: write this
-  //INVARIANT (Second Inner-Loop): TODO: write this
+  //INVARIANT (Second Outer-Loop): tab[i][] is larger than or equal to tab[i-1][].
+  //INVARIANT (Second Inner-Loop): tab[i][j] is always equal to or greater than tab[i][j-1]
   public int dynamicKnapsack(int elems, ArrayList<Integer> weights, ArrayList<Integer> prices, int backpackSize) {
     Debug debugger = new Debug();
     int returnValue = 0;
@@ -166,12 +166,18 @@ public class Knapsack<T extends Comparable<T>> {
 
       int i = 0;
       int j = 0;
-      // Second Outer For Loop Initialization: TODO: write this
+      /*INITIALIZATION Second Outer-Loop: Before the first iteration of the loop our invariant holds vacuously true because tab[i-1] does not exist.*/
       while(i < elems - 1) {
-        // Second Outer For Loop Maintenance: TODO: write this
-        // Second Inner For Loop Initialization: TODO: write this
+        /*MAINTENANCE Second Outer-Loop: At the beginning of each iteration of the for loop the invariant holds true because during the loop tab can only be increased, never decreased.*/
+        if(i > 0) {
+          debugger.assertGreatEquals(tab[i][j], tab[i-1][j]);
+        }
+        /*INITIALIZATION Second Inner-Loop: Before the first iteration of the for loop the invariant holds vacuously because tab[i][j-1] does not exist.*/
         while(j < backpackSize) {
-          // Second Inner For Loop Maintenance: TODO: write this
+          /*MAINTENANCE Second Inner-Loop: AT the beginning of each iteration of hte for loop the invariant holds true because during the loop tab can only be increased, never decreased.*/
+          if(j > 0){
+            debugger.assertGreatEquals(tab[i][j], tab[i][j-1]);
+          }
           if(weights.get(i).intValue() <= j && prices.get(i).intValue() + tab[i][j - weights.get(i).intValue()] > tab[i][j]) {
[…]

### 2015-03-13 Programming Project C/01knapsack.tex
minor update?

@@
 \newline
-%% TODO: Write pseudocode invariants here for Dynamic solution
+\textbf{Dynamic Solution First Outer-Loop Invariant}: tab[1 .. x-1] has the same number of non-null elements in it as both Weights[1 ... x - 1] and Prices[1 ... x-1].
+\newline
+\textbf{Dynamic Solution First Outer-Loop Invariant Initialization}: Before the first iteration of the loop the invariant holds true vacuously because tab[0 ...
+x-1], weights[0 ...
+x-1] and prices[0 ...
+x-1] all do not exist.
+\newline
+\textbf{Dynamic Solution First Outer-Loop Invariant Maintenance}: At the beginning of each iteration because x is incremented by one at the same rate that one element is added to tab.
+\newline
+\textbf{Dynamic Solution First Outer-Loop Invariant Termination}: At the termination of the loop the invariant holds true because x has been incremented at the same rate that an element has been added to tab.
+\newline
+\newline
+\textbf{Dynamic Solution First Inner-Loop Invariant}: tab[x] has y number of non-null elements in it.
+\newline
+\textbf{Dynamic Solution First Inner-Loop Invariant Initialization}: Before the first iteration of the loop the invariant holds true because y = 0 and tab[x] has 0 elements in it.
+\newline
+\textbf{Dynamic Solution First Inner-Loop Invariant Maintenance}: At the beginning of each iteration of the loop the invariant holds true because because y is incremented at the same rate that elements are added to tab[x].
+\newline
+\textbf{Dynamic Solution First Inner-Loop Invariant Termination}: After the termination of the loop the invariant holds true because elements have been added to tab[x] at the same rate in which y has been incremented.
 \newline

### 2015-03-13 Programming Project C/01knapsack.tex
finished loop invariants

@@
 \newline
+\textbf{Dynamic Solution Second Outer-Loop Invariant}: tab[i][j] is larger than or equal to tab[i-1][j]
+\newline
+\textbf{Dynamic Solution Second Outer-Loop Invariant Initialization}: Before the first iteration of the loop the invariant holds true vacuously because tab[-1] does not exist.
+\newline
+\textbf{Dynamic Solution Second Outer-Loop Invariant Maintenance}: At the beginning of each iteration of the loop the invariant holds true because during the loop the values are incremented, but they are never decreased.
+\newline
+\textbf{Dynamic Solution Second Outer-Loop Invariant Termination}: After the loop has terminated the invariant holds true because the values in tab have never been decreased, only increased, and therefore tab[i][j] will be greater than or equal to tab[i-1][j]
+\newline
+\newline
+\textbf{Dynamic Solution Second Inner-Loop Invariant}: tab[i][j] is always equal to or greater than tab[i][j-1]
+\newline
+\textbf{Dynamic Solution Second Inner-Loop Invariant Initialization}: Before the first iteration of the loop the invariant holds true vacuously because tab[i][j-1] does not exist.
+\newline
+\textbf{Dynamic Solution Second Inner-Loop Invariant Maintenance}: At the beginning of each iteration of the loop the invariant holds true because the values in tab have never been decreased, only increased, therefore tab[i][j] is going to be larger than or equal to tab[i][j-1].
+\newline
+\textbf{Dynamic Solution Second Inner-Loop Invariant Termination}: After the termination of the loop the invariant holds true because the values in tab have never been decremeneted, only incremented.
+\newline
+\newline
 

### 2015-03-13 Programming Project C/01knapsack.tex
added experimental analysis section

@@
 \section{Experimental Analysis}
-%%TODO: write this (make sure to mension the thing about the negative correlation between n and the answer, which can be attribute to the difference in size when numbers are larger
 \begin{figure}[!]
@@
 \end{figure}
+The insertion sort, greedy solution and dynamic solution demonstrated in Algorithm 1, 2 and 3 was implemented in Java and executed on an HP SpectreXT TouchSmart with a 4 core Intel i7 Ivybridge processor clocked at 1.9GHz running Ubuntu Gnome 14.10 64-bit with 8GB of DDR3 RAM.
+The expected growth of insertion sort as the number of elements (n) grows large can be represented as $ \theta{n^2} $ where $ \theta{} $ represents the asympotically tightly bound running time.
+\cite{textbook}
+
+The greedy solution also follows a similair $ \theta{n^2} $ run time, see Figure 1.
+This is because an insertion sort was chosen to implement as the sorting algorithm used in the greedy solution.
+An insertion sort was chosen for the sorting algorithm because, while it is not the fatest, requires less auxillary space.
+While run time is an important factor, the space required to run the algorithm is also very important, and should not be neglacted.
+As seen in Table 2, the dynamic solution unavoidably requried too much auxillary space and Java gave an error when values of $n$ became incresingly large.
+It also important to note that the data graphed in Figure 1 was executed while the machine executed superfolous commands such as checking for system updates.
+To aid this issue, averages of 10 trails were taken and used in Figure 1.
+
+The run time for the dynamic solution as seen in figure one $ \theta{n*c} $ where n is the number of elements and c is the size of the knapsack.
+In this implementation for values of $n$ greater than or equal to 100, a value of $2n$ was used for $c$.
+That would make the run time of the specific implementation of the dynamic solution $2n^2$ but after dropping constants becomes $ \theta{n^2} $.
+In Figure 1, the reason that the dynamic solution takes so much more time to complete is because there is a constant of 2 for $n^2$ which can be attributed to doubling the real time it took to return the results.
+Because the dynamic solution works by storing answers originally found in auxillary space, the values of $n$ cannot be very high without causing a Java error, as seen in Table 2.
+
+In Table 3 there looks to be a negative correlation between the size of $n$ and the percent error.
+This negative correlation can be attributed to the fact that the percent error formula takes the difference between the two results and divides it by the correct result.
+For example, if the result for a greedy solution was 2 and the result for the dynamic solution was 3 the formula would read $ \frac{2}{3} * 100 $ rendering a 66.66\% error, however the difference between the two is only 2.
+If the values were 101 and 103 the pecent error formula would read $ \frac{2}{103} * 100 $ and renders 1.9\% error.
+The second value is substantially lower, even though the difference between the two numbers remains the same.
 

### 2015-03-13 Programming Project C/01knapsack.tex
should be complete

@@
 \begin{abstract}
-%%TODO: write this
+The present study investigates two implementations for a 0/1 Knapsack problem, a greedy solution, and a dynamic solution.
+The dynamic solution will always result in the correct answer but takes $\theta{n*m}$ time to complete.
+The greedy solution will generally result in a reasonable answer, and even sometimes the correct answer, and take $\theta{n}$ time to complete.
+Which solution to use will depend on what is most important to the problem at hand, speed, or accuracy.
 
@@
 \section{Motivation}
-A 0/1 Knapsack problem is most often described as how does a theif steal the most valuable items with the amount of space he has available to carry them.
+A 0/1 Knapsack problem is most often described as how does a thief steal the most valuable items with the amount of space he has available to carry them.
 However, the applications of a 0/1 Knapsack far beyond just this.
@@
 Unfortunately, solving a 0/1 Knapsack problem often times proves difficult because in order to find the best solution, all possible solutions must be attempted, a method known as brute force.
-It is not pratical to brute force a problem of this kind because it would take $ \theta{n!} $ to solve, which is unacceptable for large sets of numbers.
-However, there are other methods beside brute force that allow for the correct or 'good-enought' solution to be found.
+It is not practical to brute force a problem of this kind because it would take $ \theta{n!} $ to solve, which is unacceptable for large sets of numbers.
+However, there are other methods beside brute force that allow for the correct or 'good-enough' solution to be found.
 One of these methods is known as dynamic programming, which trades off speed for space.
@@
 \section{Procedure}
-A multidimentional array insertion sort can be implemented in a multitude of languages using the pseudocode provided in Algorithm 1.
+A multidimensional array insertion sort can be implemented in a multitude of languages using the pseudocode provided in Algorithm 1.
 \newline
@@
 \newline
-\textbf{Greedy Solution Second-Loop Invaraint}: The aggregate of the weights corresponding to the prices included in $profit$ are less than or equal to $c$.
+\textbf{Greedy Solution Second-Loop Invariant}: The aggregate of the weights corresponding to the prices included in $profit$ are less than or equal to $c$.
 \newline
@@
 \newline
-\textbf{Dynamic Solution Second Inner-Loop Invariant Termination}: After the termination of the loop the invariant holds true because the values in tab have never been decremeneted, only incremented.
+\textbf{Dynamic Solution Second Inner-Loop Invariant Termination}: After the termination of the loop the invariant holds true because the values in tab have never been decremented, only incremented.
 \newline
@@
 One major issue encountered during the development of the dynamic programming solution was that when the second pair of loops would make calls to any of ArrayLists (including $tab$, which at the time was also an ArrayList implementation) there would be an out of bounds error because of a negative index value.
-The out of bounds error was eminating from the conditional in the second nested loop.
+The out of bounds error was emanating from the conditional in the second nested loop.
 The cause for this error is still unknown.
@@
 The insertion sort, greedy solution and dynamic solution demonstrated in Algorithm 1, 2 and 3 was implemented in Java and executed on an HP SpectreXT TouchSmart with a 4 core Intel i7 Ivybridge processor clocked at 1.9GHz running Ubuntu Gnome 14.10 64-bit with 8GB of DDR3 RAM.
-The expected growth of insertion sort as the number of elements (n) grows large can be represented as $ \theta{n^2} $ where $ \theta{} $ represents the asympotically tightly bound running time.
+The expected growth of insertion sort as the number of elements (n) grows large can be represented as $ \theta{n^2} $ where $ \theta{} $ represents the asymptotically tightly bound running time.
 \cite{textbook}
 
-The greedy solution also follows a similair $ \theta{n^2} $ run time, see Figure 1.
+The greedy solution also follows a similar $ \theta{n^2} $ run time, see Figure 1.
 This is because an insertion sort was chosen to implement as the sorting algorithm used in the greedy solution.
-An insertion sort was chosen for the sorting algorithm because, while it is not the fatest, requires less auxillary space.
-While run time is an important factor, the space required to run the algorithm is also very important, and should not be neglacted.
-As seen in Table 2, the dynamic solution unavoidably requried too much auxillary space and Java gave an error when values of $n$ became incresingly large.
-It also important to note that the data graphed in Figure 1 was executed while the machine executed superfolous commands such as checking for system updates.
+An insertion sort was chosen for the sorting algorithm because, while it is not the fastest, requires less axillary space.
+While run time is an important factor, the space required to run the algorithm is also very important, and should not be neglected.
+As seen in Table 2, the dynamic solution unavoidably required too much axillary space and Java gave an error when values of $n$ became increasingly large.
+It also important to note that the data graphed in Figure 1 was executed while the machine executed superfluous commands such as checking for system updates.
 To aid this issue, averages of 10 trails were taken and used in Figure 1.
@@
 In Figure 1, the reason that the dynamic solution takes so much more time to complete is because there is a constant of 2 for $n^2$ which can be attributed to doubling the real time it took to return the results.
-Because the dynamic solution works by storing answers originally found in auxillary space, the values of $n$ cannot be very high without causing a Java error, as seen in Table 2.
+Because the dynamic solution works by storing answers originally found in axillary space, the values of $n$ cannot be very high without causing a Java error, as seen in Table 2.
 
@@
 For example, if the result for a greedy solution was 2 and the result for the dynamic solution was 3 the formula would read $ \frac{2}{3} * 100 $ rendering a 66.66\% error, however the difference between the two is only 2.
-If the values were 101 and 103 the pecent error formula would read $ \frac{2}{103} * 100 $ and renders 1.9\% error.
+If the values were 101 and 103 the percent error formula would read $ \frac{2}{103} * 100 $ and renders 1.9\% error.
 The second value is substantially lower, even though the difference between the two numbers remains the same.
@@
 \section{Conclusion}
-%%TODO: write this
+Although in this specific implementation, both the dynamic solution and the greedy solution had a run-time of $\theta{n^2}$, it can be possible for the dynamic solution to have a run-time of $\theta{n*c}$ and the greedy solution to have a run-time of $\theta{n}$.
+The best algorithm to use for a given problem is really a subjective decision and depends on two things, 1.
+how accurate does the answer to the problem really need to be, and 2.
+what is the time requirement to find out what the answer is.
 
@@
 %% BEGIN REFERNCES
-\section*{References}
+%%\section*{References}
 
-%% \nocite{*}
-%% \bibliographystyle{IEEEtran}
-%% \bibliography{bib}
+\nocite{*}
+\bibliographystyle{IEEEtran}
+\bibliography{bib}
 

### 2015-03-13 Programming Project C/01knapsack.tex
final update

@@
 The present study investigates two implementations for a 0/1 Knapsack problem, a greedy solution, and a dynamic solution.
-The dynamic solution will always result in the correct answer but takes $\theta{n*m}$ time to complete.
-The greedy solution will generally result in a reasonable answer, and even sometimes the correct answer, and take $\theta{n}$ time to complete.
+The dynamic solution will always result in the correct answer but takes $\theta(n*m)$ time to complete.
+The greedy solution will generally result in a reasonable answer, and even sometimes the correct answer, and take $\theta(n)$ time to complete.
 Which solution to use will depend on what is most important to the problem at hand, speed, or accuracy.
@@
 Unfortunately, solving a 0/1 Knapsack problem often times proves difficult because in order to find the best solution, all possible solutions must be attempted, a method known as brute force.
-It is not practical to brute force a problem of this kind because it would take $ \theta{n!} $ to solve, which is unacceptable for large sets of numbers.
+It is not practical to brute force a problem of this kind because it would take $ \theta(n!) $ to solve, which is unacceptable for large sets of numbers.
 However, there are other methods beside brute force that allow for the correct or 'good-enough' solution to be found.
 One of these methods is known as dynamic programming, which trades off speed for space.
-With dynamic programming the correct solution can be found in $ \theta{n*m} $ time, where n is the amount of items and m is the size of the knapsack.
+With dynamic programming the correct solution can be found in $ \theta(n*m) $ time, where n is the amount of items and m is the size of the knapsack.
 Another method for solving this type of problem is called a greedy algorithm, which works by making the best decision on which item to take locally.
-A greedy solution to the 0/1 Knapsack problem traditionally runs in $ \theta{n} $ time.
+A greedy solution to the 0/1 Knapsack problem traditionally runs in $ \theta(n) $ time.
 Unfortunately a greedy solution is not always correct and therefore highly unreliable.
@@
 The insertion sort, greedy solution and dynamic solution demonstrated in Algorithm 1, 2 and 3 was implemented in Java and executed on an HP SpectreXT TouchSmart with a 4 core Intel i7 Ivybridge processor clocked at 1.9GHz running Ubuntu Gnome 14.10 64-bit with 8GB of DDR3 RAM.
-The expected growth of insertion sort as the number of elements (n) grows large can be represented as $ \theta{n^2} $ where $ \theta{} $ represents the asymptotically tightly bound running time.
+The expected growth of insertion sort as the number of elements (n) grows large can be represented as $ \theta(n^2) $ where $ \theta() $ represents the asymptotically tightly bound running time.
 \cite{textbook}
 
-The greedy solution also follows a similar $ \theta{n^2} $ run time, see Figure 1.
+The greedy solution also follows a similar $ \theta(n^2) $ run time, see Figure 1.
 This is because an insertion sort was chosen to implement as the sorting algorithm used in the greedy solution.
@@
 
-The run time for the dynamic solution as seen in figure one $ \theta{n*c} $ where n is the number of elements and c is the size of the knapsack.
+The run time for the dynamic solution as seen in figure one $ \theta(n*c) $ where n is the number of elements and c is the size of the knapsack.
 In this implementation for values of $n$ greater than or equal to 100, a value of $2n$ was used for $c$.
-That would make the run time of the specific implementation of the dynamic solution $2n^2$ but after dropping constants becomes $ \theta{n^2} $.
+That would make the run time of the specific implementation of the dynamic solution $2n^2$ but after dropping constants becomes $ \theta(n^2) $.
 In Figure 1, the reason that the dynamic solution takes so much more time to complete is because there is a constant of 2 for $n^2$ which can be attributed to doubling the real time it took to return the results.
@@
 \section{Conclusion}
-Although in this specific implementation, both the dynamic solution and the greedy solution had a run-time of $\theta{n^2}$, it can be possible for the dynamic solution to have a run-time of $\theta{n*c}$ and the greedy solution to have a run-time of $\theta{n}$.
+Although in this specific implementation, both the dynamic solution and the greedy solution had a run-time of $\theta(n^2)$, it can be possible for the dynamic solution to have a run-time of $\theta(n*c)$ and the greedy solution to have a run-time of $\theta(n)$.
 The best algorithm to use for a given problem is really a subjective decision and depends on two things, 1.
